### 2012-02-27 article.tex
Making the thing compile...

% PLEASE USE THIS FILE AS A TEMPLATE
% Check file iosart2c.tex for more examples
%
% Journal:
%   Journal of Ambient Intelligence and Smart Environments (jaise)
%   Web Intelligence and Agent Systems: An International Journal (wias)
%   Semantic Web: Interoperability, Usability, Applicability (SW)
% IOS Press
% Latex 2e

% options: jaise|wias|sw
% add. options: [seceqn,secfloat,secthm,crcready,onecolumn]

\documentclass[sw]{iosart2c}

%\documentclass[sw]{iosart2c}
%\documentclass[wias]{iosart2c}
%\documentclass[jaise]{iosart2c}

\usepackage[T1]{fontenc}
\usepackage{times}%
\usepackage{natbib}% for bibliography sorting/compressing
%\usepackage{amsmath}
%\usepackage{endnotes}
%\usepackage{graphics}

%%%%%%%%%%% Put your definitions here

%%%%%%%%%%% End of definitions

\pubyear{0000}
\volume{0}
\firstpage{1}
\lastpage{1}

\begin{document}

\begin{frontmatter}

%\pretitle{}
\title{The ChEMBL database as Linked Open Data}
\runningtitle{ChEMBL-RDF}
%\subtitle{}

%\review{}{}{}

% For one author:
\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Huh?}}
\address[A]{Maastricht University}
\runningauthor{E.L. Willighagen et al.}

% Two or more authors:
%\author[A]{\fnms{} \snm{}\thanks{}},
%\author[B]{\fnms{} \snm{}}
%\runningauthor{}
%\address[A]{}
%\address[B]{}

\begin{abstract}
An abstract,
\end{abstract}

\begin{keyword}
 \sep Resource Description Framework
\end{keyword}

\end{frontmatter}

%%%%%%%%%%% The article body starts:

%\section{}\label{s1}

%\subsection{}\label{s1.1}

%\begin{figure}[t]
%\includegraphics{}
%\caption{Figure caption.}\label{f1}
%\end{figure}

%\begin{table*}
%\caption{} \label{t1}
%\begin{tabular}{lll}
%\hline
%&&\\
%&&\\
%\hline
%\end{tabular}
%\end{table*}

%%%%%%%%%%% The bibliography starts:
%\begin{thebibliography}{9}
%
%\bibitem{r1}
%
%\bibitem{r2}
%
%\end{thebibliography}

\end{document}

### 2012-02-27 article.tex
Some text to give some idea where this is going

@@
 \begin{abstract}
-An abstract,
+This paper describes the continued conversion of ChEMBL data into RDF triples, using a few existing
+ontologis, and demonstrates a few new use cases, taking advantage of its linked nature.
 \end{abstract}
@@
 \begin{keyword}
- \sep Resource Description Framework
+% \sep 
+Resource Description Framework, ChEMBL
 \end{keyword}
@@
 
-%\section{}\label{s1}
+\section{Introduction}\label{s1}
+
+Linked Data, pharmaceutical research, Open PHACTS, ...
+
+This data set has been previously used in~\cite{} and \cite{} and some used practices have
+been described in \cite{}...
 
 %\subsection{}\label{s1.1}
+
+\section{Methods}\label{s2}
+
+The ChEMBL version used in this paper is ChEMBL 12, which was released on ....
+
+\subsection{Ontologies}
+
+Ontologies... CHEMINF, ...
+
+The BIBO ontology is used to represent documents.
+
+\subsection{Use cases}
+
+John/Anna: common statistics with SPARQL
+
+Ola: use in BC-DS for possible adverse reactions
+
+Andra: CitedIn
+
+The link between data and formal publication is important in many areas of
+attribution, scientist ranking, etc, as outline in \cite{Waagmeester2012}.
+ChEMBL contains many literature references, and we wish to query this data
+for CitedIn.
+
+Peter: Bio2RDF
+
+David: Chem2Bio2RDF
+
+\section{Results}\label{s3}
+
+Previously, a SPARQL end point has been set up at Uppsala University.
+
+Later, to satisfy the requirements for the Linked Open Drug Data \cite{}
+network, the data was also hosted on Kasabi, a triple store hosting service
+free for Open Data~\cite{}.
+
+\# triples
+\# links out
+\# problems
+
+For each of the common resource classes a triple pattern was defined, following the
+data available in the relational database.
+
+Assays .... anything
+
+\begin{verbatim}
+\end{verbatim}
+
+Targets ... protein, other
+
+\begin{verbatim}
+\end{verbatim}
+
+Activities ...
+
+\begin{verbatim}
+\end{verbatim}
+
+Compounds ... Three types of compounds, peptides, small chemicals, ...
+
+\begin{tiny}
+\begin{verbatim}
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2000/01/rdf-schema#label> "Bis(3-[14-Benzyl-11-(1H-indol-3-ylmethyl)-2-isobutyl-13-methyl-5-(2-methylsulfanyl-ethyl)-3,6,9,12,15,20-hexaoxo-1,4,7,10,13,16-hexaaza-bicyclo[15.2.1]icos-18-en-8-yl]-propionamide)" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2000/01/rdf-schema#subClassOf> <http://semanticscience.org/resource/CHEMINF_000000> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2002/07/owl#equivalentClass> <http://bio2rdf.org/chebi:100041> .
+<http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL406142> <http://www.w3.org/2002/07/owl#equivalentClass> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2002/07/owl#equivalentClass> <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL406142> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2000/01/rdf-schema#label> "CHEMBL406142" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.blueobelisk.org/chemistryblogs/smiles> "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://semanticscience.org/resource/CHEMINF_000200> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/smiles> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/smiles> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://semanticscience.org/resource/CHEMINF_000018> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/smiles> <http://semanticscience.org/resource/SIO_000300> "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.blueobelisk.org/chemistryblogs/inchi> "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://semanticscience.org/resource/CHEMINF_000200> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchi> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchi> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://semanticscience.org/resource/CHEMINF_000113> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchi> <http://semanticscience.org/resource/SIO_000300> "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2002/07/owl#equivalentClass> <http://rdf.openmolecules.net/?InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.blueobelisk.org/chemistryblogs/inchikey> "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://semanticscience.org/resource/CHEMINF_000200> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchikey> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchikey> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://semanticscience.org/resource/CHEMINF_000059> .
+<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchikey> <http://semanticscience.org/resource/SIO_000300> "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
+\end{verbatim}
+\end{tiny}
+
+For documents little information is replicated from the database, taking advantage
+of PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs). URIs for the
+latter can be resolved online, providing curated information on the identified
+documents. For PubMed articles ...
+
+\begin{tiny}
+\begin{verbatim}
+<http://data.kasabi.com/dataset/chembl-rdf/journal/j6c706049c2e08871b7c46a6528065736> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://purl.org/ontology/bibo/Journal> .
+<http://data.kasabi.com/dataset/chembl-rdf/journal/j6c706049c2e08871b7c46a6528065736> <http://purl.org/dc/elements/1.1/title> "J. Med. Chem." .
+
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://purl.org/ontology/bibo/Article> .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/pmid> "14695813" .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://www.w3.org/2000/01/rdf-schema#seeAlso> <http://bio2rdf.org/pubmed:14695813> .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/dc/elements/1.1/date> "2004" .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/volume> "47" .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/issue> "1" .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/pageStart> "1" .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/pageEnd> "9" .
+<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/dc/elements/1.1/isPartOf> <http://data.kasabi.com/dataset/chembl-rdf/journal/j6c706049c2e08871b7c46a6528065736> .
+\end{verbatim}
+\end{tiny}
+
+\# results of use cases
+
+\section{Discussion}
+
+outlook ...
+
+\section{Acknowledgements}
+
+Kasabi.
 
@@
 
+\printbibliography
+
 \end{document}

### 2012-03-01 article.tex
Added Ola as author
Test commit by SSH
Some further content...

@@
 % Two or more authors:
-%\author[A]{\fnms{} \snm{}\thanks{}},
+\author[A]{\fnms{Ola} \snm{Spjuth}\thanks{}},
 %\author[B]{\fnms{} \snm{}}
@@
 
+Throughout this paper, various prefixes are used to simplify RDF output, outlined
+in Table~\ref{namespaces}.
+
 \subsection{Use cases}
@@
 
-Ola: use in BC-DS for possible adverse reactions
+Ola: Use in BC-DS for possible adverse reactions.
 
@@
 
+\begin{small}
 \begin{verbatim}
+act:a31863
+      a       chembl:Activity ;
+      cito:citesAsDataSource
+              res:r6424 ;
+      chembl:forMolecule mol:m180094 ;
+      chembl:onAssay assay:a54505 ;
+      chembl:relation ">" ;
+      chembl:standardUnits
+              "nM" ;
+      chembl:standardValue
+              "100000"^^xsd:float ;
+      chembl:type "IC50" .
 \end{verbatim}
+\end{small}
 
@@
 
-%\begin{table*}
-%\caption{} \label{t1}
-%\begin{tabular}{lll}
-%\hline
-%&&\\
-%&&\\
-%\hline
-%\end{tabular}
-%\end{table*}
+\begin{table*}
+\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
+\begin{tabular}{ll}
+\hline
+Common & Vocabularies \\
+cito & Citation Typing Ontology~\cite{ont:cito} \\
+     & http://purl.org/spar/cito/ \\
+
+\hline
+ChEMBL-RDF & Namespaces\\
+act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
+chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
+res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
+assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
+mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
+\hline
+\end{tabular}
+\end{table*}
 
@@
 
-\printbibliography
+% \printbibliography
 

### 2012-03-02 article.tex
add Bio2RDF paragraph and bibliography entries for three citations

@@
 %%%%%%%%%%% Put your definitions here
+
+\bibliographystyle{plain}
 
@@
 
-Peter: Bio2RDF
+The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011}, and access to SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
+A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases.
+These links are vital to provide context for users who need to query across both Chembl and its related databases.
+The Linked Data server used by Bio2RDF has been configured for Chembl to provide URL based services for text and link searches.
 
@@
 
-Later, to satisfy the requirements for the Linked Open Drug Data \cite{}
+Later, to satisfy the requirements for the Linked Open Drug Data \cite{Samwald2011}
 network, the data was also hosted on Kasabi, a triple store hosting service
@@
 
+\bibliography{article}
+
 %%%%%%%%%%% The bibliography starts:

### 2012-03-02 article.tex
N3 syntax for examples

@@
 The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011}, and access to SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
-A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases.
+A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases \cite{TheUniProtConsortium2010}.
 These links are vital to provide context for users who need to query across both Chembl and its related databases.
@@
 \begin{verbatim}
+assay:a17
+ a chembl:Assay ;
+ chembl:extractedFrom
+  res:r11347 ;
+ chembl:hasAssayType chembl:ADMET ;
+ chembl:hasConfScore "7"^^xsd:int ;
+ chembl:hasDescription
+  "Inhibition of cytochrome P450 progesterone 15-alpha hydroxylase" ;
+ chembl:hasTarget target:t100122 .
+\end{verbatim}
+\end{small}
+
+Targets ... protein, other
+
+\begin{small}
+\begin{verbatim}
+target:t1
+ a chembl:Target ;
+ rdfs:subClassOf pro:PR_000000001 ;
+ rdfs:label "Glucoamylase" , 
+  "Maltase-glucoamylase, intestinal" ;
+ dc:identifier "uniprot:O43451" , "3.2.1.3" ;
+ dc:title "Maltase-glucoamylase" ;
+ chembl:classL1 "Enzyme" ;
+ chembl:hasDescription
+  "Maltase-glucoamylase, intestinal" ;
+ chembl:hasKeyword "Glycosidase" , 
+  "Membrane" , "Sulfation" ;
+ chembl:hasTaxonomy
+  <http://bio2rdf.org/taxonomy:9606> ;
+ chembl:organism "Homo sapiens" ;
+ = <http://bio2rdf.org/uniprot:O43451> , 
+   <http://bio2rdf.org/ec:3.2.1.3> .
+\end{verbatim}
+\end{small}
+
+Activities ...
+
+\begin{small}
+\begin{verbatim}
 act:a31863
@@
 
-Targets ... protein, other
-
-\begin{verbatim}
-\end{verbatim}
-
-Activities ...
-
-\begin{verbatim}
-\end{verbatim}
-
 Compounds ... Three types of compounds, peptides, small chemicals, ...
@@
 \begin{verbatim}
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2000/01/rdf-schema#label> "Bis(3-[14-Benzyl-11-(1H-indol-3-ylmethyl)-2-isobutyl-13-methyl-5-(2-methylsulfanyl-ethyl)-3,6,9,12,15,20-hexaoxo-1,4,7,10,13,16-hexaaza-bicyclo[15.2.1]icos-18-en-8-yl]-propionamide)" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2000/01/rdf-schema#subClassOf> <http://semanticscience.org/resource/CHEMINF_000000> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2002/07/owl#equivalentClass> <http://bio2rdf.org/chebi:100041> .
-<http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL406142> <http://www.w3.org/2002/07/owl#equivalentClass> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2002/07/owl#equivalentClass> <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL406142> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2000/01/rdf-schema#label> "CHEMBL406142" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.blueobelisk.org/chemistryblogs/smiles> "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://semanticscience.org/resource/CHEMINF_000200> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/smiles> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/smiles> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://semanticscience.org/resource/CHEMINF_000018> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/smiles> <http://semanticscience.org/resource/SIO_000300> "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.blueobelisk.org/chemistryblogs/inchi> "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://semanticscience.org/resource/CHEMINF_000200> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchi> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchi> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://semanticscience.org/resource/CHEMINF_000113> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchi> <http://semanticscience.org/resource/SIO_000300> "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.w3.org/2002/07/owl#equivalentClass> <http://rdf.openmolecules.net/?InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://www.blueobelisk.org/chemistryblogs/inchikey> "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41> <http://semanticscience.org/resource/CHEMINF_000200> <http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchikey> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchikey> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://semanticscience.org/resource/CHEMINF_000059> .
-<http://data.kasabi.com/dataset/chembl-rdf/molecule/m41/inchikey> <http://semanticscience.org/resource/SIO_000300> "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
+m41:inchikey
+      a       cheminf:CHEMINF_000059 ;
+      cheminf:SIO_000300 "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
+
+m41:smiles
+      a       cheminf:CHEMINF_000018 ;
+      cheminf:SIO_000300 "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" .
+
+m41:inchi
+      a       cheminf:CHEMINF_000113 ;
+      cheminf:SIO_000300 "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" .
+
+chemblid:CHEMBL406142
+      owl:equivalentClass mol:m41 .
+
+mol:m41
+      rdfs:label "CHEMBL406142" , "Bis(3-[14-Benzyl-11-(1H-indol-3-ylmethyl)-2-isobutyl-13-methyl-5-(2-methylsulfanyl-ethyl)-3,6,9,12,15,20-hexaoxo-1,4,7,10,13,16-hexaaza-bicyclo[15.2.1]icos-18-en-8-yl]-propionamide)" ;
+      rdfs:subClassOf cheminf:CHEMINF_000000 ;
+      cheminf:CHEMINF_000200
+              m41:inchikey , m41:smiles , m41:inchi ;
+      <http://www.blueobelisk.org/chemistryblogs/inchi>
+              "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" ;
+      <http://www.blueobelisk.org/chemistryblogs/inchikey>
+              "LMCOMIDLRGMFCZ-RIPOXUOASA-N" ;
+      <http://www.blueobelisk.org/chemistryblogs/smiles>
+              "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" ;
+      owl:equivalentClass <http://rdf.openmolecules.net/?InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1> , <http://bio2rdf.org/chebi:100041> , chemblid:CHEMBL406142 .
 \end{verbatim}
@@
 \begin{verbatim}
-<http://data.kasabi.com/dataset/chembl-rdf/journal/j6c706049c2e08871b7c46a6528065736> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://purl.org/ontology/bibo/Journal> .
-<http://data.kasabi.com/dataset/chembl-rdf/journal/j6c706049c2e08871b7c46a6528065736> <http://purl.org/dc/elements/1.1/title> "J. Med. Chem." .
+journal:j6c706049c2e08871b7c46a6528065736
+ a bibo:Journal ;
+ dc:title "J. Med. Chem." .
 
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://purl.org/ontology/bibo/Article> .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/pmid> "14695813" .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://www.w3.org/2000/01/rdf-schema#seeAlso> <http://bio2rdf.org/pubmed:14695813> .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/dc/elements/1.1/date> "2004" .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/volume> "47" .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/issue> "1" .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/pageStart> "1" .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/ontology/bibo/pageEnd> "9" .
-<http://data.kasabi.com/dataset/chembl-rdf/resource/r1> <http://purl.org/dc/elements/1.1/isPartOf> <http://data.kasabi.com/dataset/chembl-rdf/journal/j6c706049c2e08871b7c46a6528065736> .
+res:r1
+ a bibo:Article ;
+ rdfs:seeAlso <http://bio2rdf.org/pubmed:14695813> ;
+ dc:date "2004" ;
+ dc:isPartOf journal:j6c706049c2e08871b7c46a6528065736 ;
+ bibo:issue "1" ;
+ bibo:pageEnd "9" ;
+ bibo:pageStart "1" ;
+ bibo:pmid "14695813" ;
+ bibo:volume "47" .
 \end{verbatim}

### 2012-03-02 article.tex
Filling in more details: more used ontologies, ChEMBL 13, more authors

@@
 % For one author:
-\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Huh?}}
+\author[A]{\fnms{Egon} \snm{Willighagen}} % \thanks{Huh?}}
 \address[A]{Maastricht University}
@@
 
-% Two or more authors:
-\author[A]{\fnms{Ola} \snm{Spjuth}\thanks{}},
+% Two or more authors (order to be determined):
+\author[B]{\fnms{Ola} \snm{Spjuth}},
+\address[B]{Uppsala University}
+
+\author[C]{\fnms{Peter} \snm{Ansell}},
+\address[C]{xxx}
+
+\author[D]{\fnms{Antony} \snm{Williams}},
+\address[D]{xxx}
+
+\author[E]{\fnms{John} \snm{Overington}},
+\author[E]{\fnms{Anna} \snm{Gaulton}},
+\address[E]{European Bioinformatics Institute}
+
 %\author[B]{\fnms{} \snm{}}
@@
 %\address[A]{}
-%\address[B]{}
 
@@
 
-This data set has been previously used in~\cite{} and \cite{} and some used practices have
-been described in \cite{}...
+This data set has been previously used in~\cite{Willighagen2011} ...
 
@@
 
-The ChEMBL version used in this paper is ChEMBL 12, which was released on ....
+The ChEMBL version used in this paper is ChEMBL 13, which was released on February 29 2011....
 
@@
 network, the data was also hosted on Kasabi, a triple store hosting service
-free for Open Data~\cite{}.
+free for Open Data~\cite{kasabi}.
 
@@
               "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" ;
-      <http://www.blueobelisk.org/chemistryblogs/inchikey>
-              "LMCOMIDLRGMFCZ-RIPOXUOASA-N" ;
-      <http://www.blueobelisk.org/chemistryblogs/smiles>
-              "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" ;
       owl:equivalentClass <http://rdf.openmolecules.net/?InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1> , <http://bio2rdf.org/chebi:100041> , chemblid:CHEMBL406142 .
@@
 Common & Vocabularies \\
-cito & Citation Typing Ontology~\cite{ont:cito} \\
+cito    & Citation Typing Ontology~\cite{Shotton2010} \\
         & http://purl.org/spar/cito/ \\
+cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
+        & http://semanticscience.org/resource/ \\
+bibo    & Bibliography Ontology~\cite{Giasson2011} \\
+        & http://purl.org/ontology/bibo/ \\
 

### 2012-03-02 article.tex
Restructuring: use cases around chosen technologies

@@
 
-This data set has been previously used in~\cite{Willighagen2011} ...
+This data set has been previously used in~\cite{Willighagen2011} and in Foo where the data is used
+in the Foo framework (book chapter...)
 
@@
 The ChEMBL version used in this paper is ChEMBL 13, which was released on February 29 2011....
+
+The RDF triples were generated from the a local MySQL chembl\_13 database populated with a stock
+SQL dump of ChEMBL 13. PHP scripts were used to create N-Triples files. The source code is
+available under a XXX license from https://github.com/egonw/chembl.rdf.
 
@@
 
-John/Anna: common statistics with SPARQL
+A few new uses cases have been developed for this paper to demonstrate not just the intrinsic
+value of the data in the ChEMBL database as done previously, but also to show the possibilities
+the Linked Data nature offers.
 
-Ola: Use in BC-DS for possible adverse reactions.
-
-Andra: CitedIn
-
-The link between data and formal publication is important in many areas of
-attribution, scientist ranking, etc, as outline in \cite{Waagmeester2012}.
-ChEMBL contains many literature references, and we wish to query this data
-for CitedIn.
-
-The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011}, and access to SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
-A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases \cite{TheUniProtConsortium2010}.
-These links are vital to provide context for users who need to query across both Chembl and its related databases.
-The Linked Data server used by Bio2RDF has been configured for Chembl to provide URL based services for text and link searches.
-
-David: Chem2Bio2RDF
+... describe what we want to show ...
 
@@
 
-Previously, a SPARQL end point has been set up at Uppsala University.
-
-Later, to satisfy the requirements for the Linked Open Drug Data \cite{Samwald2011}
-network, the data was also hosted on Kasabi, a triple store hosting service
-free for Open Data~\cite{kasabi}.
+\subsection{Data Statistics}
 
@@
 \# problems
+
+The full set of triples are available from http://semantics.bigcat.unimaas.nl/chembl12/.
+
+FIXME!
+
+\subsection{Data Structure}
 
@@
 
-\# results of use cases
+\subsection{SPARQL end points}
+
+Previously, a SPARQL end point has been set up at Uppsala University.
+
+\subsubsection{Bio2RDF}
+
+Peter: write up the proxying... if that still works...
+
+The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011}, and access to SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
+A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases \cite{TheUniProtConsortium2010}.
+These links are vital to provide context for users who need to query across both Chembl and its related databases.
+The Linked Data server used by Bio2RDF has been configured for Chembl to provide URL based services for text and link searches.
+
+\subsubsection{Data Statistics and Validation}
+
+John/Anna: common statistics with SPARQL
+
+%\subsubsection{CitedIn}
+
+%Andra: use the SPARQL end point to find which entries in ChEMBL cite what papers
+
+%The link between data and formal publication is important in many areas of
+%attribution, scientist ranking, etc, as outline in \cite{Waagmeester2012}.
+%ChEMBL contains many literature references, and we wish to query this data
+%for CitedIn.
+
+\subsubsection{Decision Support}
+
+Egon,Antony/Valery,Ola: develop and write up the Bioclipse Decision Support use case
+
+\subsection{Linked Data on Kasabi}
+
+Later, to satisfy the requirements for the Linked Open Drug Data \cite{Samwald2011}
+network, the data was also hosted on Kasabi, a triple store hosting service
+free for Open Data~\cite{kasabi}.
+
+David: Chem2Bio2RDF
+
+%\subsection{Use Case 2: Chem2Bio2RDF}
+
+%David: Chem2Bio2RDF
+
+\subsubsection{Linking out to ChemSpider}
+
+Antony: if you can contribute ChemSpider IDs for the ChEMBL compounds ...
+
+\subsubsection{Linking out to ChEMBL}
+
+John/Anna: I can easily add links to webpages using foaf:homepage... are webpage URLs for entities stable enough?
+
+\subsubsection{Linking out to Bio2RDF}
+
+Peter/Egon: write up links to Bio2RDF for protein, pubmed, ...
 
@@
 
-Kasabi.
+Tim and Foo of Kasabi. What S.H. Name of the Bioinformatics department at Uppsala University.
 

### 2012-03-05 article.tex
Added a diagram showing the data structures in the data
Reordered author details
Completer address

@@
 %\usepackage{endnotes}
-%\usepackage{graphics}
+\usepackage{graphicx}
+%\usepackage{tikz}
+%\usetikzlibrary{snakes,arrows,shapes}
 
@@
 
-\bibliographystyle{plain}
+% \bibliographystyle{plain} 
+\bibliographystyle{unsrt}
 
@@
 % For one author:
-\author[A]{\fnms{Egon} \snm{Willighagen}} % \thanks{Huh?}}
-\address[A]{Maastricht University}
-\runningauthor{E.L. Willighagen et al.}
-
+\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Corresponding author.\\ E-mail: egon.willighagen@maastrichtuniversity.nl.}}
 % Two or more authors (order to be determined):
 \author[B]{\fnms{Ola} \snm{Spjuth}},
-\address[B]{Uppsala University}
-
 \author[C]{\fnms{Peter} \snm{Ansell}},
-\address[C]{xxx}
-
 \author[D]{\fnms{Antony} \snm{Williams}},
-\address[D]{xxx}
-
 \author[E]{\fnms{John} \snm{Overington}},
 \author[E]{\fnms{Anna} \snm{Gaulton}},
+
+\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands\\
+E-mail: egon.willighagen@maastrichtuniversity.nl}
+\runningauthor{E.L. Willighagen et al.}
+\address[B]{Uppsala University}
+\address[C]{University of Queensland}
+\address[D]{xxx}
 \address[E]{European Bioinformatics Institute}
-
-%\author[B]{\fnms{} \snm{}}
-%\runningauthor{}
-%\address[A]{}
 
@@
 
-%\begin{figure}[t]
-%\includegraphics{}
-%\caption{Figure caption.}\label{f1}
-%\end{figure}
+\begin{figure}[t]
+\includegraphics[width=0.5\textwidth]{figs/relations}
+\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
+of common classes, while others are instances.}\label{f1}
+\end{figure}
 
@@
 
-% \printbibliography
-
 \end{document}

### 2012-03-05 article.tex
Added a basic, draft introduction

@@
 
-Linked Data, pharmaceutical research, Open PHACTS, ...
+Despite the data deluge, finding new, unique, but significant patterns explaining biological
+phenomena not yet understood is limited by the availability of data. Such data analysis is
+beyond the scope is single data sets; it requires data integration from many sources, for
+example, systems biology integrating micro-array differential expression to biological
+pathways~\cite{}. Another prominent example is drug discovery where a new unique chemical entity is
+searched with just the right biological properties. This too requires linking many
+data sets~\cite{Samwald2011,OpenPHACTS}.
 
-This data set has been previously used in~\cite{Willighagen2011} and in Foo where the data is used
-in the Foo framework (book chapter...)
+ChEMBL contains biological activities for over a million chemical entities, extracted from
+literature, providing a unique resource for drug discovers~\cite{Warr2010}. Originally, this
+data was available for download and via a web interface. The former requires setting up
+the data in a relational database, while the latter limits the machine access to the data.
+Two independent teams created RDF triples for the ChEMBL data, resulting in exposure
+via Chem2Bio2RDF~\cite{} and ChEMBL-RDF~\cite{Willighagen2011}.
+
+This paper presents the history of the ChEMBL-RDF data set, details of the latest structures
+and ontologies used for the RDF version of ChEMBL 13, and a few new example use cases,
+exemplifying how the data set can be linked to other data sources to further support
+research in the life sciences.
 

### 2012-03-05 article.tex
add my institution
A bit more drafting...

@@
 
-%\subsection{}\label{s1.1}
-
 \section{Methods}\label{s2}
 
-The ChEMBL version used in this paper is ChEMBL 13, which was released on February 29 2011....
+The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2011.
+RDF triples were created from a local MySQL database to which the SQL data dump of ChEMBL was
+inserted, using a PHP script, available from source code hosting service
+GitHub~\citep{ChEMBLRDFGitHub}, a process outlined earlier in a best practices note by
+the international Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.
 
-The RDF triples were generated from the a local MySQL chembl\_13 database populated with a stock
-SQL dump of ChEMBL 13. PHP scripts were used to create N-Triples files. The source code is
-available under a XXX license from https://github.com/egonw/chembl.rdf.
-
-\subsection{Ontologies}
-
-Ontologies... CHEMINF, ...
-
-The BIBO ontology is used to represent documents.
-
-Throughout this paper, various prefixes are used to simplify RDF output, outlined
+During the life time of this project, the ontologies used in ChEMBL-RDF have changed,
+initially using a custom, ad-hoc ontology, but increasingly with community-proposed
+ontologies, making the RDF more interoperable. The ChEMBL-RDF 13 uses ontology
+standards like Bibliography Ontology for Citation Typinh Ontology for liteature
+references, and domain ontologies like the Protein Ontology and the Chemical Information
+Ontology. Throughout this paper, various prefixes are used to simplify RDF output, outlined
 in Table~\ref{namespaces}.
 
-\subsection{Use cases}
+To expose that ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
+hosted at Uppsala University, using the Open Source Virtuoso software. Use is free, but the
+querying is capped, based on the estimated computational effort. Second, resources have
+been made dereferencible using the Kasabi platform. This platform also provides a SPARQL
+end point, but requires a user account.
 
-A few new uses cases have been developed for this paper to demonstrate not just the intrinsic
-value of the data in the ChEMBL database as done previously, but also to show the possibilities
-the Linked Data nature offers.
-
-... describe what we want to show ...
+As a method to further test the access and possibilities of this Linked Data version of
+ChEMBL, a few new uses cases have been developed for this paper.
 
@@
 Common & Vocabularies \\
+bibo    & Bibliography Ontology~\cite{Giasson2011} \\
+        & http://purl.org/ontology/bibo/ \\
+cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
+        & http://semanticscience.org/resource/ \\
 cito    & Citation Typing Ontology~\cite{Shotton2010} \\
         & http://purl.org/spar/cito/ \\
-cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
-        & http://semanticscience.org/resource/ \\
-bibo    & Bibliography Ontology~\cite{Giasson2011} \\
-        & http://purl.org/ontology/bibo/ \\
+pro     & PRotein Ontology~\cite{XXX} \\
+        & http://cccccc.cc/xxxxxx/ \\
 
@@
 act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
+assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
 chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
+mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
 res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
-assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
-mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
 \hline

### 2012-03-06 article.tex
Added Janna to the team
add some information about the chembl-webapp
chembl-webapp has been revived as a project independent of the bio2rdf-webapp where it previously sat

it is available at https://github.com/ansell/chembl-webapp

@@
 \author[D]{\fnms{Antony} \snm{Williams}},
-\author[E]{\fnms{John} \snm{Overington}},
-\author[E]{\fnms{Anna} \snm{Gaulton}},
+\author[E]{\fnms{Janna} \snm{Hastings}},
+\author[F]{\fnms{John} \snm{Overington}},
+\author[F]{\fnms{Anna} \snm{Gaulton}},
 
@@
 \address[D]{xxx}
-\address[E]{European Bioinformatics Institute}
+\address[E]{ChEBI, European Bioinformatics Institute}
+\address[F]{ChEMBL, European Bioinformatics Institute}
 
@@
 
-The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2011.
+The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
 RDF triples were created from a local MySQL database to which the SQL data dump of ChEMBL was
@@
 
-Peter: write up the proxying... if that still works...
+%Peter: write up the proxying...
+if that still works...
+It was broken through disuse, but is available again in its rdf.farmbio.uu.se configuration, separately from the bio2rdf-webapp at https://github.com/ansell/chembl-webapp/
 
-The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011}, and access to SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
+The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
 A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases \cite{TheUniProtConsortium2010}.
-These links are vital to provide context for users who need to query across both Chembl and its related databases.
-The Linked Data server used by Bio2RDF has been configured for Chembl to provide URL based services for text and link searches.
+These links are vital to provide context for use cases that require a correlation between chemical structures and other scientific data.
+
+The Linked Data server used by Bio2RDF has been reconfigured for Chembl to provide URL based services for standard URI resolution, along with text and link searches \footnote{\url{https://github.com/ansell/chembl-webapp}}.
+It proxies the standard item URIs by translating URLs between those requested by users and the URIs that are available in the databases.
+For example, if the Chembl web application is running on the users local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}} will be resolved from the database using the full original URI \footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}.
+If the user requested an RDF document, using content negotiation, the original URIs will be unchanged, however, if the user requested an HTML document, the results will contain both the original RDF triples, represented using RDFa, with links that resolve using the users local machine.
+
+The links services enables the Chembl application to derive both forward links, originating in Chembl \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}}, and backward links, originating in other databases, such as LODD, Bio2RDF \footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF.
+These services are vital to efficiently navigate the Linked Data web, as it is both impractical and inefficient to require users to crawl the web before they can discover all of the relevant resources.
+These services are currently only supplied as web services from Chembl and Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data providers in the future.
+Datasets that are available in SPARQL endpoints can be queried for links efficiently using templated queries in the Chembl web application.
 
@@
 \subsubsection{Linking out to ChEMBL}
+Peter: write up the proxying... if that still works...
 
@@
         & http://purl.org/spar/cito/ \\
-pro     & PRotein Ontology~\cite{XXX} \\
-        & http://cccccc.cc/xxxxxx/ \\
+pro     & PRotein Ontology~\cite{Sidhu2006} \\
+        & http://purl.obolibrary.org/obo/ \\
 

### 2012-03-06 article.tex
Added Valery and the RSC affiliation details, emailed by Antony
Full names of the people in the Kasabi team
Added new lines (making easier-to-read git diffs) and s/Chembl/ChEBML/

@@
 \author[D]{\fnms{Antony} \snm{Williams}},
+\author[D]{\fnms{Valery} \snm{Tkachenko}},
 \author[E]{\fnms{Janna} \snm{Hastings}},
@@
 \address[C]{University of Queensland}
-\address[D]{xxx}
+\address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
 \address[E]{ChEBI, European Bioinformatics Institute}
@@
 
-The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data server \cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases \cite{Belleau2008}.
-A number of these databases are referenced in Chembl, including Chebi, Pubmed, and both the Uniprot protein and taxonomy databases \cite{TheUniProtConsortium2010}.
-These links are vital to provide context for use cases that require a correlation between chemical structures and other scientific data.
+The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data
+server~\cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases~\cite{Belleau2008}.
+A number of these databases are referenced in ChEMBL, including ChEBI, Pubmed, and both the
+Uniprot protein and taxonomy databases~\cite{TheUniProtConsortium2010}. These links are
+vital to provide context for use cases that require a correlation between chemical
+structures and other scientific data.
 
-The Linked Data server used by Bio2RDF has been reconfigured for Chembl to provide URL based services for standard URI resolution, along with text and link searches \footnote{\url{https://github.com/ansell/chembl-webapp}}.
-It proxies the standard item URIs by translating URLs between those requested by users and the URIs that are available in the databases.
-For example, if the Chembl web application is running on the users local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}} will be resolved from the database using the full original URI \footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}.
-If the user requested an RDF document, using content negotiation, the original URIs will be unchanged, however, if the user requested an HTML document, the results will contain both the original RDF triples, represented using RDFa, with links that resolve using the users local machine.
+The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
+based services for standard URI resolution, along with text and link searches
+\footnote{\url{https://github.com/ansell/chembl-webapp}}. It proxies the standard item
+URIs by translating URLs between those requested by users and the URIs that are available
+in the databases. For example, if the ChEMBL web application is running on the users
+local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the
+article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}}
+will be resolved from the database using the full original URI
+\footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}. If the
+user requested an RDF document, using content negotiation, the original URIs will be unchanged,
+however, if the user requested an HTML document, the results will contain both the
+original RDF triples, represented using RDFa, with links that resolve using the users local machine.
 
-The links services enables the Chembl application to derive both forward links, originating in Chembl \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}}, and backward links, originating in other databases, such as LODD, Bio2RDF \footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF.
-These services are vital to efficiently navigate the Linked Data web, as it is both impractical and inefficient to require users to crawl the web before they can discover all of the relevant resources.
-These services are currently only supplied as web services from Chembl and Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data providers in the future.
-Datasets that are available in SPARQL endpoints can be queried for links efficiently using templated queries in the Chembl web application.
+The links services enables the Chembl application to derive both forward links, originating in
+ChEMBL \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}},
+and backward links, originating in other databases, such as LODD, Bio2RDF
+\footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF.
+These services are vital to efficiently navigate the Linked Data web, as it is both
+impractical and inefficient to require users to crawl the web before they can discover all of the
+relevant resources. These services are currently only supplied as web services from ChEMBL and
+Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data
+providers in the future. Datasets that are available in SPARQL endpoints can be queried for
+links efficiently using templated queries in the ChEMBL web application.
 
@@
 
-Tim and Foo of Kasabi. What S.H. Name of the Bioinformatics department at Uppsala University.
+Tim Hodson and Zach Beauvais of Kasabi for their support. What S.H. Name of the Bioinformatics department at Uppsala University.
 

### 2012-03-07 article.tex
Added the main ChEMBL paper

@@
 ChEMBL contains biological activities for over a million chemical entities, extracted from
-literature, providing a unique resource for drug discovers~\cite{Warr2010}. Originally, this
+literature, providing a unique resource for drug discovers~\cite{Gaulton2012,Warr2010}. Originally, this
 data was available for download and via a web interface. The former requires setting up

### 2012-03-07 article.tex
Define \url
Added missing references

--- 2012-03-06 article.tex
@@
 \begin{document}
+
+\newcommand{\url}[1]{#1}
 
@@
 ChEMBL contains biological activities for over a million chemical entities, extracted from
-literature, providing a unique resource for drug discovers~\cite{Warr2010}. Originally, this
+literature, providing a unique resource for drug discovers~\cite{Gaulton2012,Warr2009}. Originally, this
 data was available for download and via a web interface. The former requires setting up
@@
 Two independent teams created RDF triples for the ChEMBL data, resulting in exposure
-via Chem2Bio2RDF~\cite{} and ChEMBL-RDF~\cite{Willighagen2011}.
+via Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}.
 

### 2012-03-07 article.tex
Andra will join too, making ChEMBL literature citation counting in CitedIn

@@
 \author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Corresponding author.\\ E-mail: egon.willighagen@maastrichtuniversity.nl.}}
+\runningauthor{E.L. Willighagen et al.}
 % Two or more authors (order to be determined):
+\author[A]{\fnms{Andra} \snm{Waagmeester}}
 \author[B]{\fnms{Ola} \snm{Spjuth}},
@@
 
-\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands\\
-E-mail: egon.willighagen@maastrichtuniversity.nl}
-\runningauthor{E.L. Willighagen et al.}
+\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands}
 \address[B]{Uppsala University}
@@
 
-%\subsubsection{CitedIn}
+\subsubsection{CitedIn}
 
-%Andra: use the SPARQL end point to find which entries in ChEMBL cite what papers
+Andra: use the SPARQL end point to find which entries in ChEMBL cite what papers
 
-%The link between data and formal publication is important in many areas of
-%attribution, scientist ranking, etc, as outline in \cite{Waagmeester2012}.
-%ChEMBL contains many literature references, and we wish to query this data
-%for CitedIn.
+The link between data and formal publication is important in many areas of
+attribution, scientist ranking, etc, as outline in \cite{Waagmeester2012}.
+ChEMBL contains many literature references, and we wish to query this data
+for CitedIn.
 
@@
 David: Chem2Bio2RDF
-
-%\subsection{Use Case 2: Chem2Bio2RDF}
-
-%David: Chem2Bio2RDF
 

### 2012-03-10 article.tex
Added David and Bin from Chem2Bio2RDF
Moved figures earlier into the paper, for better output
Added a screenshot of SNORQL, and wrapped the verbatim N3 examples

@@
 \author[F]{\fnms{Anna} \snm{Gaulton}},
+\author[F]{\fnms{Mark} \snm{Davies}},
+\author[G]{\fnms{Bin} \snm{Chen}},
+\author[G]{\fnms{David} \snm{Wild}}
 
@@
 \address[F]{ChEMBL, European Bioinformatics Institute}
+\address[G]{Indiana University}
 
@@
 
+\begin{table*}
+\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
+\begin{tabular}{ll}
+\hline
+Common & Vocabularies \\
+bibo    & Bibliography Ontology~\cite{Giasson2011} \\
+        & http://purl.org/ontology/bibo/ \\
+cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
+        & http://semanticscience.org/resource/ \\
+cito    & Citation Typing Ontology~\cite{Shotton2010} \\
+        & http://purl.org/spar/cito/ \\
+pro     & PRotein Ontology~\cite{Sidhu2006} \\
+        & http://purl.obolibrary.org/obo/ \\
+
+\hline
+ChEMBL-RDF & Namespaces\\
+act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
+assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
+chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
+mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
+res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
+\hline
+\end{tabular}
+\end{table*}
+
 To expose that ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
@@
 For each of the common resource classes a triple pattern was defined, following the
-data available in the relational database.
+data available in the relational database. ... see Figure~\ref{f1}.
+
+\begin{figure}[t]
+\includegraphics[width=0.5\textwidth]{figs/relations}
+\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
+of common classes, while others are instances.}\label{f1}
+\end{figure}
 
@@
  a cheminf:CHEMINF_000018 ;
-      cheminf:SIO_000300 "CC(C)C[C@@H]1N2C=CC(NC(=O)[C@H](Cc3ccccc3)N(C)C(=O)C(Cc4c[nH]c5ccccc45)NC(=O)C(CCC(=O)N)NC(=O)[C@@H](CCSCCSCC[C@H]6NC(=O)[C@H](CC(C)C)N7C=CC(NC(=O)[C@H](Cc8ccccc8)N(C)C(=O)C(Cc9c[nH]c%10ccccc9%10)NC(=O)C(CCC(=O)N)NC6=O)C7=O)NC1=O)C2=O" .
+ cheminf:SIO_000300
+  "CC(C)C[C@@H]1N2C= .... O)C7=O)NC1=O)C2=O" .
 
@@
  a cheminf:CHEMINF_000113 ;
-      cheminf:SIO_000300 "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" .
+ cheminf:SIO_000300
+  "InChI=1S/C82H102N .... 66+,67+,68+/m1/s1" .
 
@@
 mol:m41
-      rdfs:label "CHEMBL406142" , "Bis(3-[14-Benzyl-11-(1H-indol-3-ylmethyl)-2-isobutyl-13-methyl-5-(2-methylsulfanyl-ethyl)-3,6,9,12,15,20-hexaoxo-1,4,7,10,13,16-hexaaza-bicyclo[15.2.1]icos-18-en-8-yl]-propionamide)" ;
+ rdfs:label "CHEMBL406142" , "Bis(3-[1 .... yl]-propionamide)" ;
  rdfs:subClassOf cheminf:CHEMINF_000000 ;
@@
   m41:inchikey , m41:smiles , m41:inchi ;
-      <http://www.blueobelisk.org/chemistryblogs/inchi>
-              "InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1" ;
-      owl:equivalentClass <http://rdf.openmolecules.net/?InChI=1S/C82H102N16O14S2/c1-47(2)39-67-77(107)89-59(73(103)87-57(25-27-69(83)99)71(101)93-63(43-51-45-85-55-23-15-13-21-53(51)55)79(109)95(5)65(41-49-17-9-7-10-18-49)75(105)91-61-29-33-97(67)81(61)111)31-35-113-37-38-114-36-32-60-74(104)88-58(26-28-70(84)100)72(102)94-64(44-52-46-86-56-24-16-14-22-54(52)56)80(110)96(6)66(42-50-19-11-8-12-20-50)76(106)92-62-30-34-98(82(62)112)68(40-48(3)4)78(108)90-60/h7-24,29-30,33-34,45-48,57-68,85-86H,25-28,31-32,35-44H2,1-6H3,(H2,83,99)(H2,84,100)(H,87,103)(H,88,104)(H,89,107)(H,90,108)(H,91,105)(H,92,106)(H,93,101)(H,94,102)/t57?,58?,59-,60-,61?,62?,63?,64?,65+,66+,67+,68+/m1/s1> , <http://bio2rdf.org/chebi:100041> , chemblid:CHEMBL406142 .
+ owl:equivalentClass <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
+  <http://bio2rdf.org/chebi:100041> ,
+  chemblid:CHEMBL406142 .
 \end{verbatim}
@@
  dc:date "2004" ;
- dc:isPartOf journal:j6c706049c2e08871b7c46a6528065736 ;
+ dc:isPartOf
+  journal:j6c706049c2e08871b7c46a6528065736 ;
  bibo:issue "1" ;
@@
 Previously, a SPARQL end point has been set up at Uppsala University.
+
+For human-oriented browsing, a SNORQL front-end has been set up, which we
+recently updated to have a menu item for superclasses, that is classes
+of which subclasses are found in the data, as shown in Figure~\ref{f2}.
+
+\begin{figure}[bt]
+\includegraphics[width=0.5\textwidth]{snorql}
+\caption{The SNORQL interface for browsing the SPARQL end point.}\label{f2}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-\includegraphics[width=0.5\textwidth]{figs/relations}
-\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
-of common classes, while others are instances.}\label{f1}
-\end{figure}
-
-\begin{table*}
-\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
-\begin{tabular}{ll}
-\hline
-Common & Vocabularies \\
-bibo    & Bibliography Ontology~\cite{Giasson2011} \\
-        & http://purl.org/ontology/bibo/ \\
-cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
-        & http://semanticscience.org/resource/ \\
-cito    & Citation Typing Ontology~\cite{Shotton2010} \\
-        & http://purl.org/spar/cito/ \\
-pro     & PRotein Ontology~\cite{Sidhu2006} \\
-        & http://purl.obolibrary.org/obo/ \\
-
-\hline
-ChEMBL-RDF & Namespaces\\
-act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
-assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
-chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
-mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
-res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
-\hline
-\end{tabular}
-\end{table*}
-
 \bibliography{article}

### 2012-03-11 article.tex
Small edits to existing form of document, inserted comments about license. Will send seperate email regarding text for ChemSpider to insert into article.

@@
 This paper describes the continued conversion of ChEMBL data into RDF triples, using a few existing
-ontologis, and demonstrates a few new use cases, taking advantage of its linked nature.
+ontologies, and demonstrates a few new use cases, taking advantage of its linked nature.
 \end{abstract}
@@
 phenomena not yet understood is limited by the availability of data. Such data analysis is
-beyond the scope is single data sets; it requires data integration from many sources, for
+beyond the scope of single data sets; it requires data integration from many sources, for
 example, systems biology integrating micro-array differential expression to biological
@@
 
+AJW> CIte here the Open PHACTS paper submitted next week: Open PHACTS: Semantic interoperability for drug discovery
+Antony J.
+Williams, Lee Harland, Paul Groth, Stephen Pettifer, Christine Chichester, Egon L.
+Willighagen, Chris T.
+Evelo, Niklas Blomberg, Gerhard Ecker, Carole Goble and Barend Mons, Drug Disc.
+Today.
+Submitted for publication.
+Hopefully reviewed and with DOI available by the time this article is accepted.
+
 ChEMBL contains biological activities for over a million chemical entities, extracted from
-literature, providing a unique resource for drug discovers~\cite{Gaulton2012,Warr2009}. Originally, this
+literature, providing a unique resource for drug discovers~\cite{Gaulton2012,Warr2009}.
+It is updated on a fairly frequent basis as the existing data is further curated and new data is added.
+Originally, this
 data was available for download and via a web interface. The former requires setting up
@@
 RDF triples were created from a local MySQL database to which the SQL data dump of ChEMBL was
-inserted, using a PHP script, available from source code hosting service
+inserted, using a PHP script, available from the source code hosting service
 GitHub~\citep{ChEMBLRDFGitHub}, a process outlined earlier in a best practices note by
@@
 initially using a custom, ad-hoc ontology, but increasingly with community-proposed
-ontologies, making the RDF more interoperable. The ChEMBL-RDF 13 uses ontology
-standards like Bibliography Ontology for Citation Typinh Ontology for liteature
+ontologies, making the RDF more interoperable. The ChEMBL-RDF 13 dataset uses ontology
+standards such as the Bibliography Ontology for Citation Typing Ontology for literature
 references, and domain ontologies like the Protein Ontology and the Chemical Information
-Ontology. Throughout this paper, various prefixes are used to simplify RDF output, outlined
+Ontology. Throughout this paper, various prefixes are used to simplify the RDF output and they are outlined
 in Table~\ref{namespaces}.
@@
 querying is capped, based on the estimated computational effort. Second, resources have
-been made dereferencible using the Kasabi platform. This platform also provides a SPARQL
+been made dereferencable using the Kasabi platform. This platform also provides a SPARQL
 end point, but requires a user account.
@@
 Antony: if you can contribute ChemSpider IDs for the ChEMBL compounds ... 
+AJW> CC-BY-SA 3.0 licensing approved for the mappings between ChEMBL and ChemSpider. Please use the Open PHACTS data provided and attribution to ChEMBL, OPS and ChemSPider.
 

### 2012-03-11 article.tex
Added initial text about the ChemSpider link (by Antony)

@@
 
+RSC’s ChemSpider~\cite{Pence2010} is a free online database of over 26 million unique
+chemical compounds aggregated from over 400 data sources as well as chemical data extracted
+from RSC scientific articles and databases. Since its inception efforts have been made to
+utilize the platform as both a deposition platform for the community to contribute data as
+well as a platform for annotation and curation. Studies have shown that there are data
+quality issues in many of the public compound databases~\cite{Williams2011} and ChemSpider has become a
+valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
+is presently providing the chemical services underpinning the Open PHACTS semantic web
+project~\cite{Williams2012} providing access to structure, substructure and similarity searching services
+to the core architecture of the project. Specific chemical data sources containing data
+mappings between ChemSpider identifiers (CSIDs) and the original data source identifiers
+have been provided to the triple store, together with chemical identifiers including
+validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
+The data mappings between the ChemSpider IDs and the data source IDs are released to
+the community under Creative Commons licenses (CC-BY-SA 3.0). Link mappings of the
+form:
+
+\begin{tiny}
+\begin{verbatim}
+<http://rdf.chemspider.com/370>
+ skos:exactMatch
+ <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846>
+\end{verbatim}
+\end{tiny}
+
+have been constructed.
+
 \subsubsection{Linking out to ChEMBL}

### 2012-03-29 article.tex
Already cited in the first paragraph (label: OpenPHACTS); detail of citation will follow when available

@@
 data sets~\cite{Samwald2011,OpenPHACTS}.
-
-AJW> CIte here the Open PHACTS paper submitted next week: Open PHACTS: Semantic interoperability for drug discovery
-Antony J.
-Williams, Lee Harland, Paul Groth, Stephen Pettifer, Christine Chichester, Egon L.
-Willighagen, Chris T.
-Evelo, Niklas Blomberg, Gerhard Ecker, Carole Goble and Barend Mons, Drug Disc.
-Today.
-Submitted for publication.
-Hopefully reviewed and with DOI available by the time this article is accepted.
 

### 2012-03-29 article.tex
Added an initial LOD cloud, showing how ChEMBL-RDF embeds

@@
 literature, providing a unique resource for drug discovers~\cite{Gaulton2012,Warr2009}.
-It is updated on a fairly frequent basis as the existing data is further curated and new data is added.
-Originally, this
+It is updated on a fairly frequent basis as the existing data is further curated and new data is added. Originally, this
 data was available for download and via a web interface. The former requires setting up
@@
 
+\subsection{Linked Open Data}
+
+To make the data a proper citizen of the semantic web, we link out to various resources.
+The web resources linked to as well as the predicate used for that link is given in
+Figure~\ref{2}.
+Links out are made to ChemSpider and RDF OpenMolecules for compounds, Bio2RDF for
+protein targets, and CrossRef and Bio2RDF for DOIs and PubMed identifiers respectively.
+
+\begin{figure}[t]
+\includegraphics[width=0.5\textwidth]{figs/lodgraph}
+\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
+Edges are labeled by the predicates making the links.}\label{2}
+\end{figure}
+
 \subsection{SPARQL end points}

### 2012-03-29 article.tex
Some talk about recent content changes

@@
 FIXME!
+
+In contrast to earlier releases of ChEMBL-RDF, it now contains the chemical properties
+provided by the ChEMBL database as calculated by ACD/Labs and XXXXX. The data includes
+chemical properties like polar surface area, pKa and logP, counts for hydrogen bond donor
+and acceptor, and rotational bonds. This data is provided for XXXXX structures.
+
+The ChemSpider co-authors complemented the original ChEMBL data with ChemSpider identifiers
+for the chemical structures in the data set. Links were generated by ....
+
+Valery: FIXME
+
+Resulting in 1073967 ChemSpider identifiers.
 

### 2012-04-02 article.tex
Written a full abstract

@@
 \begin{abstract}
-This paper describes the continued conversion of ChEMBL data into RDF triples, using a few existing
-ontologies, and demonstrates a few new use cases, taking advantage of its linked nature.
+Making data available as RDF has the advantages that integration with other web resources brings:
+we can link out to related data, others can link to this data, it makes the data machine readable,
+is easy to extend with additional information,  and enables a representation of the data in common
+vocabularies. This combination provides a clear, open interface.
+This paper describes the continued conversion of data from the ChEMBL database into RDF triples.
+This updated version of ChEMBL-RDF now uses recently introduced, such as the CHEMINF and CiTO ontologies,
+exposes more information from the database, and is now available as dereferencable, linked data.
+To demonstrate the new features, we here present a few new use cases, showing integration with
+other web resources using semantic web technologies.
 \end{abstract}

### 2012-04-02 article.tex
Fixed column overflows

@@
 
-The full set of triples are available from http://semantics.bigcat.unimaas.nl/chembl12/.
+The full set of triples are available from the website http://semantics.bigcat.unimaas.nl/chembl12/.
 
@@
 \begin{figure}[t]
-\includegraphics[width=0.5\textwidth]{figs/relations}
+\includegraphics[width=0.45\textwidth]{figs/relations}
 \caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
@@
  chembl:hasDescription
-  "Inhibition of cytochrome P450 progesterone 15-alpha hydroxylase" ;
+  "Inhibition of ... hydroxylase" ;
  chembl:hasTarget target:t100122 .
@@
   "Maltase-glucoamylase, intestinal" ;
- dc:identifier "uniprot:O43451" , "3.2.1.3" ;
+ dc:identifier "uniprot:O43451" ,
+   "3.2.1.3" ;
  dc:title "Maltase-glucoamylase" ;
@@
  a cheminf:CHEMINF_000059 ;
- cheminf:SIO_000300 "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
+ cheminf:SIO_000300
+   "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .
 
@@
 mol:m41
- rdfs:label "CHEMBL406142" , "Bis(3-[1 .... yl]-propionamide)" ;
+ rdfs:label "CHEMBL406142" , 
+   "Bis(3-[1 .... yl]-propionamide)" ;
  rdfs:subClassOf cheminf:CHEMINF_000000 ;
@@
   m41:inchikey , m41:smiles , m41:inchi ;
- owl:equivalentClass <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
+ owl:equivalentClass 
+  <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
   <http://bio2rdf.org/chebi:100041> ,
@@
 \begin{figure}[t]
-\includegraphics[width=0.5\textwidth]{figs/lodgraph}
+\includegraphics[width=0.45\textwidth]{figs/lodgraph}
 \caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
@@
 \begin{figure}[bt]
-\includegraphics[width=0.5\textwidth]{snorql}
+\includegraphics[width=0.45\textwidth]{snorql}
 \caption{The SNORQL interface for browsing the SPARQL end point.}\label{f2}
@@
 based services for standard URI resolution, along with text and link searches
-\footnote{\url{https://github.com/ansell/chembl-webapp}}. It proxies the standard item
+\cite{WebAppGitHub}. It proxies the standard item
 URIs by translating URLs between those requested by users and the URIs that are available
 in the databases. For example, if the ChEMBL web application is running on the users
-local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the
-article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}}
-will be resolved from the database using the full original URI
-\footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}. If the
+local machine, e.g. \url{http://localhost:8080/chembl/}, then a request for the
+article with identifier ``a31863'', \url{http://localhost:8080/-chembl/article/a31863},
+will be resolved from the database using the full original URI,
+\url{http://data.kasabi.com/-dataset/chembl-rdf/13/activity/a31863}. If the
 user requested an RDF document, using content negotiation, the original URIs will be unchanged,
@@
 
-The links services enables the Chembl application to derive both forward links, originating in
-ChEMBL \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}},
-and backward links, originating in other databases, such as LODD, Bio2RDF
-\footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF.
+The links services enables the ChEMBL application to derive both forward links, originating in
+ChEMBL, e.g. \url{http://localhost:8080/chembl/linkstonamespace/-targetns/originalns:identifier},
+and backward links, originating in other databases, such as LODD, Bio2RDF, e.g.
+\url{http://bio2rdf.org/linksns/targetns/originalns:ident-ifier}, or Chem2Bio2RDF.
 These services are vital to efficiently navigate the Linked Data web, as it is both

### 2012-04-02 article.tex
Restructured things

@@
 querying is capped, based on the estimated computational effort. Second, resources have
-been made dereferencable using the Kasabi platform. This platform also provides a SPARQL
+been made dereferencable using the Kasabi platform~\cite{kasabi}. This platform also provides a SPARQL
 end point, but requires a user account.
@@
 
-\subsection{Data Statistics}
+We here present the updated ChEMBL-RDF.
+
+\subsection{ChEMBL-RDF}
+
+\subsubsection{Data Statistics}
 
@@
 
-The ChemSpider co-authors complemented the original ChEMBL data with ChemSpider identifiers
-for the chemical structures in the data set. Links were generated by ....
+\subsubsection{Data Statistics and Validation}
 
-Valery: FIXME
+John/Anna: common statistics with SPARQL
 
-Resulting in 1073967 ChemSpider identifiers.
+\subsubsection{Added Data}
 
-\subsection{Data Structure}
+RSC's ChemSpider~\cite{Pence2010} is a free online database of over 26 million unique
+chemical compounds aggregated from over 400 data sources as well as chemical data extracted
+from RSC scientific articles and databases. Since its inception efforts have been made to
+utilize the platform as both a deposition platform for the community to contribute data as
+well as a platform for annotation and curation. Studies have shown that there are data
+quality issues in many of the public compound databases~\cite{Williams2011} and ChemSpider has become a
+valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
+is presently providing the chemical services underpinning the Open PHACTS semantic web
+project~\cite{Williams2012} providing access to structure, substructure and similarity searching services
+to the core architecture of the project. Specific chemical data sources containing data
+mappings between ChemSpider identifiers (CSIDs) and the original data source identifiers
+have been provided to the triple store, together with chemical identifiers including
+validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
+The data mappings between the ChemSpider IDs and the data source IDs are released to
+the community under Creative Commons licenses (CC-BY-SA 3.0; attribution should be
+made to the original ChEMBL database, Open PHACTS, and ChemSpider).
+
+Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
+are also available via a CHEMINF representation:
+
+\begin{tiny}
+\begin{verbatim}
+<http://rdf.chemspider.com/370>
+ skos:exactMatch
+ <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
+\end{verbatim}
+\end{tiny}
+
+FIXME: add ChemSpider IDs with CHEMINF.
+
+\subsubsection{Data Structure}
 
@@
 
+\subsubsection{Linking out to Bio2RDF}
+
+Peter/Egon: write up links to Bio2RDF for protein, pubmed, ...
+
+\subsubsection{Linking out to ChemSpider}
+
+\subsubsection{Linking out to RDF OpenMolecules}
+
+\subsubsection{Linking out to CrossRef}
+
 \subsection{SPARQL end points}
@@
 
-\subsubsection{Bio2RDF}
+\section{Applications}
+
+\subsection{Bio2RDF}
 
@@
 
-\subsubsection{Data Statistics and Validation}
-
-John/Anna: common statistics with SPARQL
-
-\subsubsection{CitedIn}
+\subsection{CitedIn}
 
@@
 
-\subsubsection{Decision Support}
+\subsection{Decision Support}
 
@@
 
-\subsection{Linked Data on Kasabi}
+\subsection{Chem2Bio2RDF}
 
-Later, to satisfy the requirements for the Linked Open Drug Data \cite{Samwald2011}
-network, the data was also hosted on Kasabi, a triple store hosting service
-free for Open Data~\cite{kasabi}.
-
-David: Chem2Bio2RDF
-
-\subsubsection{Linking out to ChemSpider}
-
-Antony: if you can contribute ChemSpider IDs for the ChEMBL compounds ... 
-AJW> CC-BY-SA 3.0 licensing approved for the mappings between ChEMBL and ChemSpider. Please use the Open PHACTS data provided and attribution to ChEMBL, OPS and ChemSPider.
-
-RSC’s ChemSpider~\cite{Pence2010} is a free online database of over 26 million unique
-chemical compounds aggregated from over 400 data sources as well as chemical data extracted
-from RSC scientific articles and databases. Since its inception efforts have been made to
-utilize the platform as both a deposition platform for the community to contribute data as
-well as a platform for annotation and curation. Studies have shown that there are data
-quality issues in many of the public compound databases~\cite{Williams2011} and ChemSpider has become a
-valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
-is presently providing the chemical services underpinning the Open PHACTS semantic web
-project~\cite{Williams2012} providing access to structure, substructure and similarity searching services
-to the core architecture of the project. Specific chemical data sources containing data
-mappings between ChemSpider identifiers (CSIDs) and the original data source identifiers
-have been provided to the triple store, together with chemical identifiers including
-validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
-The data mappings between the ChemSpider IDs and the data source IDs are released to
-the community under Creative Commons licenses (CC-BY-SA 3.0). Link mappings of the
-form:
-
-\begin{tiny}
-\begin{verbatim}
-<http://rdf.chemspider.com/370>
- skos:exactMatch
- <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846>
-\end{verbatim}
-\end{tiny}
-
-have been constructed.
-
-\subsubsection{Linking out to ChEMBL}
-Peter: write up the proxying... if that still works...
-
-John/Anna: I can easily add links to webpages using foaf:homepage... are webpage URLs for entities stable enough?
-
-\subsubsection{Linking out to Bio2RDF}
-
-Peter/Egon: write up links to Bio2RDF for protein, pubmed, ...
+David/Bin: Chem2Bio2RDF
 
@@
 Tim Hodson and Zach Beauvais of Kasabi for their support. What S.H. Name of the Bioinformatics department at Uppsala University.
+
+FIXME: add IMI
 

### 2012-04-03 article.tex
Completing the text of the triple structures

@@
 \hline
-ChEMBL-RDF & Namespaces\\
+ChEMBL-RDF & Namespace\\
+\hline
+chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
+\hline
+ChEMBL-RDF & Prefixes\\
+\hline
 act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
 assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
-chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
 mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
@@
 end point, but requires a user account.
+
+An initial custom vocabulary has been defined, reflecting the concepts and links found in the
+data structure of the ChEMBL database. This vocabulary has only partially been converted into
+an OWL ontology, and is being phased out and replaced by common ontologies. This has not happened
+for all data yet, and terms from the ChEMBL-RDF namespace (\textit{http://rdf.farmbio.uu.se/chembl/onto/\#},
+prefix: \textit{chembl}) are still in use.
 
@@
 
-Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
-are also available via a CHEMINF representation:
-
-\begin{tiny}
-\begin{verbatim}
-<http://rdf.chemspider.com/370>
- skos:exactMatch
- <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
-\end{verbatim}
-\end{tiny}
-
-FIXME: add ChemSpider IDs with CHEMINF.
-
 \subsubsection{Data Structure}
 
-For each of the common resource classes a triple pattern was defined, following the
-data available in the relational database. ... see Figure~\ref{f1}.
+For each of the common resource classes, a triple pattern was defined, following the
+data available in the relational database. Figure~\ref{f1} shows how the various resource
+classes are linked together. This section will show how data in the ChEMBL database
+is exposed at an triple level.
 
@@
 
-Assays .... anything
+The core concept in the ChEMBL database is that of the biological activity. This
+is the type of information that the ChEMBL database extracts from literature.
+Activities are at this moment still using an informal, custom vocabulary: the
+type, fields and links to other resources are using the \textit{chembl}
+namespace. While ChEMBL provides both the original activities as found in literature
+and standardized values allowing comparison between studies, the triples only
+make the latter available. The CiTO ontology is used to point to the paper from
+which the data was extracted.
 
-\begin{small}
+\begin{footnotesize}
+\begin{verbatim}
+act:a31863
+ a       chembl:Activity ;
+ cito:citesAsDataSource
+  res:r6424 ;
+ chembl:forMolecule mol:m180094 ;
+ chembl:onAssay assay:a54505 ;
+ chembl:relation ">" ;
+ chembl:standardUnits
+  "nM" ;
+ chembl:standardValue
+  "100000"^^xsd:float ;
+ chembl:type "IC50" .
+\end{verbatim}
+\end{footnotesize}
+
+The activities themselves are measured against assays, which make up a second important
+resource type. Various assay types are found in the database: chembl:ADMET, chembl:Binding,
+chembl:Functional, chembl:Property, and chembl:Unassigned. The assay information link
+activities to a target (see below) and has a short description. The confidence information
+assigned by the ChEMBL is exposed too, and was previously used to improve activity
+prediction using Bayesian statistics~\cite{Willighagen2011}.
+
+\begin{footnotesize}
 \begin{verbatim}
@@
  a chembl:Assay ;
- chembl:extractedFrom
+ cito:citesAsDataSource
   res:r11347 ;
@@
 \end{verbatim}
-\end{small}
+\end{footnotesize}
 
-Targets ... protein, other
+The assays measure activities against particular biological targets. The ChEMBL database
+recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
+chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
+chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
+explicit types.
 
-\begin{small}
+\begin{footnotesize}
 \begin{verbatim}
@@
   <http://bio2rdf.org/taxonomy:9606> ;
- chembl:organism "Homo sapiens" ;
- = <http://bio2rdf.org/uniprot:O43451> , 
-   <http://bio2rdf.org/ec:3.2.1.3> .
+ chembl:organism "Homo sapiens" .
 \end{verbatim}
-\end{small}
-
-Activities ...
-
-\begin{small}
-\begin{verbatim}
-act:a31863
- a       chembl:Activity ;
- cito:citesAsDataSource
-  res:r6424 ;
- chembl:forMolecule mol:m180094 ;
- chembl:onAssay assay:a54505 ;
- chembl:relation ">" ;
- chembl:standardUnits
-  "nM" ;
- chembl:standardValue
-  "100000"^^xsd:float ;
- chembl:type "IC50" .
-\end{verbatim}
-\end{small}
+\end{footnotesize}
 
@@
 
-\begin{tiny}
+\begin{footnotesize}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{footnotesize}
+
+The molecular properties available from ChEMBL are now too exposed in triple format,
+and like the InChI and InChIKey, these are provided using the CHEMINF ontology structure:
+
+\begin{footnotesize}
+\begin{verbatim}
+mol:m1
+ cheminf:CHEMINF_000200 m1:alogp .
+
+m1:alogp
+ a cheminf:CHEMINF_000305 ;
+ cheminf:SIO_000300
+  "3.344"^^xsd:double .
+\end{verbatim}
+\end{footnotesize}
+
+Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
+are also available via a CHEMINF representation:
+
+\begin{tiny}
+\begin{verbatim}
+<http://rdf.chemspider.com/370>
+ skos:exactMatch
+ <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
+\end{verbatim}
 \end{tiny}
+
+FIXME: add ChemSpider IDs with CHEMINF.
 
@@
 \end{verbatim}
-\end{tiny}
+\end{footnotesize}
 

### 2012-04-04 article.tex
Added draft text and references on Decision Support.

@@
 \address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands}
-\address[B]{Uppsala University}
+\address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
 \address[C]{University of Queensland}
@@
 
+Bioclipse Decision Support (Bioclipse DS)~\cite{Spjuth:2011uq} is a user-oriented tool for providing on-time and on-demand information on chemical structures, based on the Bioclipse workbench.
+Such information can include calculated properties, data from database queries, and results from predictive models.
+Bioclipse DS has previously been demonstrated on predictive modeling in drug safety assessment~\cite{Spjuth:2011uq} and also been linked to invoke and present results from distributed toxicity predictions from the OpenTox infrastructure~\cite{Willighagen:2011kx}.
+
+In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and Chemspider, which enables users to for chemical structures in Bioclipse look up near neighbors on Chemspider via the Chemspider Web API (SOAP) and query ChEMBL-RDF for properties regarding any measured interactions.
+The results are presented alongside predictive models in Bioclipse, and can be used as decision support when evaluating chemical structures and consider strategies for e.g.
+optimization.
+
+[TODO: SCREENSHOT from Bioclipse-DS-ChEMBL-RDF]
+
 \subsection{Chem2Bio2RDF}

### 2012-04-05 article.tex
add address for my university
Completed a draft introduction into the graph structure

@@
 \address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
-\address[C]{University of Queensland}
+\address[C]{University of Queensland, St Lucia, Qld 4072, Australia}
 \address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
@@
 
+More than five thousand different activity types are captured by the ChEMBL database.
+The top five types are Potency (43\%), IC50 (13\%), MIC (4.6\%), Inhibition (3.7\%),
+and Ki (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
+using, an ontology.
+
 The activities themselves are measured against assays, which make up a second important
@@
 
-Compounds ... Three types of compounds, peptides, small chemicals, ...
+For drug discovery, the drugs themselves are the main topic of study.
+ChEMBL contains many different drug types, mostly small molecules,
+but also peptides, proteins, oligosaccharides, oligonucleotides, and
+even antibodies. All compounds are represented as classes, following
+the design of the CHEMINF ontology. The compound typing is expressed
+by subclassing other classes. For example, proteins subclass the
+Protein concept from the PRotein Ontology (PR\_000000001), small
+molecules subclass the Chemical Entity concept from the CHEMINF
+ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
+subclass their respective matches in the CHEBI ontology (CHEBI\_50699
+and CHEBI\_7754). For each drug compound the name and synonyms are
+provided as labels. When InChI and InChIKeys are available for a drug, then these are
+provided via the CHEMINF formalism (here abbreviated):
 
@@
 \begin{verbatim}
+mol:m41
+ rdfs:label "CHEMBL406142" , 
+   "Bis(3-[1 .... yl]-propionamide)" ;
+ rdfs:subClassOf cheminf:CHEMINF_000000 ;
+ cheminf:CHEMINF_000200
+  m41:inchikey , m41:smiles , m41:inchi ;
+
 m41:inchikey
@@
  owl:equivalentClass mol:m41 .
-
-mol:m41
- rdfs:label "CHEMBL406142" , 
-   "Bis(3-[1 .... yl]-propionamide)" ;
- rdfs:subClassOf cheminf:CHEMINF_000000 ;
- cheminf:CHEMINF_000200
-  m41:inchikey , m41:smiles , m41:inchi ;
- owl:equivalentClass 
-  <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
-  <http://bio2rdf.org/chebi:100041> ,
-  chemblid:CHEMBL406142 .
 \end{verbatim}
@@
 
-The molecular properties available from ChEMBL are now too exposed in triple format,
-and like the InChI and InChIKey, these are provided using the CHEMINF ontology structure:
+For small molecules, molecular properties are often available from ChEMBL, and as of ChEMBL-RDF 13
+these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
+CHEMINF ontology approach. Here is, for example, the ALogP value:
 
@@
 
-Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
-are also available via a CHEMINF representation:
-
-\begin{tiny}
-\begin{verbatim}
-<http://rdf.chemspider.com/370>
- skos:exactMatch
- <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
-\end{verbatim}
-\end{tiny}
-
-FIXME: add ChemSpider IDs with CHEMINF.
-
 For documents little information is replicated from the database, taking advantage
@@
 latter can be resolved online, providing curated information on the identified
-documents. For PubMed articles ...
+documents. For each paper, basic properties are provided using the BIBO
+ontology:
 
@@
  a bibo:Article ;
- rdfs:seeAlso <http://bio2rdf.org/pubmed:14695813> ;
  dc:date "2004" ;
@@
 \subsubsection{Linking out to ChemSpider}
+
+Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
+are also available via a CHEMINF representation:
+
+\begin{tiny}
+\begin{verbatim}
+<http://rdf.chemspider.com/370>
+ skos:exactMatch
+ <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
+\end{verbatim}
+\end{tiny}
 

### 2012-04-11 article.tex
Drafted the LOD section

@@
   "Membrane" , "Sulfation" ;
- chembl:hasTaxonomy
-  <http://bio2rdf.org/taxonomy:9606> ;
  chembl:organism "Homo sapiens" .
@@
 
-To make the data a proper citizen of the semantic web, we link out to various resources.
-The web resources linked to as well as the predicate used for that link is given in
-Figure~\ref{2}.
-Links out are made to ChemSpider and RDF OpenMolecules for compounds, Bio2RDF for
-protein targets, and CrossRef and Bio2RDF for DOIs and PubMed identifiers respectively.
+To make the data a proper citizen of the semantic web, we link out to various resources,
+which is visualized in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
+using the new complementary index and RDF OpenMolecules using InChI values, to Bio2RDF
+for protein targets using the uniprot identifier, and CrossRef and Bio2RDF for literature
+references using the DOIs and PubMed identifiers, respectively.
 
@@
 
-Peter/Egon: write up links to Bio2RDF for protein, pubmed, ...
+The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data
+server~\cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases~\cite{Belleau2008}.
+A number of these databases are referenced in ChEMBL, including ChEBI, Pubmed, and both the
+Uniprot protein and taxonomy databases~\cite{TheUniProtConsortium2010}. These links are
+vital to provide context for use cases that require a correlation between chemical
+structures and other scientific data.
+
+For protein links out are based on the FOOBAR for the species, the EC code for proteins,
+as well as UniProt identifiers.
+
+FIXME: should we link out to the new UniProt RDF?
+
+\begin{tiny}
+\begin{verbatim}
+target:t101191
+ chembl:hasTaxonomy <http://bio2rdf.org/taxonomy:9606> ;
+ owl:sameAs <http://bio2rdf.org/ec:2.7.11.1> ;
+ owl:sameAs <http://bio2rdf.org/uniprot:Q8IYT8> .
+\end{verbatim}
+\end{tiny}
+
+For papers with PubMed identifiers we also link out to Bio2RDF:
+
+\begin{tiny}
+\begin{verbatim}
+res:r23 rdfs:seeAlso <http://bio2rdf.org/pubmed:15149661> .
+\end{verbatim}
+\end{tiny}
 
@@
 \begin{verbatim}
-<http://rdf.chemspider.com/370>
- skos:exactMatch
- <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
+<http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846>
+ skos:exactMatch <http://rdf.chemspider.com/370> .
 \end{verbatim}
@@
 
+The InChI is a unique identifier for (small) organic molecules, and has been previously used
+to define unique IRIs for molecules~\cite{???}. While IRIs are theoretically unlimited in length,
+in practice web browsers and servers do limit the length of IRIs. Virtuoso is, unfortunately,
+a system which supports only IRIs of up to a certain length. Therefore, only for smaller molecules
+InChI-based links are created. Almost 1.3 million links are created, looking like:
+
+\begin{tiny}
+\begin{verbatim}
+mol:m62687 owl:equivalentClass
+ <http://rdf.openmolecules.net/?InChI=1S/CH4/h1H4> .
+\end{verbatim}
+\end{tiny}
+
+Notice here the use of owl:equivalentClass to match the formalism in CHEMINF that defines molecules
+are classes, rather than instances.
+
 \subsubsection{Linking out to CrossRef}
+
+Besides the PubMed identifiers used to link from literature references to Bio2RDF (see earlier),
+ChEMBL uses DOIs alternatively, which we use to link out to the RDF provided by CrossRef~\citep{CrossRefRDF}:
+
+\begin{tiny}
+\begin{verbatim}
+res:r2032 owl:sameAs <http://dx.doi.org/10.1016/0960-894X(96)00111-4> .
+\end{verbatim}
+\end{tiny}
 
@@
 It was broken through disuse, but is available again in its rdf.farmbio.uu.se configuration, separately from the bio2rdf-webapp at https://github.com/ansell/chembl-webapp/
-
-The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data
-server~\cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases~\cite{Belleau2008}.
-A number of these databases are referenced in ChEMBL, including ChEBI, Pubmed, and both the
-Uniprot protein and taxonomy databases~\cite{TheUniProtConsortium2010}. These links are
-vital to provide context for use cases that require a correlation between chemical
-structures and other scientific data.
 

### 2012-04-11 article.tex
Removed the separate section on SPARQL end points

@@
 
-\subsection{SPARQL end points}
-
-Previously, a SPARQL end point has been set up at Uppsala University.
-
-For human-oriented browsing, a SNORQL front-end has been set up, which we
-recently updated to have a menu item for superclasses, that is classes
-of which subclasses are found in the data, as shown in Figure~\ref{f2}.
-
-\begin{figure}[bt]
-\includegraphics[width=0.45\textwidth]{snorql}
-\caption{The SNORQL interface for browsing the SPARQL end point.}\label{f2}
-\end{figure}
-
 \section{Applications}

### 2012-04-11 article.tex
Added an intro for the application section

@@
 
+The ChEMBL-RDF is increasingly used, and we here present a few applications. The first application
+describes how it uses the SPARQL end point can be used to make the data available as linked data
+via the Bio2RDF platform. The second application takes advantage of the bibliographic information
+exposed as machine readable data, in calculating citation statistics. The third application shows
+an integration of ChEMBL-RDF with ChemSpider to provide an extension for the decision support
+platform in Bioclipse. The last application shows how Chem2Bio2RDF can combine the ChEMBL data
+with other life sciences databases, showing the power of the Linked Data approach.
+
 \subsection{Bio2RDF}

### 2012-04-15 article.tex
change rdfs:seeAlso for pubmed to skos:exactMatch
Checked state, and updated internal notes
More complete addresses

@@
 
-\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands}
+\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands}
 \address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
@@
 \address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
-\address[E]{ChEBI, European Bioinformatics Institute}
-\address[F]{ChEMBL, European Bioinformatics Institute}
-\address[G]{Indiana University}
+\address[E]{Chemoinformatics and Metabolism, European Bioinformatics Institute, POSTAL CODE, Hinxton, United Kingdom}
+\address[F]{EMBL-European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom}
+\address[G]{School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.}
 
@@
 \begin{verbatim}
-res:r23 rdfs:seeAlso <http://bio2rdf.org/pubmed:15149661> .
+res:r23 skos:exactMatch <http://bio2rdf.org/pubmed:15149661> .
 \end{verbatim}
@@
 \subsection{Bio2RDF}
-
-%Peter: write up the proxying...
-if that still works...
-It was broken through disuse, but is available again in its rdf.farmbio.uu.se configuration, separately from the bio2rdf-webapp at https://github.com/ansell/chembl-webapp/
 

### 2012-04-15 article.tex
Consistent referal to OpenMolecules RDF

@@
 which is visualized in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
-using the new complementary index and RDF OpenMolecules using InChI values, to Bio2RDF
+using the new complementary index and OpenMolecules RDF using InChI values, to Bio2RDF
 for protein targets using the uniprot identifier, and CrossRef and Bio2RDF for literature
@@
 
-\subsubsection{Linking out to RDF OpenMolecules}
+\subsubsection{Linking out to OpenMolecules RDF}
 
 The InChI is a unique identifier for (small) organic molecules, and has been previously used
-to define unique IRIs for molecules~\cite{???}. While IRIs are theoretically unlimited in length,
+to define unique IRIs for molecules~\cite{Bradley2009,Willighagen2011}. While IRIs are theoretically unlimited in length,
 in practice web browsers and servers do limit the length of IRIs. Virtuoso is, unfortunately,

### 2012-04-15 article.tex
Easier to read table

@@
 \hline
-Common & Vocabularies \\
+\multicolumn{2}{l}{\textbf{Common Vocabularies}} \\
 bibo    & Bibliography Ontology~\cite{Giasson2011} \\
@@
 
-\hline
-ChEMBL-RDF & Namespace\\
-\hline
+\multicolumn{2}{l}{\textbf{ChEMBL-RDF Namespace}} \\
 chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
-\hline
-ChEMBL-RDF & Prefixes\\
-\hline
+
+\multicolumn{2}{l}{\textbf{ChEMBL-RDF Prefixes }}\\
 act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\

### 2012-04-15 article.tex
Writing of the discussion section

@@
 Besides the PubMed identifiers used to link from literature references to Bio2RDF (see earlier),
-ChEMBL uses DOIs alternatively, which we use to link out to the RDF provided by CrossRef~\citep{CrossRefRDF}:
+ChEMBL uses DOIs alternatively, which we use to link out to the RDF provided by CrossRef~\citep{Bilder2011}:
 
@@
 
-Egon,Antony/Valery,Ola: develop and write up the Bioclipse Decision Support use case
+% Egon,Antony/Valery,Ola: develop and write up the Bioclipse Decision Support use case
 
@@
 
-outlook ...
+While we show here that the triple version of the ChEMBL data is very useful, the current triples are by no means static: ChEMBL-RDF
+will keep evolve, following new developments in the Linked Open Data world. For example, it is likely that over time we will link out
+to more Linked Data resources. More importantly, we wish to adopt more ontologies, of which the BioAssay Ontology is planned
+next~\cite{Visser2011}.
 
@@
 Tim Hodson and Zach Beauvais of Kasabi for their support. What S.H. Name of the Bioinformatics department at Uppsala University.
-
-FIXME: add IMI
 

### 2012-04-15 article.tex
OK, no clue what further discussion to write up

@@
 will keep evolve, following new developments in the Linked Open Data world. For example, it is likely that over time we will link out
-to more Linked Data resources. More importantly, we wish to adopt more ontologies, of which the BioAssay Ontology is planned
-next~\cite{Visser2011}.
+to more Linked Data resources. More importantly, we wish to adopt more common ontologies, of which the BioAssay Ontology is planned
+to be the first next to be adopted~\cite{Visser2011}. This should make the ChEMBL-RDF triples even more interoperable.
 
@@
 
-Tim Hodson and Zach Beauvais of Kasabi for their support. What S.H. Name of the Bioinformatics department at Uppsala University.
+Tim Hodson and Zach Beauvais of Kasabi and A. L\"ovgren at the at Uppsala Biomedical Centre at Uppsala University for their
+support.
 

### 2012-04-15 article.tex
Added text on the :has_role :drug

@@
         & http://purl.org/ontology/bibo/ \\
+chebi   & Chemical Entities of Biological Interest~\cite{DeMatos2010} \\
+        & http://purl.org/obo/owl/CHEBI\# \\
 cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
@@
         & http://purl.org/spar/cito/ \\
-pro     & PRotein Ontology~\cite{Sidhu2006} \\
+obo / pro & OBO \& PRotein Ontology~\cite{Sidhu2006} \\
           & http://purl.obolibrary.org/obo/ \\
@@
 ChEMBL contains many different drug types, mostly small molecules,
-but also peptides, proteins, oligosaccharides, oligonucleotides, and
-even antibodies. All compounds are represented as classes, following
-the design of the CHEMINF ontology. The compound typing is expressed
+but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
+even cells. All compounds are represented as classes, following
+the design of the CHEMINF ontology. These entities do not have a common
+superclass in common, but can easily be identified as having the 
+the role of being a drug. This is triplified using the OBO and ChEBI
+ontologies:
+
+\begin{footnotesize}
+\begin{verbatim}
+mol:m4 obo:has_role chebi:CHEBI_23888 .
+\end{verbatim}
+\end{footnotesize}
+
+The entity typing is expressed
 by subclassing other classes. For example, proteins subclass the

### 2012-04-15 article.tex
edit the abstract, introduction and methods sections

@@
 \begin{abstract}
-Making data available as RDF has the advantages that integration with other web resources brings:
-we can link out to related data, others can link to this data, it makes the data machine readable,
-is easy to extend with additional information,  and enables a representation of the data in common
-vocabularies. This combination provides a clear, open interface.
+Making data available as Linked Data using RDF is advantageous because it promotes integration with other web resources.
+RDF documents can natively link to related data, and others can link back using URIs. RDF makes the data machine-readable,
+using extensible vocabularies for additional information. This combination provides a new, open, interface for scientists to use ChEMBL.
 This paper describes the continued conversion of data from the ChEMBL database into RDF triples.
-This updated version of ChEMBL-RDF now uses recently introduced, such as the CHEMINF and CiTO ontologies,
-exposes more information from the database, and is now available as dereferencable, linked data.
-To demonstrate the new features, we here present a few new use cases, showing integration with
+This updated version of ChEMBL-RDF now uses recently introduced ontologies, including CHEMINF and CiTO,
+exposes more information from the database, and is now available as dereferencable, Linked Data.
+To demonstrate the new features, we present new use cases showing the benefits of integration with
 other web resources using semantic web technologies.
@@
 
-Despite the data deluge, finding new, unique, but significant patterns explaining biological
-phenomena not yet understood is limited by the availability of data. Such data analysis is
-beyond the scope of single data sets; it requires data integration from many sources, for
-example, systems biology integrating micro-array differential expression to biological
-pathways~\cite{}. Another prominent example is drug discovery where a new unique chemical entity is
-searched with just the right biological properties. This too requires linking many
-data sets~\cite{Samwald2011,OpenPHACTS}.
+The current scientific data deluge, in which datasets are growing faster than scientists can analyse and 
+curate them, provides several new challenges for information systems. Scientists wish to 
+discover new, unique, and significant patterns in datasets, which explaining biological 
+phenomena not yet understood. This process requires integration of datasets that are both 
+sparse, and growing independent of one another. The discovery of new patterns showing the 
+causes and effects of various biological phenomena is beyond the scope of single datasets.
+For example, systems biologists integrate micro-array differential expression datasets to 
+biological pathways, using various other datasets to provide evidence for the links~\cite{}. 
+Another prominent example is drug discovery where a new unique chemical entity is discovered 
+based on descriptions of its biological properties. This process requires the effective 
+linkage of many scientific datasets~\cite{Samwald2011,OpenPHACTS}.
 
-ChEMBL contains biological activities for over a million chemical entities, extracted from
-literature, providing a unique resource for drug discovers~\cite{Gaulton2012,Warr2009}.
-It is updated on a fairly frequent basis as the existing data is further curated and new data is added. Originally, this
-data was available for download and via a web interface. The former requires setting up
-the data in a relational database, while the latter limits the machine access to the data.
-Two independent teams created RDF triples for the ChEMBL data, resulting in exposure
-via Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}.
+ChEMBL contains descriptions for biological activities involving over a million chemical 
+entities, extracted from literature. This provides a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
+It is updated on a fairly frequent basis as the existing data is further curated and new data is added. 
+The ChEMBL dataset is available for download and can be browsed using a web interface. 
+The former requires scientists to import the data into a relational database, while the 
+latter limits the machine access to the data. Two independent teams have previously mapped 
+the ChEMBL dataset to RDF, Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}.
 
-This paper presents the history of the ChEMBL-RDF data set, details of the latest structures
-and ontologies used for the RDF version of ChEMBL 13, and a few new example use cases,
-exemplifying how the data set can be linked to other data sources to further support
-research in the life sciences.
+This paper presents the current state of the ChEMBL-RDF dataset, details of the latest structures
+and ontologies used to map version 13 of ChEMBL to RDF, along with new example use cases, showing 
+links to other datasets using their RDF URIs to further support research in the life sciences.
 
@@
 The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
-RDF triples were created from a local MySQL database to which the SQL data dump of ChEMBL was
-inserted, using a PHP script, available from the source code hosting service
-GitHub~\citep{ChEMBLRDFGitHub}, a process outlined earlier in a best practices note by
-the international Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.
+The SQL datadump provided by EMBL was inserted into a local MySQL database. A set of SQL queries 
+were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
+and are available from the source code hosting service, GitHub~\citep{ChEMBLRDFGitHub}. This process 
+has been outlined in a best practices note by the W3C Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.
 
-During the life time of this project, the ontologies used in ChEMBL-RDF have changed,
-initially using a custom, ad-hoc ontology, but increasingly with community-proposed
+In the previous RDF conversion the RDF triples in ChEMBL-RDF were represented using properties and c
+lasses from a custom, ad-hoc ontology~\cite{Willighagen2011}. The current version used community-proposed
 ontologies, making the RDF more interoperable. The ChEMBL-RDF 13 dataset uses ontology
-standards such as the Bibliography Ontology for Citation Typing Ontology for literature
-references, and domain ontologies like the Protein Ontology and the Chemical Information
-Ontology. Throughout this paper, various prefixes are used to simplify the RDF output and they are outlined
-in Table~\ref{namespaces}.
+standards such as the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology~\cite{Hastings2011} for literature
+references, and domain ontologies like the Protein Ontology\cite{Sidhu2006} and the Chemical Information
+Ontology\cite{Hastings2011}. Throughout this paper, various shorthand prefixes are used to denote different ontology 
+namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.
 
@@
 querying is capped, based on the estimated computational effort. Second, resources have
-been made dereferencable using the Kasabi platform~\cite{kasabi}. This platform also provides a SPARQL
-end point, but requires a user account.
+been made dereferencable using the Kasabi platform~\cite{kasabi}. The Kasabi hosting provides a SPARQL
+end point which requires a user account. A user account can be free for casual usage, or paid for consistent usage.
 
-An initial custom vocabulary has been defined, reflecting the concepts and links found in the
-data structure of the ChEMBL database. This vocabulary has only partially been converted into
-an OWL ontology, and is being phased out and replaced by common ontologies. This has not happened
-for all data yet, and terms from the ChEMBL-RDF namespace (\textit{http://rdf.farmbio.uu.se/chembl/onto/\#},
-prefix: \textit{chembl}) are still in use.
+Although many parts of the CHEMBL-RDF dataset use community based ontologies, some 
+terms from the previous ontology are still used. These terms were created under the 
+URI namespace \textit{http://rdf.farmbio.uu.se/chembl/onto/\#}, and are referenced here using the
+prefix: \textit{chembl}.
 
-As a method to further test the access and possibilities of this Linked Data version of
-ChEMBL, a few new uses cases have been developed for this paper.
+%As a method to further test the access and possibilities of this Linked Data version of
+%ChEMBL,  new uses cases have been developed for this paper.
 

### 2012-04-15 article.tex
rearrange Results section and edit some parts

@@
 
-Although many parts of the CHEMBL-RDF dataset use community based ontologies, some 
-terms from the previous ontology are still used. These terms were created under the 
+Although many parts of the CHEMBL-RDF 13 dataset now use community based ontologies, some 
+terms from the previous ad-hoc ontology are still used. These terms were created under the 
 URI namespace \textit{http://rdf.farmbio.uu.se/chembl/onto/\#}, and are referenced here using the
@@
 
-We here present the updated ChEMBL-RDF.
+We here present the updated ChEMBL-RDF, including a description of the mapping process 
+and the identification of links to other datasets.
 
 \subsection{ChEMBL-RDF}
-
-\subsubsection{Data Statistics}
-
-\# triples
-\# links out
-\# problems
-
-The full set of triples are available from the website http://semantics.bigcat.unimaas.nl/chembl12/.
-
-FIXME!
-
-In contrast to earlier releases of ChEMBL-RDF, it now contains the chemical properties
-provided by the ChEMBL database as calculated by ACD/Labs and XXXXX. The data includes
-chemical properties like polar surface area, pKa and logP, counts for hydrogen bond donor
-and acceptor, and rotational bonds. This data is provided for XXXXX structures.
-
-\subsubsection{Data Statistics and Validation}
-
-John/Anna: common statistics with SPARQL
-
-\subsubsection{Added Data}
-
-RSC's ChemSpider~\cite{Pence2010} is a free online database of over 26 million unique
-chemical compounds aggregated from over 400 data sources as well as chemical data extracted
-from RSC scientific articles and databases. Since its inception efforts have been made to
-utilize the platform as both a deposition platform for the community to contribute data as
-well as a platform for annotation and curation. Studies have shown that there are data
-quality issues in many of the public compound databases~\cite{Williams2011} and ChemSpider has become a
-valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
-is presently providing the chemical services underpinning the Open PHACTS semantic web
-project~\cite{Williams2012} providing access to structure, substructure and similarity searching services
-to the core architecture of the project. Specific chemical data sources containing data
-mappings between ChemSpider identifiers (CSIDs) and the original data source identifiers
-have been provided to the triple store, together with chemical identifiers including
-validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
-The data mappings between the ChemSpider IDs and the data source IDs are released to
-the community under Creative Commons licenses (CC-BY-SA 3.0; attribution should be
-made to the original ChEMBL database, Open PHACTS, and ChemSpider).
 
@@
 is the type of information that the ChEMBL database extracts from literature.
-Activities are at this moment still using an informal, custom vocabulary: the
+In CHEMBL-RDF activities are mostly represented using the custom CHEMBL-RDF vocabulary, where the
 type, fields and links to other resources are using the \textit{chembl}
@@
 
+\subsubsection{Data Statistics}
+
+\# triples
+\# links out
+\# problems
+
+The full set of triples are available from the website http://semantics.bigcat.unimaas.nl/chembl12/.
+
+FIXME!
+
+In contrast to earlier releases of ChEMBL-RDF, it now contains the chemical properties
+provided by the ChEMBL database as calculated by ACD/Labs and XXXXX. The data includes
+chemical properties like polar surface area, pKa and logP, counts for hydrogen bond donor
+and acceptor, and rotational bonds. This data is provided for XXXXX structures.
+
+\subsubsection{Data Statistics and Validation}
+
+John/Anna: common statistics with SPARQL
+
 \subsection{Linked Open Data}
@@
 
+ChemSpider~\cite{Pence2010} is a free online database provided by RSC. It contains over 26 million unique
+chemical compounds aggregated from over 400 data sources as well as chemical data extracted
+from RSC scientific articles and databases. Since its inception, efforts have been made to
+utilize it as both a deposition platform for the community to contribute novel data, as
+well as a platform for annotation and curation for existing data. Studies have shown that there are data
+quality issues in many of the public compound databases~\cite{Williams2011}. ChemSpider has become a
+valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
+is presently providing the chemical structure, substructure and similarity searching services 
+underpinning the Open PHACTS semantic web project~\cite{Williams2012}. Specific chemical data 
+sources containing data mappings between ChemSpider identifiers (CSIDs) and the original data 
+source identifiers have been provided to the triple store, together with chemical identifiers 
+including validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
+The data mappings between the CSIDs and ChEMBL IDs are released to the 
+community under the Creative Common Attribution-Share Alike license (CC-BY-SA 3.0). Attribution 
+should be made to the original ChEMBL database, Open PHACTS, and ChemSpider.
+
 Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
@@
 
-Besides the PubMed identifiers used to link from literature references to Bio2RDF (see earlier),
-ChEMBL uses DOIs alternatively, which we use to link out to the RDF provided by CrossRef~\citep{Bilder2011}:
+In addition to the PubMed identifiers used to link from literature references to Bio2RDF,
+ChEMBL provides DOIs, which we use to link out to the RDF provided by CrossRef~\citep{Bilder2011}:
 
@@
 
-The ChEMBL-RDF is increasingly used, and we here present a few applications. The first application
-describes how it uses the SPARQL end point can be used to make the data available as linked data
+ChEMBL-RDF can be used to explore interrelated scientific datasets, and we here present a few applications. 
+The first application describes how it uses the SPARQL end point can be used to make the data available as linked data
 via the Bio2RDF platform. The second application takes advantage of the bibliographic information
@@
 The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
-based services for standard URI resolution, along with text and link searches
-\cite{WebAppGitHub}. It proxies the standard item
-URIs by translating URLs between those requested by users and the URIs that are available
-in the databases. For example, if the ChEMBL web application is running on the users
+based services for standard URI resolution, along with text and link searches~\cite{Ansell2011}. 
+The configuration files along with build scripts for the ChEMBL-RDF Linked Data server are available on GitHub~\cite{WebAppGitHub}. 
+It proxies the standard item URIs by translating URLs between those requested by users and the URIs that are available
+in SPARQL endpoints. For example, if the ChEMBL web application is running on the users
 local machine, e.g. \url{http://localhost:8080/chembl/}, then a request for the
@@
 \url{http://data.kasabi.com/-dataset/chembl-rdf/13/activity/a31863}. If the
-user requested an RDF document, using content negotiation, the original URIs will be unchanged,
+user requested an RDF document using content negotiation, the original URIs will be unchanged,
 however, if the user requested an HTML document, the results will contain both the
@@
 
-Bioclipse Decision Support (Bioclipse DS)~\cite{Spjuth:2011uq} is a user-oriented tool for providing on-time and on-demand information on chemical structures, based on the Bioclipse workbench.
-Such information can include calculated properties, data from database queries, and results from predictive models.
-Bioclipse DS has previously been demonstrated on predictive modeling in drug safety assessment~\cite{Spjuth:2011uq} and also been linked to invoke and present results from distributed toxicity predictions from the OpenTox infrastructure~\cite{Willighagen:2011kx}.
+Bioclipse Decision Support (Bioclipse DS)~\cite{Spjuth:2011uq} is a user-oriented tool for 
+providing on-time and on-demand information on chemical structures, based on the Bioclipse 
+workbench. Such information can include calculated properties, data from database queries, 
+and results from predictive models. Bioclipse DS has previously been demonstrated on predictive 
+modeling in drug safety assessment~\cite{Spjuth:2011uq} and also been linked to invoke and 
+present results from distributed toxicity predictions from the OpenTox infrastructure~\cite{Willighagen:2011kx}.
 
-In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and Chemspider, which enables users to for chemical structures in Bioclipse look up near neighbors on Chemspider via the Chemspider Web API (SOAP) and query ChEMBL-RDF for properties regarding any measured interactions.
-The results are presented alongside predictive models in Bioclipse, and can be used as decision support when evaluating chemical structures and consider strategies for e.g.
-optimization.
+In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and Chemspider. 
+This enables users browsing chemical structures in Bioclipse to look up near neighbors on Chemspider 
+via the Chemspider Web API (SOAP) and query ChEMBL-RDF for properties regarding any measured interactions. 
+The results are presented alongside predictive models in Bioclipse, and can be used as 
+decision support when evaluating chemical structures and consider strategies for optimization.
 

### 2012-04-16 article.tex
more edits

@@
 
-Although many parts of the CHEMBL-RDF 13 dataset now use community based ontologies, some 
+Although many parts of the ChEMBL-RDF 13 dataset now use community based ontologies, some 
 terms from the previous ad-hoc ontology are still used. These terms were created under the 
@@
 is the type of information that the ChEMBL database extracts from literature.
-In CHEMBL-RDF activities are mostly represented using the custom CHEMBL-RDF vocabulary, where the
+In ChEMBL-RDF activities are mostly represented using the custom ChEMBL-RDF vocabulary, where the
 type, fields and links to other resources are using the \textit{chembl}
@@
 mol:m41
- rdfs:label "CHEMBL406142" , 
+ rdfs:label "ChEMBL406142" , 
    "Bis(3-[1 .... yl]-propionamide)" ;
@@
 
-chemblid:CHEMBL406142
+chemblid:ChEMBL406142
  owl:equivalentClass mol:m41 .
@@
 
-To make the data a proper citizen of the semantic web, we link out to various resources,
-which is visualized in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
-using the new complementary index and OpenMolecules RDF using InChI values, to Bio2RDF
-for protein targets using the uniprot identifier, and CrossRef and Bio2RDF for literature
-references using the DOIs and PubMed identifiers, respectively.
+To integrate ChEMBL-RDF with other RDF versions of scientific datasets, we link out to various resources.
+These links are shown in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
+using the complementary index and OpenMolecules RDF using InChI values. Protein links are given to Bio2RDF
+using the Uniprot identifier. Literature references are directed to CrossRef and Bio2RDF 
+using the DOIs and PubMed identifiers, respectively.
 
@@
 
+%FIXME:
 For protein links out are based on the FOOBAR for the species, the EC code for proteins,
@@
 
-FIXME: should we link out to the new UniProt RDF?
+FIXME: should we link out to the new UniProt RDF? PA: Code is available on the OpenPHACTS branch to link to http://purl.uniprot.org/uniprot/
 
@@
 \begin{verbatim}
-<http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846>
+<http://data.kasabi.com/dataset/chembl-rdf/chemblid/ChEMBL324846>
  skos:exactMatch <http://rdf.chemspider.com/370> .
@@
 to define unique IRIs for molecules~\cite{Bradley2009,Willighagen2011}. While IRIs are theoretically unlimited in length,
-in practice web browsers and servers do limit the length of IRIs. Virtuoso is, unfortunately,
-a system which supports only IRIs of up to a certain length. Therefore, only for smaller molecules
-InChI-based links are created. Almost 1.3 million links are created, looking like:
+in practice web browsers and servers limit the length of IRIs. Virtuoso is, unfortunately,
+a system which supports only IRIs of up to a certain length. Therefore, InChI-based links are only created for smaller molecules. 
+Almost 1.3 million links to OpenMolecules RDF were created in a similar manner to:
 
@@
 The configuration files along with build scripts for the ChEMBL-RDF Linked Data server are available on GitHub~\cite{WebAppGitHub}. 
-It proxies the standard item URIs by translating URLs between those requested by users and the URIs that are available
+It proxies the standard ChEMBL-RDF URIs by translating URLs between those requested by users and the URIs that are available
 in SPARQL endpoints. For example, if the ChEMBL web application is running on the users
@@
 ChEMBL, e.g. \url{http://localhost:8080/chembl/linkstonamespace/-targetns/originalns:identifier},
-and backward links, originating in other databases, such as LODD, Bio2RDF, e.g.
-\url{http://bio2rdf.org/linksns/targetns/originalns:ident-ifier}, or Chem2Bio2RDF.
+and backward links, originating in other databases, such as LODD, Bio2RDF,
+\footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, and Chem2Bio2RDF.
 These services are vital to efficiently navigate the Linked Data web, as it is both
-impractical and inefficient to require users to crawl the web before they can discover all of the
+impractical and inefficient to require users to crawl the entire web before they can discover 
 relevant resources. These services are currently only supplied as web services from ChEMBL and
@@
 providers in the future. Datasets that are available in SPARQL endpoints can be queried for
-links efficiently using templated queries in the ChEMBL web application.
+links efficiently using simple queries as demonstrated in the ChEMBL web application.
 
@@
 
-While we show here that the triple version of the ChEMBL data is very useful, the current triples are by no means static: ChEMBL-RDF
+While we show here that the RDF version of the ChEMBL data is very useful, the current RDF triples are by no means static: ChEMBL-RDF
 will keep evolve, following new developments in the Linked Open Data world. For example, it is likely that over time we will link out
 to more Linked Data resources. More importantly, we wish to adopt more common ontologies, of which the BioAssay Ontology is planned
-to be the first next to be adopted~\cite{Visser2011}. This should make the ChEMBL-RDF triples even more interoperable.
+to be the next~\cite{Visser2011}. These steps will make the ChEMBL-RDF triples even more interoperable.
 

### 2012-04-16 article.tex
Filling in Data Statistics and Validation section

@@
 
-John/Anna: common statistics with SPARQL
+Each release of the ChEMBL database is accompanied by set of release counts, which help 
+provide a concise content overview of the release in question. The ChEMBL 13 release counts
+are as follows:
+
+\begin{itemize}
+  \item 8,845 targets
+  \item 1,143,682 distinct molecules
+  \item 6,933,068 activities
+  \item 617,681 assays
+  \item 44,682 documents
+\end{itemize}
+
+These counts are important as firstly they show continued growth in ChEMBL data, as seen 
+in all previous releases. Secondly, these counts can help validate any transformations, 
+which have been applied to the relational data model traditionally used to store the ChEMBL 
+data. The following SPARQL queries attempt to provide a high level validation of the 
+ChEMBL-RDF by regenerating the ChEMBL 13 release statistics.
+
+The following SPARQL query returns 8,845 ChEMBL targets:
+
+\begin{tiny}
+\begin{verbatim}
+select count(*) WHERE {
+  ?s ?p chembl:Target .
+}
+\end{verbatim}
+\end{tiny}
+
+The following SPARQL query returns 1,143,682 ChEMBL molecules:
+
+NOTE: Full set of ChEMBL molecules have been created by querying molecule_dictionary table
+NOTE: I have added extra triple for each molecule: <http://data.kasabi.com/dataset/chembl-rdf/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
+NOTE: Modified molecule compound structure can be queried here: https://wwwdev.ebi.ac.uk/chembl/sparql
+
+\begin{tiny}
+\begin{verbatim}
+select count(*) WHERE {
+  ?s ?p chembl:Molecule .
+}
+\end{verbatim}
+\end{tiny}
+
+The following SPARQL query returns 617,681 ChEMBL assays:
+
+\begin{tiny}
+\begin{verbatim}
+select count(*) WHERE {
+  ?s ?p chembl:Assay .
+}
+\end{verbatim}
+\end{tiny}
+
+The following SPARQL query returns the 6,933,068 ChEMBL activities:
+
+\begin{tiny}
+\begin{verbatim}
+select count(*) WHERE {
+  ?s ?p chembl:Activity .
+}
+\end{verbatim}
+\end{tiny}
+
+The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 less than appears 
+in ChEMBL 13 release notes, as it does not include doc_id -1, which refers to the 
+unpublished dataset.
+
+\begin{tiny}
+\begin{verbatim}
+select count(*) WHERE {
+  ?s ?p bibo:Article .
+}
+\end{verbatim}
+\end{tiny}
+
+Recreation of the ChEMBL release statistics, by querying the ChEMBL-RDF with SPARQL 
+provides end users with high degree of confidence that the ChEMBL-RDF contains same 
+core content as source relational database.
 

### 2012-04-16 article.tex
Adding Wellcome Trust acknowledgement
Acknowledgement text change
Added Compound Selectivity Use Case. The SPARQL query still causing problems, so may want to include this Use Case

@@
 
+\subsection{Compound Selectivity}
+
+NOTE: May want to exclude this Use Case as query has not successful run yet, keeps causing virtuoso 
+to timeout or run out of memory. Currently playing with local server setup to get query to run.
+
+Designing a molecule, which is selective to one target over another will often be considered a 
+successful outcome in a drug design process. If the target in question is a protein, it is easy
+to understand that as sequence identity amongst residues, which contribute to potential small 
+molecule binding sites remains high so does the issue of selectivity. That is, the molecule being 
+designed may also bind to the equivalent binding site in the  closely related protein 
+target and may lead to undesirable consequences. The ChEMBL data model links molecules to 
+targets using different activity types recorded in the literature. Using activity types, 
+which act as a measure of binding affinity, such as IC50 or Ki and applying activity value cuts
+offs it is possible to identify molecules which have higher binding affinity to certain targets 
+compared to others. Taking this one step further it is possible to identify a set molecules, 
+which have a high affinity to protein A (e.g. IC50 value < 50nM) and low affinity to protein 
+B (e.g. IC50 value > 200nM).
+
+The following SPARQL query identifies a set of molecules, which based on data curated from the 
+literature and stored in the ChEMBL-RDF data model, selectively bind Human Cyclin-Dependent Kinase 2 
+(UniProt: P24941) over Human Cyclin-Dependent Kinase 4 (UniProt: P11802).
+
+NOTE: Query should return 65 molecules
+
+\begin{tiny}
+\begin{verbatim}
+select distinct ?molecule WHERE 
+{ 
+  ?assay ?p chembl:Assay ;
+  chembl:hasTarget ?target .
+  ?activity ?p chembl:Activity ;
+  chembl:forMolecule ?molecule ;
+  chembl:onAssay ?assay ;
+  chembl:type ?type ;
+  chembl:standardUnits ?unit ;
+  chembl:standardValue ?value;
+  chembl:relation ?relation
+  FILTER (
+    ?type = "IC50"  &&
+    ?unit = "nM"    &&
+    ?value  < 50    &&
+    ?relation = "="
+  ) .
+  ?target ?p chembl:Target;
+  dc:identifier ?id
+  FILTER (
+    ?id = "uniprot:P24941"
+  ) .
+  { 
+    select ?molecule WHERE 
+    {
+      ?assay ?p chembl:Assay ;
+      chembl:hasTarget ?target .
+      ?activity ?p chembl:Activity ;
+      chembl:forMolecule ?molecule ;
+      chembl:onAssay ?assay;
+      chembl:type ?type ;
+      chembl:standardUnits ?unit ;
+      chembl:standardValue ?value ;
+      chembl:relation ?relation
+      FILTER (
+        ?type = "IC50"  &&
+        ?unit = "nM"    &&
+        ?value  > 200   &&
+        ?relation = "="
+      ) .
+      ?target ?p chembl:Target;
+      dc:identifier ?id
+      FILTER (
+        ?id = "uniprot:P11802"
+      ) .
+    }
+  }
+}
+\end{verbatim}
+\end{tiny}
+
+Analysis on the set of molecules returned by this query can be used to help identify small 
+molecule features, which may increase target-binding specificity. For queries which link 
+small molecules to targets, by traversing bioactivity data in the ChEMBL database, it is 
+also important to consider the parameters associated with the assay to target mappings. 
+These additional parameters include a relationship type, a multi flag (for poorly defined
+targets) a complex flag (for protein complex targets) and a curation level. These different 
+factors are summarized in the ChEMBL confidence score, which ranges from 9 (direct single 
+protein target) to 0 (uncurated). In order to return the largest possible dataset, the 
+confidence score has been ignored in this example compound selectivity use case.
+
 \section{Discussion}
@@
 support.
+
+The data in ChEMBL is made available by funding from the Wellcome Trust [086151/Z/08/Z]
 

### 2012-04-17 article.tex
Added statement of funding.
fix latex syntax and hide internal notes

@@
 
-NOTE: Full set of ChEMBL molecules have been created by querying molecule_dictionary table
-NOTE: I have added extra triple for each molecule: <http://data.kasabi.com/dataset/chembl-rdf/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
-NOTE: Modified molecule compound structure can be queried here: https://wwwdev.ebi.ac.uk/chembl/sparql
+%Mark Davies notes:
+%NOTE: Full set of ChEMBL molecules have been created by querying molecule\_dictionary table
+%NOTE: I have added extra triple for each molecule: <http://data.kasabi.com/dataset/chembl-rdf/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
+%NOTE: Modified molecule compound structure can be queried here: https://wwwdev.ebi.ac.uk/chembl/sparql
 
@@
 The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 less than appears 
-in ChEMBL 13 release notes, as it does not include doc_id -1, which refers to the 
+in ChEMBL 13 release notes, as it does not include doc\_id -1, which refers to the 
 unpublished dataset.
@@
 Tim Hodson and Zach Beauvais of Kasabi and A. L\"ovgren at the at Uppsala Biomedical Centre at Uppsala University for their
-support.
+support. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE.
 

### 2012-04-18 article.tex
make it clear that there is (or will be soon) a WAR file available for download from GitHub
Added screenshot for Bioclipse-DS.

@@
 based services for standard URI resolution, along with text and link searches~\cite{Ansell2011}. 
-The configuration files along with build scripts for the ChEMBL-RDF Linked Data server are available on GitHub~\cite{WebAppGitHub}. 
+A Java Web Archive (WAR) file along with the configuration files and build scripts for the ChEMBL-RDF Linked Data server are available on GitHub~\cite{WebAppGitHub}. 
 It proxies the standard ChEMBL-RDF URIs by translating URLs between those requested by users and the URIs that are available
@@
 
-\subsection{Decision Support}
+\subsection{Bioclipse Decision Support}
 
@@
 
-Bioclipse Decision Support (Bioclipse DS)~\cite{Spjuth:2011uq} is a user-oriented tool for 
-providing on-time and on-demand information on chemical structures, based on the Bioclipse 
-workbench. Such information can include calculated properties, data from database queries, 
+Bioclipse Decision Support (Bioclipse DS)~\cite{Spjuth:2011uq} is a user-oriented tool based on the Bioclipse 
+workbench for providing on-time and on-demand information on chemical structures. 
+Such information can include calculated properties, data from database queries, 
 and results from predictive models. Bioclipse DS has previously been demonstrated on predictive 
@@
 In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and Chemspider. 
-This enables users browsing chemical structures in Bioclipse to look up near neighbors on Chemspider 
-via the Chemspider Web API (SOAP) and query ChEMBL-RDF for properties regarding any measured interactions. 
-The results are presented alongside predictive models in Bioclipse, and can be used as 
-decision support when evaluating chemical structures and consider strategies for optimization.
+This enables users browsing chemical structures in Bioclipse to look up near neighbors in Chemspider 
+via the Chemspider Web API (SOAP), and for the found compounds to query ChEMBL-RDF for 
+available interaction data.
+The results are presented alongside predictive models in Bioclipse (see Fig \ref{fig:bioclipse-ds}, and can be used for decision support when evaluating chemical structures and consider
+strategies for optimization.
 
-[TODO: SCREENSHOT from Bioclipse-DS-ChEMBL-RDF]
+\begin{figure*}[!ht]
+		\begin{center}
+		\includegraphics[width=16cm]{bioclipse-ds.png}
+		\newline
+		\caption[wee]{Screnshot from Bioclipse Decision Support with results from a Chsmpider + ChEMBL-RDF search.
+The top middle canvas contains the query structure, the top right canvas shows the near neighbors in Chemspider (via a similarStructure search), the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower center canvas shows the found interactions for this compound.}
+	\label{fig:bioclipse-ds}
+		\end{center}
+\end{figure*}
 

### 2012-05-01 article.tex
Removed a typo
Changed to the preferred reference
Consistent spelling of ChemSpider

@@
 
-In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and Chemspider. 
-This enables users browsing chemical structures in Bioclipse to look up near neighbors in Chemspider 
-via the Chemspider Web API (SOAP), and for the found compounds to query ChEMBL-RDF for 
+In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and ChemSpider. 
+This enables users browsing chemical structures in Bioclipse to look up near neighbors in ChemSpider 
+via the ChemSpider Web API (SOAP), and for the found compounds to query ChEMBL-RDF for 
 available interaction data.
@@
 		\newline
-		\caption[wee]{Screnshot from Bioclipse Decision Support with results from a Chsmpider + ChEMBL-RDF search.
-The top middle canvas contains the query structure, the top right canvas shows the near neighbors in Chemspider (via a similarStructure search), the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower center canvas shows the found interactions for this compound.}
+		\caption[wee]{Screnshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
+The top middle canvas contains the query structure, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search), the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower center canvas shows the found interactions for this compound.}
 	\label{fig:bioclipse-ds}
@@
 
-Tim Hodson and Zach Beauvais of Kasabi and A. L\"ovgren at the at Uppsala Biomedical Centre at Uppsala University for their
+Tim Hodson and Zach Beauvais of Kasabi and A. L\"ovgren at the BMC Computing Department at Uppsala University for their
 support. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE.

### 2012-05-01 article.tex
Clarified 'free'. Being first author, the use has to be more exact than normally

@@
 
-ChemSpider~\cite{Pence2010} is a free online database provided by RSC. It contains over 26 million unique
+ChemSpider~\cite{Pence2010} is a freely accessible, online database provided by RSC. It contains over 26 million unique
 chemical compounds aggregated from over 400 data sources as well as chemical data extracted
@@
 source identifiers have been provided to the triple store, together with chemical identifiers 
-including validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
+including validated chemical names (systematic, generic and trivial), SMILES, and InChIs. 
 The data mappings between the CSIDs and ChEMBL IDs are released to the 

### 2012-05-01 article.tex
Migrated to the BMC template, as we are now submitting to JChemInf

@@
-% PLEASE USE THIS FILE AS A TEMPLATE
-% Check file iosart2c.tex for more examples
-%
-% Journal:
-%   Journal of Ambient Intelligence and Smart Environments (jaise)
-%   Web Intelligence and Agent Systems: An International Journal (wias)
-%   Semantic Web: Interoperability, Usability, Applicability (SW)
-% IOS Press
-% Latex 2e
+\documentclass[10pt]{bmc_article}
 
-% options: jaise|wias|sw
-% add. options: [seceqn,secfloat,secthm,crcready,onecolumn]
+\usepackage{a4wide}  % Formatting web addresses  
+\usepackage{url}  % Formatting web addresses  
+\usepackage{ifthen}  % Conditional 
+\usepackage{multicol}   %Columns
+\usepackage[utf8]{inputenc} %unicode support
+%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
+%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
+\urlstyle{rm}
+\usepackage[citestyle=numeric-comp]{biblatex}                                                                     
+\setlength{\bibitemsep}{0cm}                                                                                                                
+\renewbibmacro{in:}{%                                                                                                                       
+  \ifentrytype{article}{}{%                                                                                                                 
+  \printtext{\bibstring{in}\intitlepunct}}}                                                                                                 
+\addbibresource{article.bib}
 
-\documentclass[sw]{iosart2c}
-
-%\documentclass[sw]{iosart2c}
-%\documentclass[wias]{iosart2c}
-%\documentclass[jaise]{iosart2c}
-
-\usepackage[T1]{fontenc}
-\usepackage{times}%
-\usepackage{natbib}% for bibliography sorting/compressing
 %\usepackage{amsmath}
@@
 %\usepackage{tikz}
-%\usetikzlibrary{snakes,arrows,shapes}
 
-%%%%%%%%%%% Put your definitions here
+\newboolean{publ}
 
-% \bibliographystyle{plain} 
-\bibliographystyle{unsrt}
+%Review style settings
+\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}
 
-%%%%%%%%%%% End of definitions
-
-\pubyear{0000}
-\volume{0}
-\firstpage{1}
-\lastpage{1}
+%Publication style settings
+%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}
 
 \begin{document}
-
-\newcommand{\url}[1]{#1}
-
-\begin{frontmatter}
+\begin{bmcformat}
 
@@
 \title{The ChEMBL database as Linked Open Data}
-\runningtitle{ChEMBL-RDF}
-%\subtitle{}
 
-%\review{}{}{}
+\author{
+  Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
+  \email{Egon L Willighagen - egon.willighagen@maastrichtuniversity.nl}
+\and
+  Andra Waagmeester$^1$%
+  \email{Andra Waagmeester - FIXME}
+\and
+  Ola Spjuth$^2$%
+  \email{Ola Spjuth - FIXME}
+\and
+  Peter Ansell$^3$%
+  \email{Peter Ansell - FIXME}
+\and
+  Antony Williams$^4$%
+  \email{Antony Williams - FIXME}
+\and
+  Valery Tkachenko$^4$%
+  \email{Valery Tkachenko - FIXME}
+\and
+  Janna Hastings$^5$%
+  \email{Janna Hastings - FIXME}
+\and
+  John Overington$^6$%
+  \email{John Overington - FIXME}
+\and
+  Anna Gaulton$^6$%
+  \email{Anna Gaulton - FIXME}
+\and
+  Mark Davies$^6$%
+  \email{Mark Davies - FIXME}
+\and
+  Bin Chen$^7$%
+  \email{Bin Chen - FIXME}
+\and
+  David Wild$^7$%
+  \email{David Wild - FIXME}
+}
 
-% For one author:
-\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Corresponding author.\\ E-mail: egon.willighagen@maastrichtuniversity.nl.}}
-\runningauthor{E.L. Willighagen et al.}
-% Two or more authors (order to be determined):
-\author[A]{\fnms{Andra} \snm{Waagmeester}}
-\author[B]{\fnms{Ola} \snm{Spjuth}},
-\author[C]{\fnms{Peter} \snm{Ansell}},
-\author[D]{\fnms{Antony} \snm{Williams}},
-\author[D]{\fnms{Valery} \snm{Tkachenko}},
-\author[E]{\fnms{Janna} \snm{Hastings}},
-\author[F]{\fnms{John} \snm{Overington}},
-\author[F]{\fnms{Anna} \snm{Gaulton}},
-\author[F]{\fnms{Mark} \snm{Davies}},
-\author[G]{\fnms{Bin} \snm{Chen}},
-\author[G]{\fnms{David} \snm{Wild}}
+\address{
+    \iid(1)Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands  \\
+    \iid(2)Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden \\
+    \iid(3)University of Queensland, St Lucia, Qld 4072, Australia \\
+    \iid(4)Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A. \\
+    \iid(5)Chemoinformatics and Metabolism, European Bioinformatics Institute, POSTAL CODE, Hinxton, United Kingdom \\
+    \iid(6)EMBL-European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\ 
+    \iid(7)School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.
+}
 
-\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands}
-\address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
-\address[C]{University of Queensland, St Lucia, Qld 4072, Australia}
-\address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
-\address[E]{Chemoinformatics and Metabolism, European Bioinformatics Institute, POSTAL CODE, Hinxton, United Kingdom}
-\address[F]{EMBL-European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom}
-\address[G]{School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.}
+\maketitle
 
 \begin{abstract}
+\paragraph*{Background:}
 Making data available as Linked Data using RDF is advantageous because it promotes integration with other web resources.
@@
 using extensible vocabularies for additional information. This combination provides a new, open, interface for scientists to use ChEMBL.
+\paragraph*{Results:}
 This paper describes the continued conversion of data from the ChEMBL database into RDF triples.
@@
 other web resources using semantic web technologies.
+\paragraph*{Conclusions:}
 \end{abstract}
-
-\begin{keyword}
-% \sep 
-Resource Description Framework, ChEMBL
-\end{keyword}
-
-\end{frontmatter}
 
@@
 
-\section{Introduction}\label{s1}
+\section*{Introduction}\label{s1}
 
@@
 
-\section{Methods}\label{s2}
+\section*{Methods}\label{s2}
 
@@
 were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
-and are available from the source code hosting service, GitHub~\citep{ChEMBLRDFGitHub}. This process 
+and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
 has been outlined in a best practices note by the W3C Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.
@@
 
-\section{Results}\label{s3}
+\section*{Results}\label{s3}
 
@@
 In addition to the PubMed identifiers used to link from literature references to Bio2RDF,
-ChEMBL provides DOIs, which we use to link out to the RDF provided by CrossRef~\citep{Bilder2011}:
+ChEMBL provides DOIs, which we use to link out to the RDF provided by CrossRef~\cite{Bilder2011}:
 
@@
 
-\section{Applications}
+\section*{Applications}
 
@@
 
-\section{Discussion}
+\section*{Discussion}
 
@@
 
-\section{Acknowledgements}
+\section*{Authors contributions}
+EW initiated the project, created the initial RDF version of the ChEMBL data, and encouraged the use cases.
+AW extended CitedIn to support citation info in ChEMBL-RDF.
+OS, AW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
+JH supported the project with CHEMINF representations.
+PA integrated ChEMBL-RDF into Bio2RDF.
+JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
+All authors contributed to the continued development of ChEMBL-RDF, as well as to the writing of the paper,
+and approved the final version.
+
+\section*{Acknowledgements}
 
@@
 
-\bibliography{article}
+{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
+\printbibliography
+}
 
-%%%%%%%%%%% The bibliography starts:
-%\begin{thebibliography}{9}
-%
-%\bibitem{r1}
-%
-%\bibitem{r2}
-%
-%\end{thebibliography}
+\ifthenelse{\boolean{publ}}{\end{multicols}}{}
 
+\end{bmcformat}
 \end{document}

### 2012-05-01 article.tex
Cleaned up author lines, fixing some whitespace issues

@@
 
-\author{
-  Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
-  \email{Egon L Willighagen - egon.willighagen@maastrichtuniversity.nl}
-\and
-  Andra Waagmeester$^1$%
-  \email{Andra Waagmeester - FIXME}
-\and
-  Ola Spjuth$^2$%
-  \email{Ola Spjuth - FIXME}
-\and
-  Peter Ansell$^3$%
-  \email{Peter Ansell - FIXME}
-\and
-  Antony Williams$^4$%
-  \email{Antony Williams - FIXME}
-\and
-  Valery Tkachenko$^4$%
-  \email{Valery Tkachenko - FIXME}
-\and
-  Janna Hastings$^5$%
-  \email{Janna Hastings - FIXME}
-\and
-  John Overington$^6$%
-  \email{John Overington - FIXME}
-\and
-  Anna Gaulton$^6$%
-  \email{Anna Gaulton - FIXME}
-\and
-  Mark Davies$^6$%
-  \email{Mark Davies - FIXME}
-\and
-  Bin Chen$^7$%
-  \email{Bin Chen - FIXME}
-\and
-  David Wild$^7$%
-  \email{David Wild - FIXME}
+\author{Egon~L.~Willighagen$^{1}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl}\and
+Andra~Waagmeester$^1$\email{Andra Waagmeester - FIXME}\and
+Ola Spjuth$^2$\email{Ola Spjuth - FIXME}\and
+Peter~Ansell$^3$\email{Peter Ansell - FIXME}\and
+Antony~Williams$^4$\email{Antony Williams - FIXME}\and
+Valery~Tkachenko$^4$\email{Valery Tkachenko - FIXME}\and
+Janna~Hastings$^5$\email{Janna Hastings - FIXME}\and
+John~Overington$^6$\email{John Overington - FIXME}\and
+Anna~Gaulton$^6$\email{Anna Gaulton - FIXME}\and
+Mark~Davies$^6$\email{Mark Davies - FIXME}\and
+Bin~Chen$^7$\email{Bin Chen - FIXME}\and
+David~Wild$^7$\email{David Wild - FIXME}
 }
 
-\address{
-    \iid(1)Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands  \\
+\address{\iid(1)Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands  \\
 \iid(2)Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden \\

### 2012-05-02 article.tex
First citation is ref. 1
add email address

@@
 \urlstyle{rm}
-\usepackage[citestyle=numeric-comp]{biblatex}                                                                     
+\usepackage[citestyle=numeric-comp,sorting=none]{biblatex}
 \setlength{\bibitemsep}{0cm}                                                                                                                
@@
 Ola Spjuth$^2$\email{Ola Spjuth - FIXME}\and
-Peter~Ansell$^3$\email{Peter Ansell - FIXME}\and
+Peter~Ansell$^3$\email{Peter Ansell - p\_ansell@yahoo.com}\and
 Antony~Williams$^4$\email{Antony Williams - FIXME}\and

### 2012-05-28 article.tex
Small tweakings

@@
 For example, systems biologists integrate micro-array differential expression datasets to 
-biological pathways, using various other datasets to provide evidence for the links~\cite{}. 
+biological pathways, using various other datasets to provide evidence for the links~\cite{FIXME}. 
 Another prominent example is drug discovery where a new unique chemical entity is discovered 
@@
 The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
-The SQL datadump provided by EMBL was inserted into a local MySQL database. A set of SQL queries 
+The SQL datadump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
 were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
 and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
-has been outlined in a best practices note by the W3C Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.
+has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest Group~\cite{Marshall2012}.
 
@@
 \caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
+\begin{center}
 \begin{tabular}{ll}
@@
 \end{tabular}
+\end{center}
 \end{table*}
@@
 \begin{figure}[t]
-\includegraphics[width=0.45\textwidth]{figs/relations}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figs/relations}
 \caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
 of common classes, while others are instances.}\label{f1}
+\end{center}
 \end{figure}
@@
 
-\begin{footnotesize}
+\begin{small}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{footnotesize}
+\end{small}
 
@@
 
-\begin{tiny}
+\begin{small}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{tiny}
+\end{small}
 
@@
 \begin{figure}[t]
+{\center
 \includegraphics[width=0.45\textwidth]{figs/lodgraph}
@@
 Edges are labeled by the predicates making the links.}\label{2}
+}
 \end{figure}

### 2012-05-28 article.tex
More tuning: added two references, unnumbered sections

@@
 For example, systems biologists integrate micro-array differential expression datasets to 
-biological pathways, using various other datasets to provide evidence for the links~\cite{FIXME}. 
+biological pathways, using various other datasets to provide evidence for the links~\cite{Staab2007}. 
 Another prominent example is drug discovery where a new unique chemical entity is discovered 
@@
 
-\subsection{ChEMBL-RDF}
+\subsection*{ChEMBL-RDF}
 
-\subsubsection{Data Structure}
+\subsubsection*{Data Structure}
 
@@
 
-\subsubsection{Data Statistics and Validation}
+\subsubsection*{Data Statistics and Validation}
 
@@
 are as follows:
-
 \begin{itemize}
@@
 
-\subsection{Linked Open Data}
+\subsection*{Linked Open Data}
 
@@
 
-\subsubsection{Linking out to Bio2RDF}
+\subsubsection*{Linking out to Bio2RDF}
 
@@
 \end{verbatim}
-\end{tiny}
+\end{small}
 
-\subsubsection{Linking out to ChemSpider}
+\subsubsection*{Linking out to ChemSpider}
 
@@
 \end{verbatim}
-\end{tiny}
+\end{small}
 
-\subsubsection{Linking out to OpenMolecules RDF}
+\subsubsection*{Linking out to OpenMolecules RDF}
 
@@
 
-\subsubsection{Linking out to CrossRef}
+\subsubsection*{Linking out to CrossRef}
 
@@
 
-\subsection{Bio2RDF}
+\subsection*{Bio2RDF}
 
@@
 
-\subsection{CitedIn}
+\subsection*{CitedIn}
 
@@
 
-\subsection{Bioclipse Decision Support}
+\subsection*{Bioclipse Decision Support}
 
@@
 
-\subsection{Chem2Bio2RDF}
+\subsection*{Chem2Bio2RDF}
 
@@
 
-\subsection{Compound Selectivity}
+\subsection*{Compound Selectivity}
 
@@
 to more Linked Data resources. More importantly, we wish to adopt more common ontologies, of which the BioAssay Ontology is planned
-to be the next~\cite{Visser2011}. These steps will make the ChEMBL-RDF triples even more interoperable.
+to be the next~\cite{Visser2011}. These steps will make the ChEMBL-RDF triples even more interoperable. Part of this future work
+is already being performed within the context of the Open PHACTS project~\cite{Williams2012}.
 

### 2012-07-09 article.tex
Tuned up the abstract

@@
 \paragraph*{Background:}
-Making data available as Linked Data using RDF is advantageous because it promotes integration with other web resources.
-RDF documents can natively link to related data, and others can link back using URIs. RDF makes the data machine-readable,
-using extensible vocabularies for additional information. This combination provides a new, open, interface for scientists to use ChEMBL.
+Making data available as Linked Data using Resource Description Framework (RDF) promotes integration with other web resources.
+RDF documents can natively link to related data, and others can link back using Universal Resource Identifiers (URIs).
+RDF makes the data machine-readable, using extensible vocabularies for additional information, making it easier
+to scale up inference and data analysis.
 \paragraph*{Results:}
-This paper describes the continued conversion of data from the ChEMBL database into RDF triples.
+This paper describes the continued project to convert data from the ChEMBL database into RDF triples.
 This updated version of ChEMBL-RDF now uses recently introduced ontologies, including CHEMINF and CiTO,
-exposes more information from the database, and is now available as dereferencable, Linked Data.
-To demonstrate the new features, we present new use cases showing the benefits of integration with
-other web resources using semantic web technologies.
+exposes more information from the database, and is now available as dereferencable, linked data.
+To demonstrate these new features, we present novel use cases showing further integration with
+other web resources, including Bio2RDF, Chem2Bio2RDF, and ChemSpider, and showing the use of standard
+ontologies.
 \paragraph*{Conclusions:}
+The use cases show that by using open standards and ontologies it is possible to link the ChEMBL database
+to other databases, such that the integration can understood by software. 
 \end{abstract}

### 2012-07-09 article.tex
Workaround for a stupid bug in TexMaker that does not find the end of the section title properly
Tuned the intro and methods section

@@
 
-\section*{Introduction}\label{s1}
+\section*{Introduction}
+\label{s1}
 
-The current scientific data deluge, in which datasets are growing faster than scientists can analyse and 
-curate them, provides several new challenges for information systems. Scientists wish to 
+The current scientific data deluge, in which datasets grow faster in size and complexity than scientists can handle,
+provides several new challenges for information systems. Scientists wish to 
 discover new, unique, and significant patterns in datasets, which explaining biological 
@@
 The former requires scientists to import the data into a relational database, while the 
-latter limits the machine access to the data. Two independent teams have previously mapped 
-the ChEMBL dataset to RDF, Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}.
+latter limits the machine access to the data.
 
-This paper presents the current state of the ChEMBL-RDF dataset, details of the latest structures
-and ontologies used to map version 13 of ChEMBL to RDF, along with new example use cases, showing 
+Using the right to download, modify, and redistribute the ChEMBL data, two independent teams
+have previously mapped  the ChEMBL dataset to the Resource Description Framework (RDF),
+Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}. RDF is a framework where
+knowledge is represented by small, so-called triples reflecting very basic facts. Things
+in RDF are represented by Universal Resource Identifiers (URI) and when the more specialized
+Universal Resource Locator (URL) is used, it creates \emph{linked data}~\cite{Samwald2011}. RDF is
+developed as an open standard by the World Wide Web Consortium, and provides a knowledge
+representation framework that applies to many, if not any, domain, making it a perfect
+approach for data integration in the life sciences.
+
+This paper presents an update on the ChEMBL-RDF dataset, and details of the latest structures
+and ontologies used to map ChEMBL 13 to RDF, along with new use cases, showing 
 links to other datasets using their RDF URIs to further support research in the life sciences.
 
-\section*{Methods}\label{s2}
+\section*{Methods}
+\label{s2}
 
@@
 and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
-has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest Group~\cite{Marshall2012}.
+has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
+Group~\cite{Marshall2012}. This process has not changed much.
 
-In the previous RDF conversion the RDF triples in ChEMBL-RDF were represented using properties and c
-lasses from a custom, ad-hoc ontology~\cite{Willighagen2011}. The current version used community-proposed
-ontologies, making the RDF more interoperable. The ChEMBL-RDF 13 dataset uses ontology
-standards such as the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology~\cite{Hastings2011} for literature
-references, and domain ontologies like the Protein Ontology\cite{Sidhu2006} and the Chemical Information
-Ontology\cite{Hastings2011}. Throughout this paper, various shorthand prefixes are used to denote different ontology 
+The previous ChEMBL-RDF conversion used, however, custom predicates and classes to represent RDF triples, defining
+an ad-hoc ontology~\cite{Willighagen2011}. The current version, instead, uses community-proposed
+ontologies, making the RDF more interoperable. The new triple dataset uses standard ontologies
+including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology~\cite{Hastings2011} for literature
+references and citations, and domain ontologies like the Protein Ontology\cite{Sidhu2006} and the Chemical Information
+Ontology\cite{Hastings2011}. Throughout this paper, various prefixes are used to denote different ontology 
 namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.
@@
 been made dereferencable using the Kasabi platform~\cite{kasabi}. The Kasabi hosting provides a SPARQL
-end point which requires a user account. A user account can be free for casual usage, or paid for consistent usage.
+end point which requires a user account. A user account can be acquired for free for casual usage,
+or paid for uncapped  use.
 
@@
 
-\section*{Results}\label{s3}
+\section*{Results}
+\label{s3}
 

### 2012-07-09 article.tex
Cleaning up the first part of the Results section

@@
 
-\subsection*{ChEMBL-RDF}
-
-\subsubsection*{Data Structure}
+\subsection*{Data Structure}
 
@@
 data available in the relational database. Figure~\ref{f1} shows how the various resource
-classes are linked together. This section will show how data in the ChEMBL database
-is exposed at an triple level.
+classes are linked together, detailed in this section at a triple level.
 
@@
 is the type of information that the ChEMBL database extracts from literature.
-In ChEMBL-RDF activities are mostly represented using the custom ChEMBL-RDF vocabulary, where the
-type, fields and links to other resources are using the \textit{chembl}
+Standard ontologies are being explored for future release for annotation
+of activities, but, currently, the
+type, fields and links to other resources are still using the \textit{chembl}
 namespace. While ChEMBL provides both the original activities as found in literature
 and standardized values allowing comparison between studies, the triples only
-make the latter available. The CiTO ontology is used to point to the paper from
-which the data was extracted.
+make the latter available. The link to literature is encoded using the CiTO ontology.
+This results in a set of triples that look like:
 
@@
 
-More than five thousand different activity types are captured by the ChEMBL database.
+More than five thousand different activity types, represented by the chembl:type predicate,
+are captured by the ChEMBL database.
 The top five types are Potency (43\%), IC50 (13\%), MIC (4.6\%), Inhibition (3.7\%),
@@
 resource type. Various assay types are found in the database: chembl:ADMET, chembl:Binding,
-chembl:Functional, chembl:Property, and chembl:Unassigned. The assay information link
-activities to a target (see below) and has a short description. The confidence information
-assigned by the ChEMBL is exposed too, and was previously used to improve activity
-prediction using Bayesian statistics~\cite{Willighagen2011}.
+chembl:Functional, chembl:Property, and chembl:Unassigned.
 
@@
  chembl:hasAssayType chembl:ADMET ;
- chembl:hasConfScore "7"^^xsd:int ;
  chembl:hasDescription
-  "Inhibition of ... hydroxylase" ;
- chembl:hasTarget target:t100122 .
+  "Inhibition of ... hydroxylase" .
 \end{verbatim}
@@
 
-The assays measure activities against particular biological targets. The ChEMBL database
+Each assay measures activities against one or more particular biological targets. To each
+target the ChEMBL database associates a confidence score. Because this score is specific
+for each target, the following construct is used:
+
+\begin{small}
+\begin{verbatim}
+assay:a17 chembl:hasTargetScore assay:a17/score/t100122 .
+assay:a17/score/t100122
+ chembl:forTarget target:t100122 ;
+ chembl:hasConfScore "7"^^xsd:int .  
+\end{verbatim}
+\end{small}
+
+The ChEMBL database
 recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
@@
 
-\subsubsection{Data Statistics}
-
-\# triples
-\# links out
-\# problems
-
-The full set of triples are available from the website http://semantics.bigcat.unimaas.nl/chembl12/.
-
-FIXME!
-
-In contrast to earlier releases of ChEMBL-RDF, it now contains the chemical properties
-provided by the ChEMBL database as calculated by ACD/Labs and XXXXX. The data includes
-chemical properties like polar surface area, pKa and logP, counts for hydrogen bond donor
-and acceptor, and rotational bonds. This data is provided for XXXXX structures.
-
-\subsubsection*{Data Statistics and Validation}
+\subsection*{Data Statistics and Validation}
 

### 2012-07-09 article.tex
Tuned more results

@@
 chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
-explicit types.
+explicit types, and effectively uses a closed-world approach. A target
+specification can then look like:
 
@@
 but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
-even cells. All compounds are represented as classes, following
-the design of the CHEMINF ontology. These entities do not have a common
-superclass in common, but can easily be identified as having the 
+even cells. ChEMBL-RDF follows the approach used by the CHEMINF ontology,
+and encodes drugs as classes, rather than instances, and they subclass
+classes defined by several ontologies for these drug types. For example, proteins subclass the
+Protein concept from the PRotein Ontology (PR\_000000001), small
+molecules subclass the Chemical Entity concept from the CHEMINF
+ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
+subclass their respective matches in the CHEBI ontology (CHEBI\_50699
+and CHEBI\_7754).
+
+It should also be noted that the drugs are not subclassed from
+a general drug class, or any other common superclass. However, to annotate
+the drugs to be drugs, they are identified as having the 
 the role of being a drug. This is triplified using the OBO and ChEBI
-ontologies:
+ontologies, where CHEBI\_23888 is the drug role:
 
@@
 
-The entity typing is expressed
-by subclassing other classes. For example, proteins subclass the
-Protein concept from the PRotein Ontology (PR\_000000001), small
-molecules subclass the Chemical Entity concept from the CHEMINF
-ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
-subclass their respective matches in the CHEBI ontology (CHEBI\_50699
-and CHEBI\_7754). For each drug compound the name and synonyms are
+The name and synonyms for each drug are
 provided as labels. When InChI and InChIKeys are available for a drug, then these are
@@
 
-For small molecules, molecular properties are often available from ChEMBL, and as of ChEMBL-RDF 13
+Similarly, properties of small molecules are available from ChEMBL, and as of ChEMBL-RDF 13
 these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
@@
 
-For documents little information is replicated from the database, taking advantage
-of PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs). URIs for the
-latter can be resolved online, providing curated information on the identified
-documents. For each paper, basic properties are provided using the BIBO
-ontology:
+The literature from which all this data was extracted, is replicated from the database,
+but only a subset of information about that literature is provided. In all cases,
+it uses the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) to link
+to authoritative databases to provide full detail, as outlined later in this paper.
+The provided properties are using the BIBO ontology:
 
@@
 \begin{verbatim}
-journal:j6c706049c2e08871b7c46a6528065736
- a bibo:Journal ;
- dc:title "J. Med. Chem." .
-
 res:r1
@@
  bibo:volume "47" .
+
+journal:j6c706049c2e08871b7c46a6528065736
+ a bibo:Journal ;
+ dc:title "J. Med. Chem." .
 \end{verbatim}

### 2012-07-09 article.tex
Final tunes of the structure description

@@
 These links are shown in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
-using the complementary index and OpenMolecules RDF using InChI values. Protein links are given to Bio2RDF
-using the Uniprot identifier. Literature references are directed to CrossRef and Bio2RDF 
+using the complementary index and OpenMolecules RDF using InChI values. Protein links are given to Bio2RDF~\cite{Belleau2008}
+using the Uniprot identifier~\cite{TheUniProtConsortium2010}. Literature references are directed to CrossRef and Bio2RDF 
 using the DOIs and PubMed identifiers, respectively.
@@
 %FIXME:
-For protein links out are based on the FOOBAR for the species, the EC code for proteins,
-as well as UniProt identifiers.
-
-FIXME: should we link out to the new UniProt RDF? PA: Code is available on the OpenPHACTS branch to link to http://purl.uniprot.org/uniprot/
+%For protein links out are based on the FOOBAR for the species, the EC code for proteins,
+%as well as UniProt identifiers.
+%
+%FIXME: should we link out to the new UniProt RDF? PA: Code is available on the OpenPHACTS branch to link to http://purl.uniprot.org/uniprot/
+%
+% Egon: I suggest to skip that for now, as I don't want to taint these results with Open PHACTS work
 

### 2012-07-09 article.tex
Updated the URI to match the current kasabi hosting

@@
 
-ChemSpider~\cite{Pence2010} is a freely accessible, online database provided by RSC. It contains over 26 million unique
+ChemSpider~\cite{Pence2010} is a freely accessible, online database provided by Royal Society of Chemistry (RSC).
+It contains over 26 million unique
 chemical compounds aggregated from over 400 data sources as well as chemical data extracted
@@
 will be resolved from the database using the full original URI,
-\url{http://data.kasabi.com/-dataset/chembl-rdf/13/activity/a31863}. If the
+\url{http://data.kasabi.com/dataset/chembl-rdf/activity/a31863}. If the
 user requested an RDF document using content negotiation, the original URIs will be unchanged,

### 2012-07-09 article.tex
More tuning

@@
 via the ChemSpider Web API (SOAP), and for the found compounds to query ChEMBL-RDF for 
-available interaction data.
-The results are presented alongside predictive models in Bioclipse (see Fig \ref{fig:bioclipse-ds}, and can be used for decision support when evaluating chemical structures and consider
-strategies for optimization.
+available interaction data. The results are presented alongside predictive models in Bioclipse
+(see Fig~\ref{fig:bioclipse-ds}), and can be used for decision support when evaluating chemical
+structures and consider strategies for optimization.
 
@@
 targets using different activity types recorded in the literature. Using activity types, 
-which act as a measure of binding affinity, such as IC50 or Ki and applying activity value cuts
+which act as a measure of binding affinity, such as IC$_{50}$ or K$_{i}$ and applying activity value cuts
 offs it is possible to identify molecules which have higher binding affinity to certain targets 
 compared to others. Taking this one step further it is possible to identify a set molecules, 
-which have a high affinity to protein A (e.g. IC50 value < 50nM) and low affinity to protein 
-B (e.g. IC50 value > 200nM).
+which have a high affinity to protein A (e.g. IC$_{50}$ value $<$ 50~nM) and low affinity to protein 
+B (e.g. IC$_{50}$ value $>$ 200~nM).
 

### 2012-07-10 article.tex
Updated DS screenshot and caption.

@@
 		\begin{center}
-		\includegraphics[width=16cm]{bioclipse-ds.png}
+		\includegraphics[width=15cm]{bioclipse-ds.png}
 		\newline
 		\caption[wee]{Screnshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
-The top middle canvas contains the query structure, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search), the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower center canvas shows the found interactions for this compound.}
+The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
 	\label{fig:bioclipse-ds}

### 2012-07-17 article.tex
Updated the URIs to match linkedchemistry.info

@@
 \multicolumn{2}{l}{\textbf{ChEMBL-RDF Prefixes }}\\
-act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
-assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
-mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
-res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
+act    & http://linkedchemistry.info/chembl/activity/ \\
+assay  & http://linkedchemistry.info/chembl/assay/ \\
+mol    & http://linkedchemistry.info/chembl/molecule/ \\
+res    & http://linkedchemistry.info/chembl/resource/ \\
 \hline
@@
 %NOTE: Full set of ChEMBL molecules have been created by querying molecule\_dictionary table
-%NOTE: I have added extra triple for each molecule: <http://data.kasabi.com/dataset/chembl-rdf/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
+%NOTE: I have added extra triple for each molecule: <http://linkedchemistry.info/chembl/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
 %NOTE: Modified molecule compound structure can be queried here: https://wwwdev.ebi.ac.uk/chembl/sparql
@@
 \begin{verbatim}
-<http://data.kasabi.com/dataset/chembl-rdf/chemblid/ChEMBL324846>
+<http://linkedchemistry.info/chembl/chemblid/ChEMBL324846>
  skos:exactMatch <http://rdf.chemspider.com/370> .
@@
 will be resolved from the database using the full original URI,
-\url{http://data.kasabi.com/dataset/chembl-rdf/activity/a31863}. If the
+\url{http://linkedchemistry.info/chembl/activity/a31863}. If the
 user requested an RDF document using content negotiation, the original URIs will be unchanged,

### 2012-07-17 article.tex
Updated the ack and contrib sections for the recent activity

@@
 EW initiated the project, created the initial RDF version of the ChEMBL data, and encouraged the use cases.
-AW extended CitedIn to support citation info in ChEMBL-RDF.
+AW extended CitedIn to support citation info in ChEMBL-RDF and supported setting up the http://linkedchemistry.info/ host.
 OS, AW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
-JH supported the project with CHEMINF representations.
-PA integrated ChEMBL-RDF into Bio2RDF.
+JH supported the project with CHEMINF representations. PA integrated ChEMBL-RDF into Bio2RDF.
 JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
@@
 
-Tim Hodson and Zach Beauvais of Kasabi and A. L\"ovgren at the BMC Computing Department at Uppsala University for their
-support. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE.
-
-The data in ChEMBL is made available by funding from the Wellcome Trust [086151/Z/08/Z]
+The authors acknowledge A. L\"ovgren at the BMC Computing Department at Uppsala University for his
+support in hosting the SPARQL end point. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE.
+The data in ChEMBL is made available by funding from the Wellcome Trust [086151/Z/08/Z].
 

### 2012-07-17 article.tex
Updated the rest of the paper for the phase-out of Kasabi

@@
 querying is capped, based on the estimated computational effort. Second, resources have
-been made dereferencable using the Kasabi platform~\cite{kasabi}. The Kasabi hosting provides a SPARQL
-end point which requires a user account. A user account can be acquired for free for casual usage,
-or paid for uncapped  use.
+been made dereferencable via the \url{http://linkedchemistry.info} resource. The Kasabi
+platform~\cite{kasabi} was used for a period of time, until that project was discontinued.
+The Uppsala SPARQL end point still uses resource IRIs based on this, and follow a similar
+IRI pattern where linkedchemistry.info/chembl is replaced by data.kasabi.com/dataset/chembl-rdf.
 
@@
 OS, AW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
-JH supported the project with CHEMINF representations. PA integrated ChEMBL-RDF into Bio2RDF.
+JH supported the project with CHEMINF representations.
+PA integrated ChEMBL-RDF into Bio2RDF.
 JO, AG, and MD critically validated and summarized the ChEMBL data statistics.

### 2012-07-17 article.tex
Some fixes in the Refs section; still need a proper biblatex style for BioMed Central

@@
 \urlstyle{rm}
-\usepackage[citestyle=numeric-comp,sorting=none]{biblatex}
+\usepackage[citestyle=numeric-comp,sorting=none,firstinits=true,doi=false,isbn=false,url=false]{biblatex}
 \setlength{\bibitemsep}{0cm}                                                                                                                
@@
 based on descriptions of its biological properties. This process requires the effective 
-linkage of many scientific datasets~\cite{Samwald2011,OpenPHACTS}.
+linkage of many scientific datasets~\cite{Samwald2011,Williams2012}.
 

### 2012-07-17 article.tex
Added Bin's contribution

@@
 
-David/Bin: Chem2Bio2RDF
+Chem2Bio2RDF created a single RDF repository covering over 20 public data resources pertaining to drugs, chemical compounds, carcinogens,
+protein targets, genes, diseases, side effects, pathways and their relations~\cite{Chen2010}. The entities and their relations were
+further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative search~\cite{Chen2012a} and
+mining~\cite{Chen2012b}.  ChEMBL-RDF set (version 13) has been uploaded into Chem2Bio2RDF triple store, enabling to perform
+queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChi keys to present chemical
+compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link each other. For instance,
+to investigate the relations between drug side effects and their targets (usually off-targets), a query was created to link the
+targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
+In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets (IC$_50$ $<$ 10um). The top two most common targets
+of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
+the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
+The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and Sider in
+Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
+SPARQL endpoint~\cite{Chen2010}.
 

### 2012-07-18 article.tex
Reverted from biblatex to bibtex, as I did not quickly get a driver hacked up for BioMed Central
A bit of spelling checking

@@
 \urlstyle{rm}
-\usepackage[citestyle=numeric-comp,sorting=none,firstinits=true,doi=false,isbn=false,url=false]{biblatex}
-\setlength{\bibitemsep}{0cm}                                                                                                                
-\renewbibmacro{in:}{%                                                                                                                       
-  \ifentrytype{article}{}{%                                                                                                                 
-  \printtext{\bibstring{in}\intitlepunct}}}                                                                                                 
-\addbibresource{article.bib}
+%\usepackage[citestyle=numeric-comp,sorting=none,firstinits=true,doi=false,isbn=false,url=false]{biblatex}
+\usepackage{cite}
+%\setlength{\bibitemsep}{0cm}                                                                                                                
+%\renewbibmacro{in:}{%                                                                                                                       
+%  \ifentrytype{article}{}{%                                                                                                                 
+%  \printtext{\bibstring{in}\intitlepunct}}}                                                                                                 
+%\addbibresource{article.bib}
 
@@
 The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
-The SQL datadump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
+The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
 were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
@@
 using the complementary index and OpenMolecules RDF using InChI values. Protein links are given to Bio2RDF~\cite{Belleau2008}
-using the Uniprot identifier~\cite{TheUniProtConsortium2010}. Literature references are directed to CrossRef and Bio2RDF 
+using the UniProt identifier~\cite{TheUniProtConsortium2010}. Literature references are directed to CrossRef and Bio2RDF 
 using the DOIs and PubMed identifiers, respectively.
@@
 server~\cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases~\cite{Belleau2008}.
-A number of these databases are referenced in ChEMBL, including ChEBI, Pubmed, and both the
-Uniprot protein and taxonomy databases~\cite{TheUniProtConsortium2010}. These links are
+A number of these databases are referenced in ChEMBL, including ChEBI, PubMed, and both the
+UniProt protein and taxonomy databases~\cite{TheUniProtConsortium2010}. These links are
 vital to provide context for use cases that require a correlation between chemical
@@
 		\newline
-		\caption[wee]{Screnshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
-The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
+		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
+The top left canvas contains the query structure, in this case the drug carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
 	\label{fig:bioclipse-ds}
@@
 mining~\cite{Chen2012b}.  ChEMBL-RDF set (version 13) has been uploaded into Chem2Bio2RDF triple store, enabling to perform
-queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChi keys to present chemical
+queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChI keys to present chemical
 compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link each other. For instance,
@@
 the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
-The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and Sider in
+The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and SIDER in
 Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
@@
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
-\printbibliography
+ \bibliographystyle{bmc_article}  % Style BST file
+  \bibliography{article}
 }

### 2012-07-19 article.tex
Reordered the applications, and completed a full draft for the discussion

@@
 The use cases show that by using open standards and ontologies it is possible to link the ChEMBL database
-to other databases, such that the integration can understood by software. 
+to other databases. Using those links data integration can understood by software and create a foundation
+for integrative cheminformatics applications, such as the presented decision support.
 \end{abstract}
@@
 
+This first applications shows how ChEMBL-RDF can be integrated into other linked data services.
 The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
@@
 
+\subsection*{Chem2Bio2RDF}
+
+A second resource in the linked open data network in which ChEMBL-RDF has been integrated is Chem2Bio2RDF.
+Chem2Bio2RDF created a single RDF repository covering over 20 public data resources pertaining to drugs, chemical compounds, carcinogens,
+protein targets, genes, diseases, side effects, pathways and their relations~\cite{Chen2010}. The entities and their relations were
+further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative search~\cite{Chen2012a} and
+mining~\cite{Chen2012b}.  ChEMBL-RDF set (version 13) has been uploaded into Chem2Bio2RDF triple store, enabling to perform
+queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChI keys to present chemical
+compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link each other. For instance,
+to investigate the relations between drug side effects and their targets (usually off-targets), a query was created to link the
+targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
+In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets (IC$_50$ $<$ 10um). The top two most common targets
+of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
+the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
+The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and Sider in
+Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
+SPARQL endpoint~\cite{Chen2010}.
+
 \subsection*{CitedIn}
@@
 % Egon,Antony/Valery,Ola: develop and write up the Bioclipse Decision Support use case
+The use of the RDF open standard also simplifies the integration of data resources with 
+scientific software, as we have shown previously~\cite{Willighagen2011,Willighagen:2011kx}.
+We here demonstrate a new application following this idea.
 
@@
 		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
-The top left canvas contains the query structure, in this case the drug carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
+The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
 	\label{fig:bioclipse-ds}
@@
 \end{figure*}
-
-\subsection*{Chem2Bio2RDF}
-
-Chem2Bio2RDF created a single RDF repository covering over 20 public data resources pertaining to drugs, chemical compounds, carcinogens,
-protein targets, genes, diseases, side effects, pathways and their relations~\cite{Chen2010}. The entities and their relations were
-further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative search~\cite{Chen2012a} and
-mining~\cite{Chen2012b}.  ChEMBL-RDF set (version 13) has been uploaded into Chem2Bio2RDF triple store, enabling to perform
-queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChI keys to present chemical
-compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link each other. For instance,
-to investigate the relations between drug side effects and their targets (usually off-targets), a query was created to link the
-targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
-In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets (IC$_50$ $<$ 10um). The top two most common targets
-of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
-the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
-The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and SIDER in
-Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
-SPARQL endpoint~\cite{Chen2010}.
 
@@
 
-While we show here that the RDF version of the ChEMBL data is very useful, the current RDF triples are by no means static: ChEMBL-RDF
-will keep evolve, following new developments in the Linked Open Data world. For example, it is likely that over time we will link out
-to more Linked Data resources. More importantly, we wish to adopt more common ontologies, of which the BioAssay Ontology is planned
-to be the next~\cite{Visser2011}. These steps will make the ChEMBL-RDF triples even more interoperable. Part of this future work
-is already being performed within the context of the Open PHACTS project~\cite{Williams2012}.
+This papers shows the evolution of the ChEMBL-RDF resource in standardizing access to the
+ChEMBL data in a machine-readable manner, by using semantic web technologies. This standardization
+applies to several aspects of data exchange, including the data serialization format and
+access options, the data structure, and the ontologies used to explain what the data is about.
+We touched upon choices made to implement these approaches, such as a selection of
+RDF technologies, and chosen ontologies.
+
+For example, for linking to other databases we selected a few predicates to link to
+other databases: rdfs:seeAlso, skos:exactMatch, owl:equivalentClass, owl:sameAs. Each
+has an increasing level of implications, ranging from "you may also be interested in" to
+"both objects are identical and all properties apply to both objects". Additionally,
+the CiTO ontology has been used to link data to the literature it was extracted from,
+which provides link that lies in between rdfs:seeAlso and skos:exactMatch.
+
+Another ontology that we adopted in this ChEMBL-RDF 13 is the cheminformatics ontology
+CHEMINF, allowing us to encode numeric and non-numeric descriptors of chemical entities
+in the same standardized way. Using this approach we expose identity descriptors such
+as the SMILES string, InChI, and InChIKey, and molecular properties including the
+logP, rotatable bond count, and hydrogen bond donor and acceptor counts.
+
+However, not all information is yet using standardized ontologies. For example, the
+activity types (IC50, EC50, Potency) are not exposed using an ontology, making it
+still impossible to reason or just group over similar activity types. Likewise,
+measurement units are provided as literals, making it needlessly hard to compare
+concentrations in nM and mM. For this kind of data, two new ontologies were
+recently developed. For example, the BioAssay Ontology addresses many of our
+remaining concerns around making the experimental data machine readable~\cite{Visser2011},
+while the Quantity, Unit, Dimension and Type (QUDT) ontology makes it possible
+to express measurements in a machine readable way~\cite{QUDT}. Both approaches are
+currently explored to be applied to ChEMBL-RDF as part of work in the
+Open PHACTS project~\cite{Williams2012}.
+
+Finally, we stress that nothing in this paper cannot be done with other approaches, such as
+more common relational databases. In fact, that is exactly what the ChEMBL website does.
+Similarly, there is nothing principle component analysis does that you cannot do with drawings
+on paper too; it is just silly. We have here presented a linked data approach to the ChEMBL data
+showing standardized and simplified integration to other resources, and that is what
+cheminformatics is about.
 

### 2012-07-19 article.tex
I'm the corresponding author

@@
 
-\author{Egon~L.~Willighagen$^{1}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl}\and
+\author{Egon~L.~Willighagen$^{1,*}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl}\and
 Andra~Waagmeester$^1$\email{Andra Waagmeester - FIXME}\and

### 2012-07-19 article.tex
Added email addresses

--- 2012-07-19 article.tex
@@
 
-\author{Egon~L.~Willighagen$^{1}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl}\and
-Andra~Waagmeester$^1$\email{Andra Waagmeester - FIXME}\and
-Ola Spjuth$^2$\email{Ola Spjuth - FIXME}\and
+\author{Egon~L.~Willighagen$^{1,*}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl}\and
+Andra~Waagmeester$^1$\email{Andra Waagmeester - andra.waagmeester@maastrichtuniversity.nl}\and
+Ola Spjuth$^2$\email{Ola Spjuth - ola.spjuth@farmbio.uu.se}\and
 Peter~Ansell$^3$\email{Peter Ansell - p\_ansell@yahoo.com}\and
-Antony~Williams$^4$\email{Antony Williams - FIXME}\and
-Valery~Tkachenko$^4$\email{Valery Tkachenko - FIXME}\and
-Janna~Hastings$^5$\email{Janna Hastings - FIXME}\and
-John~Overington$^6$\email{John Overington - FIXME}\and
-Anna~Gaulton$^6$\email{Anna Gaulton - FIXME}\and
-Mark~Davies$^6$\email{Mark Davies - FIXME}\and
-Bin~Chen$^7$\email{Bin Chen - FIXME}\and
-David~Wild$^7$\email{David Wild - FIXME}
+Antony~J.~Williams$^4$\email{Antony Williams - WilliamsA@rsc.org}\and
+Valery~Tkachenko$^4$\email{Valery Tkachenko - TkachenkoV@rsc.org}\and
+Janna~Hastings$^5$\email{Janna Hastings - hastings@ebi.ac.uk}\and
+John~Overington$^6$\email{John Overington - jpo@ebi.ac.uk}\and
+Anna~Gaulton$^6$\email{Anna Gaulton - agaulton@ebi.ac.uk}\and
+Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk}\and
+Bin~Chen$^7$\email{Bin Chen - binchen@indiana.edu}\and
+David~Wild$^7$\email{David Wild - djwild@indiana.edu}
 }

### 2012-07-20 article.tex
CitedIn additions

@@
 \subsection*{CitedIn}
+The impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this narrowly defined calculation increasingly fails to capture the true reach of a paper.
+CitedIn (http://www.citedin.org), an online resource was developed to make scientific impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public resources.
+ChEMBL contains many literature references.
+Currently ChEMBL's SPARQL endpoint is queried by CitedIn and thee results are presented to its users.
+(see Figure xx).
+Counting citations in ChEMBL is done by the following SPARQL query.
 
-Andra: use the SPARQL end point to find which entries in ChEMBL cite what papers
-
-The link between data and formal publication is important in many areas of
-attribution, scientist ranking, etc, as outline in \cite{Waagmeester2012}.
-ChEMBL contains many literature references, and we wish to query this data
-for CitedIn.
+\begin{small}
+\begin{verbatim}
+SELECT * WHERE {
+                  ?paper <http://purl.org/ontology/bibo/pmid> \"$pmid\" .
+                  ?paper <http://purl.org/ontology/bibo/pmid> ?pmid .
+                  ?object ?cites ?paper .
+}
+\end{verbatim}
+\end{small}
 

### 2012-07-21 article.tex
Updated with a working SPARQL

@@
 
-The following SPARQL query identifies a set of molecules, which based on data curated from the 
+The following SPARQL query identifies a set of 65 molecules, which based on data curated from the 
 literature and stored in the ChEMBL-RDF data model, selectively bind Human Cyclin-Dependent Kinase 2 
@@
 
-NOTE: Query should return 65 molecules
-
 \begin{small}
 \begin{verbatim}
-select distinct ?molecule WHERE 
-{ 
-  ?assay ?p chembl:Assay ;
-  chembl:hasTarget ?target .
-  ?activity ?p chembl:Activity ;
+select distinct ?molecule WHERE {
+  ?target1 owl:sameAs <http://bio2rdf.org/uniprot:P24941> .
+  ?assay1 chembl:hasTarget ?target1 .
+  ?activity1 chembl:onAssay ?assay1 ;
     chembl:forMolecule ?molecule ;
-  chembl:onAssay ?assay ;
-  chembl:type ?type ;
-  chembl:standardUnits ?unit ;
-  chembl:standardValue ?value;
-  chembl:relation ?relation
-  FILTER (
-    ?type = "IC50"  &&
-    ?unit = "nM"    &&
-    ?value  < 50    &&
-    ?relation = "="
-  ) .
-  ?target ?p chembl:Target;
-  dc:identifier ?id
-  FILTER (
-    ?id = "uniprot:P24941"
-  ) .
-  { 
-    select ?molecule WHERE 
-    {
-      ?assay ?p chembl:Assay ;
-      chembl:hasTarget ?target .
-      ?activity ?p chembl:Activity ;
+    chembl:type "IC50" ;
+    chembl:standardUnits "nM" ;
+    chembl:standardValue ?value1 ;
+    chembl:relation "="
+    FILTER ( ?value1 < 50 ) .
+  ?target2 owl:sameAs <http://bio2rdf.org/uniprot:P11802> .
+  ?assay2 chembl:hasTarget ?target2 .
+  ?activity2 chembl:onAssay ?assay2 ;
     chembl:forMolecule ?molecule ;
-      chembl:onAssay ?assay;
-      chembl:type ?type ;
-      chembl:standardUnits ?unit ;
-      chembl:standardValue ?value ;
-      chembl:relation ?relation
-      FILTER (
-        ?type = "IC50"  &&
-        ?unit = "nM"    &&
-        ?value  > 200   &&
-        ?relation = "="
-      ) .
-      ?target ?p chembl:Target;
-      dc:identifier ?id
-      FILTER (
-        ?id = "uniprot:P11802"
-      ) .
-    }
-  }
+    chembl:type "IC50" ;
+    chembl:standardUnits "nM" ;
+    chembl:standardValue ?value2 ;
+    chembl:relation "="
+    FILTER ( ?value2 > 200 ) .
 }

### 2012-07-21 article.tex
Added the CitedIn screenshot, and updated the text a bit accordingly

@@
 
+\begin{figure*}[!ht]
+		\begin{center}
+		\includegraphics[width=15cm]{cited.png}
+		\newline
+		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
+The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
+	\label{fig:bioclipse-ds}
+		\end{center}
+\end{figure*}
+
 \subsection*{CitedIn}
+
+Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as in CitedIn, (http://www.citedin.org).
 The impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this narrowly defined calculation increasingly fails to capture the true reach of a paper.
-CitedIn (http://www.citedin.org), an online resource was developed to make scientific impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public resources.
+CitedIn was developed to make scientific impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public resources.
 ChEMBL contains many literature references.
-Currently ChEMBL's SPARQL endpoint is queried by CitedIn and thee results are presented to its users.
-(see Figure xx).
+The SPARQL endpoint is queried by CitedIn and the results are presented to its users.
+Figure shows the result for the Nature paper by Gamo presenting the malari structures, which is highly cited in ChEMBL~\cite{Gamo2010}.
 Counting citations in ChEMBL is done by the following SPARQL query.
@@
 \begin{verbatim}
-SELECT * WHERE {
-                  ?paper <http://purl.org/ontology/bibo/pmid> \"$pmid\" .
-                  ?paper <http://purl.org/ontology/bibo/pmid> ?pmid .
-                  ?object ?cites ?paper .
+SELECT count(?object) WHERE {
+  ?paper bibo:pmid \"$pmid\" .
+  ?object cito:citesAsDataSource ?paper .
 }

### 2012-07-26 article.tex
Fixed the caption and png location
Addressed the comments from Antony Williams

@@
 to other databases. Using those links data integration can understood by software and create a foundation
-for integrative cheminformatics applications, such as the presented decision support.
+for integrated cheminformatics applications, such as the presented decision support.
 \end{abstract}
@@
 provides several new challenges for information systems. Scientists wish to 
-discover new, unique, and significant patterns in datasets, which explaining biological 
+discover new, unique, and significant patterns in datasets that explain biological 
 phenomena not yet understood. This process requires integration of datasets that are both 
-sparse, and growing independent of one another. The discovery of new patterns showing the 
+sparse, and growing independently of one another. The discovery of new patterns showing the 
 causes and effects of various biological phenomena is beyond the scope of single datasets.
@@
 Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}. RDF is a framework where
-knowledge is represented by small, so-called triples reflecting very basic facts. Things
-in RDF are represented by Universal Resource Identifiers (URI) and when the more specialized
+knowledge is represented by small so-called triples reflecting very basic facts. Any fact can
+be given on anything, and RDF therefore talks about \textit{things}~\cite{Miller:04:RP}. Things in RDF
+are represented by Universal Resource Identifiers (URI) and when the more specialized
 Universal Resource Locator (URL) is used, it creates \emph{linked data}~\cite{Samwald2011}. RDF is
@@
 This paper presents an update on the ChEMBL-RDF dataset, and details of the latest structures
-and ontologies used to map ChEMBL 13 to RDF, along with new use cases, showing 
+and ontologies used to map ChEMBL to RDF, along with new use cases, showing 
 links to other datasets using their RDF URIs to further support research in the life sciences.
@@
 has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
-Group~\cite{Marshall2012}. This process has not changed much.
+Group~\cite{Marshall2012}. This process has not changed much since its original description.
 
-The previous ChEMBL-RDF conversion used, however, custom predicates and classes to represent RDF triples, defining
+However, The previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
 an ad-hoc ontology~\cite{Willighagen2011}. The current version, instead, uses community-proposed
@@
 references and citations, and domain ontologies like the Protein Ontology\cite{Sidhu2006} and the Chemical Information
-Ontology\cite{Hastings2011}. Throughout this paper, various prefixes are used to denote different ontology 
+Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
 namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.
@@
 
-To expose that ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
+To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
 hosted at Uppsala University, using the Open Source Virtuoso software. Use is free, but the
@@
 
-We here present the updated ChEMBL-RDF, including a description of the mapping process 
+We present here the updated ChEMBL-RDF, including a description of the mapping process 
 and the identification of links to other datasets.
@@
 provided as labels. When InChI and InChIKeys are available for a drug, then these are
-provided via the CHEMINF formalism (here abbreviated):
+provided via the CHEMINF formalism (abbreviated here):
 
@@
 these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
-CHEMINF ontology approach. Here is, for example, the ALogP value:
+CHEMINF ontology approach. Here is, for example, the ALogP value (as distributed with ChEMBL and
+calculated with ACD/Labs software):
 
@@
 The literature from which all this data was extracted, is replicated from the database,
-but only a subset of information about that literature is provided. In all cases,
+but only a subset of information about that literature is provided. In all cases
 it uses the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) to link
@@
 
-Each release of the ChEMBL database is accompanied by set of release counts, which help 
+Each release of the ChEMBL database is accompanied by a set of release counts, which help 
 provide a concise content overview of the release in question. The ChEMBL 13 release counts
@@
 The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 less than appears 
-in ChEMBL 13 release notes, as it does not include doc\_id -1, which refers to the 
+in ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as it does not include doc\_id -1, which refers to the 
 unpublished dataset.
@@
 Recreation of the ChEMBL release statistics, by querying the ChEMBL-RDF with SPARQL 
-provides end users with high degree of confidence that the ChEMBL-RDF contains same 
-core content as source relational database.
+provides end users with a high degree of confidence that the ChEMBL-RDF contains the same 
+core content as the source relational database.
 
@@
 ChemSpider~\cite{Pence2010} is a freely accessible, online database provided by Royal Society of Chemistry (RSC).
-It contains over 26 million unique
+It contains over 28 million unique
 chemical compounds aggregated from over 400 data sources as well as chemical data extracted
@@
 Notice here the use of owl:equivalentClass to match the formalism in CHEMINF that defines molecules
-are classes, rather than instances.
+as classes, rather than instances.
 
@@
 
-The links services enables the ChEMBL application to derive both forward links, originating in
+These links service enables the ChEMBL application to derive both forward links, originating in
 ChEMBL, e.g. \url{http://localhost:8080/chembl/linkstonamespace/-targetns/originalns:identifier},
-and backward links, originating in other databases, such as LODD, Bio2RDF,
-\footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, and Chem2Bio2RDF.
+and backward links, originating in other databases, such as LODD, Bio2RDF
+(\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}), and Chem2Bio2RDF.
 These services are vital to efficiently navigate the Linked Data web, as it is both
@@
 further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative search~\cite{Chen2012a} and
-mining~\cite{Chen2012b}.  ChEMBL-RDF set (version 13) has been uploaded into Chem2Bio2RDF triple store, enabling to perform
+mining~\cite{Chen2012b}. The ChEMBL-RDF set has been uploaded into the Chem2Bio2RDF triple store, enabling
 queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChI keys to present chemical
@@
 the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
-The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and Sider in
+The targets can be also linked to side effects via protein-protein interactions by searching across ChEMBL, HPRD and Sider in
 Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
@@
 		\begin{center}
-		\includegraphics[width=15cm]{cited.png}
+		\includegraphics[width=15cm]{citedin.png}
 		\newline
-		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
-The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
-	\label{fig:bioclipse-ds}
+		\caption[wee]{Screenshot of the CitedIn web service showing a paper cited 94671 times in the ChEMBL database.}
+	\label{fig:citedin}
 		\end{center}
@@
 The SPARQL endpoint is queried by CitedIn and the results are presented to its users.
-Figure shows the result for the Nature paper by Gamo presenting the malari structures, which is highly cited in ChEMBL~\cite{Gamo2010}.
+Figure~\ref{fig:citedin} shows the result for the Nature paper by Gamo presenting the malaria structures, which is highly cited in ChEMBL~\cite{Gamo2010}.
 Counting citations in ChEMBL is done by the following SPARQL query.
@@
 
-NOTE: May want to exclude this Use Case as query has not successful run yet, keeps causing virtuoso 
-to timeout or run out of memory. Currently playing with local server setup to get query to run.
-
 Designing a molecule, which is selective to one target over another will often be considered a 
@@
 to understand that as sequence identity amongst residues, which contribute to potential small 
-molecule binding sites remains high so does the issue of selectivity. That is, the molecule being 
+molecule binding sites remains high, so does the issue of selectivity. That is, the molecule being 
 designed may also bind to the equivalent binding site in the  closely related protein 
@@
 targets using different activity types recorded in the literature. Using activity types, 
-which act as a measure of binding affinity, such as IC$_{50}$ or K$_{i}$ and applying activity value cuts
+which act as a measure of binding affinity, such as $IC_{50}$ or $K_{i}$ and applying activity value cuts
 offs it is possible to identify molecules which have higher binding affinity to certain targets 
-compared to others. Taking this one step further it is possible to identify a set molecules, 
-which have a high affinity to protein A (e.g. IC$_{50}$ value $<$ 50~nM) and low affinity to protein 
-B (e.g. IC$_{50}$ value $>$ 200~nM).
+compared to others. Taking this one step further it is possible to identify a set molecules
+which have a high affinity to protein A (e.g. $IC_{50}$ value $<$ 50~nM) and low affinity to protein 
+B (e.g. $IC_{50}$ value $>$ 200~nM).
 
@@
 
-This papers shows the evolution of the ChEMBL-RDF resource in standardizing access to the
+This paper shows the evolution of the ChEMBL-RDF resource in standardizing access to the
 ChEMBL data in a machine-readable manner, by using semantic web technologies. This standardization
@@
 the CiTO ontology has been used to link data to the literature it was extracted from,
-which provides link that lies in between rdfs:seeAlso and skos:exactMatch.
+which provides links that lie in between rdfs:seeAlso and skos:exactMatch.
 
@@
 Similarly, there is nothing principle component analysis does that you cannot do with drawings
-on paper too; it is just silly. We have here presented a linked data approach to the ChEMBL data
+on paper too; it is just inappropriate to do so when computer-based approaches are so much
+more efficient. We have presented here a linked data approach to the ChEMBL data
 showing standardized and simplified integration to other resources, and that is what
@@
 AW extended CitedIn to support citation info in ChEMBL-RDF and supported setting up the http://linkedchemistry.info/ host.
-OS, AW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
+OS, AJW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
 JH supported the project with CHEMINF representations.

### 2012-07-26 article.tex
Addressed some of Bin's comments, while others are discussed for a larger audience via the mailing list

@@
 The use cases show that by using open standards and ontologies it is possible to link the ChEMBL database
-to other databases. Using those links data integration can understood by software and create a foundation
+to other databases. Using those links data integration can be understood by software and create a foundation
 for integrated cheminformatics applications, such as the presented decision support.
@@
 targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
-In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets (IC$_50$ $<$ 10um). The top two most common targets
+In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets ($IC_50$ $<$ 10um). The top two most common targets
 of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
@@
 OS, AJW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
-JH supported the project with CHEMINF representations.
-PA integrated ChEMBL-RDF into Bio2RDF.
+JH supported the project with CHEMINF representations. PA integrated ChEMBL-RDF into Bio2RDF.
 JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
+BC, DW contributed the Chem2Bio2RDF use cases.
 All authors contributed to the continued development of ChEMBL-RDF, as well as to the writing of the paper,

### 2012-07-26 article.tex
Moved Mark earlier into the author list

@@
 Janna~Hastings$^5$\email{Janna Hastings - hastings@ebi.ac.uk}\and
+Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk}\and
 John~Overington$^6$\email{John Overington - jpo@ebi.ac.uk}\and
 Anna~Gaulton$^6$\email{Anna Gaulton - agaulton@ebi.ac.uk}\and
-Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk}\and
 Bin~Chen$^7$\email{Bin Chen - binchen@indiana.edu}\and

### 2012-08-01 article.tex
Janna's overall edits

@@
 \iid(4)Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A. \\
-\iid(5)Chemoinformatics and Metabolism, European Bioinformatics Institute, POSTAL CODE, Hinxton, United Kingdom \\
-\iid(6)EMBL-European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\ 
+\iid(5)Cheminformatics and Metabolism, European Bioinformatics Institute, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\
+\iid(6)Chemogenomics, European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\ 
 \iid(7)School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.
@@
 RDF documents can natively link to related data, and others can link back using Universal Resource Identifiers (URIs).
-RDF makes the data machine-readable, using extensible vocabularies for additional information, making it easier
+RDF makes the data machine-readable and uses extensible vocabularies for additional information, making it easier
 to scale up inference and data analysis.
 \paragraph*{Results:}
-This paper describes the continued project to convert data from the ChEMBL database into RDF triples.
-This updated version of ChEMBL-RDF now uses recently introduced ontologies, including CHEMINF and CiTO,
+This paper describes recent developments in an ongoing project converting data from the ChEMBL database into RDF triples.
+Relative to earlier versions, this updated version of ChEMBL-RDF uses recently introduced ontologies, including CHEMINF and CiTO,
 exposes more information from the database, and is now available as dereferencable, linked data.
@@
 other web resources, including Bio2RDF, Chem2Bio2RDF, and ChemSpider, and showing the use of standard
-ontologies.
+ontologies for querying.
 \paragraph*{Conclusions:}
-The use cases show that by using open standards and ontologies it is possible to link the ChEMBL database
-to other databases. Using those links data integration can be understood by software and create a foundation
-for integrated cheminformatics applications, such as the presented decision support.
+We have illustrated the advantage of using open standards and ontologies to link the ChEMBL database
+to other databases. Using those links and the knowledge encoded in standards and ontologies, the ChEMBL-RDF
+resource creates a foundation for integrated semantic web cheminformatics applications, 
+such as the presented decision support.
 \end{abstract}
@@
 
-The current scientific data deluge, in which datasets grow faster in size and complexity than scientists can handle,
+The current scientific data deluge, in which datasets grow faster in size and complexity than scientists can keep up with,
 provides several new challenges for information systems. Scientists wish to 
@@
 sparse, and growing independently of one another. The discovery of new patterns showing the 
-causes and effects of various biological phenomena is beyond the scope of single datasets.
+causes and effects of various biological phenomena is often beyond the scope of single datasets.
 For example, systems biologists integrate micro-array differential expression datasets to 
 biological pathways, using various other datasets to provide evidence for the links~\cite{Staab2007}. 
-Another prominent example is drug discovery where a new unique chemical entity is discovered 
-based on descriptions of its biological properties. This process requires the effective 
+Another prominent example is drug discovery, in which a new unique chemical entity is designed or discovered 
+based on descriptions of its required biological properties. This process requires the effective 
 linkage of many scientific datasets~\cite{Samwald2011,Williams2012}.
@@
 ChEMBL contains descriptions for biological activities involving over a million chemical 
-entities, extracted from literature. This provides a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
+entities, extracted from the scientific literature, providing a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
 It is updated on a fairly frequent basis as the existing data is further curated and new data is added. 
@@
 The former requires scientists to import the data into a relational database, while the 
-latter limits the machine access to the data.
+latter limits the machine access to the data. ChEMBL also provides web services for programmatic
+access, but these are limited to specific types of query as encoded in the web service API.
 
@@
 Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}. RDF is a framework where
-knowledge is represented by small so-called triples reflecting very basic facts. Any fact can
-be given on anything, and RDF therefore talks about \textit{things}~\cite{Miller:04:RP}. Things in RDF
+knowledge is represented by small so-called triples, reflecting ``facts''. A fact can
+be given on any topic, and RDF therefore talks about \textit{things} very generically~\cite{Miller:04:RP}. Things in RDF
 are represented by Universal Resource Identifiers (URI) and when the more specialized
@@
 developed as an open standard by the World Wide Web Consortium, and provides a knowledge
-representation framework that applies to many, if not any, domain, making it a perfect
-approach for data integration in the life sciences.
+representation framework that applies to many, if not all, domains, making it a successful 
+approach to data integration in the life sciences.
 
-This paper presents an update on the ChEMBL-RDF dataset, and details of the latest structures
+This paper presents an update on the ChEMBL-RDF resource, including details of the latest structures
 and ontologies used to map ChEMBL to RDF, along with new use cases, showing 
@@
 The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
-were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
+were then executed to generate RDF triples using a set of PHP scripts. These scripts are Open Source 
 and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
@@
 
-However, The previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
+However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
 an ad-hoc ontology~\cite{Willighagen2011}. The current version, instead, uses community-proposed
-ontologies, making the RDF more interoperable. The new triple dataset uses standard ontologies
-including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology~\cite{Hastings2011} for literature
-references and citations, and domain ontologies like the Protein Ontology\cite{Sidhu2006} and the Chemical Information
+ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
+The new triple dataset uses standard ontologies
+including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology~\cite{Shotton2010} for literature
+references and citations, and domain ontologies such the Protein Ontology\cite{Sidhu2006} and the Chemical Information
 Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
@@
 
-To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
-hosted at Uppsala University, using the Open Source Virtuoso software. Use is free, but the
-querying is capped, based on the estimated computational effort. Second, resources have
+To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point is being
+hosted at Uppsala University, using the Open Source Virtuoso software. Usage is free, but the
+allowed volume of querying is capped, based on the estimated computational effort. Second, resources have
 been made dereferencable via the \url{http://linkedchemistry.info} resource. The Kasabi
 platform~\cite{kasabi} was used for a period of time, until that project was discontinued.
-The Uppsala SPARQL end point still uses resource IRIs based on this, and follow a similar
+The Uppsala SPARQL end point still uses resource IRIs based on the Kasabi approach, following a similar
 IRI pattern where linkedchemistry.info/chembl is replaced by data.kasabi.com/dataset/chembl-rdf.
@@
 Although many parts of the ChEMBL-RDF 13 dataset now use community based ontologies, some 
-terms from the previous ad-hoc ontology are still used. These terms were created under the 
+terms from the previous ad-hoc ontology are still used because no community alternative is yet available. 
+These terms were created under the 
 URI namespace \textit{http://rdf.farmbio.uu.se/chembl/onto/\#}, and are referenced here using the
@@
 
-The core concept in the ChEMBL database is that of the biological activity. This
-is the type of information that the ChEMBL database extracts from literature.
-Standard ontologies are being explored for future release for annotation
-of activities, but, currently, the
-type, fields and links to other resources are still using the \textit{chembl}
-namespace. While ChEMBL provides both the original activities as found in literature
+The core concept in the ChEMBL database is that of the biological activity. 
+While ChEMBL provides both the original activities as found in literature
 and standardized values allowing comparison between studies, the triples only
-make the latter available. The link to literature is encoded using the CiTO ontology.
+make the latter available. The links to the literature are encoded using the CiTO ontology.
 This results in a set of triples that look like:
@@
 and Ki (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
-using, an ontology.
+linked to, an ontology.
 
 The activities themselves are measured against assays, which make up a second important
-resource type. Various assay types are found in the database: chembl:ADMET, chembl:Binding,
-chembl:Functional, chembl:Property, and chembl:Unassigned.
+resource type. Various assay types are found in the database: \verb+chembl:ADMET+, \verb+chembl:Binding+,
+\verb+chembl:Functional+, \verb+chembl:Property+, and \verb+chembl:Unassigned+.
 
@@
 The ChEMBL database
-recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
-chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
-chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
-explicit types, and effectively uses a closed-world approach. A target
-specification can then look like:
+recognizes various types: \verb+pro:PR_0000001+, \verb+chembl:ADMET+, \verb+chembl:CELL-LINE+,
+\verb+chembl:NUCLEIC-ACID+, \verb+chembl:ORGANISM+, \verb+chembl:SUBCELLULAR+, \verb+chembl:TISSUE+,
+\verb+chembl:UNCHECKED+, and \verb+chembl:UNKNOWN+. The latter two are currently defined as
+explicit types, thus effectively implementing a closed-world approach.
+
+A target specification can then look like:
 
@@
 classes defined by several ontologies for these drug types. For example, proteins subclass the
-Protein concept from the PRotein Ontology (PR\_000000001), small
-molecules subclass the Chemical Entity concept from the CHEMINF
-ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
+Protein class from the PRotein Ontology (PR\_000000001), small
+molecules subclass the Chemical Entity class from the CHEMINF
+ontology (CHEMINF\_000000) (which is itself equivalent to the Chemical Entity class in ChEBI, CHEBI\_24431), 
+and oligosaccharides and oligonucleotides
 subclass their respective matches in the CHEBI ontology (CHEBI\_50699
@@
 
-It should also be noted that the drugs are not subclassed from
-a general drug class, or any other common superclass. However, to annotate
-the drugs to be drugs, they are identified as having the 
-the role of being a drug. This is triplified using the OBO and ChEBI
-ontologies, where CHEBI\_23888 is the drug role:
+It should also be noted that the relationship of these structural classes of drug 
+to a general drug class is that of having a \textit{role} rather than a subclass relationship. 
+This is triplified using the OBO and ChEBI ontologies, where CHEBI\_23888 is the drug role:
 
@@
 these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
-CHEMINF ontology approach. Here is, for example, the ALogP value (as distributed with ChEMBL and
+CHEMINF ontology. Here is, for example, the ALogP value (as distributed with ChEMBL and
 calculated with ACD/Labs software):
@@
 The literature from which all this data was extracted, is replicated from the database,
-but only a subset of information about that literature is provided. In all cases
-it uses the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) to link
-to authoritative databases to provide full detail, as outlined later in this paper.
-The provided properties are using the BIBO ontology:
+but only a subset of the possible information about that literature is provided. In all cases 
+the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) are used to link
+to authoritative databases that provide full detail about the references, as outlined later in this paper.
+The properties are provided using the BIBO ontology:
 
@@
 
-The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 less than appears 
-in ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as it does not include doc\_id -1, which refers to the 
-unpublished dataset.
+The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 fewer than appears 
+in the ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as the result does not include the doc\_id -1, as this refers to the 
+unpublished dataset, not itself a valid \verb+bibo:Article+.
 
@@
 
-Recreation of the ChEMBL release statistics, by querying the ChEMBL-RDF with SPARQL 
+Reproduction of the ChEMBL release statistics through querying ChEMBL-RDF with SPARQL 
 provides end users with a high degree of confidence that the ChEMBL-RDF contains the same 
@@
 It contains over 28 million unique
-chemical compounds aggregated from over 400 data sources as well as chemical data extracted
+chemical compounds aggregated from over 400 data sources, as well as chemical data extracted
 from RSC scientific articles and databases. Since its inception, efforts have been made to
@@
 
-Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
+Link mappings are provided with \verb+skos:exactMatch+ predicates, while the ChemSpider identifiers
 are also available via a CHEMINF representation:
@@
 to define unique IRIs for molecules~\cite{Bradley2009,Willighagen2011}. While IRIs are theoretically unlimited in length,
-in practice web browsers and servers limit the length of IRIs. Virtuoso is, unfortunately,
-a system which supports only IRIs of up to a certain length. Therefore, InChI-based links are only created for smaller molecules. 
+in practice many web browsers and servers limit the length of IRIs. Virtuoso is, unfortunately,
+one such system that supports only IRIs of up to a certain length. Therefore, InChI-based links are only created for smaller molecules. 
 Almost 1.3 million links to OpenMolecules RDF were created in a similar manner to:
@@
 
-Notice here the use of owl:equivalentClass to match the formalism in CHEMINF that defines molecules
-as classes, rather than instances.
+Notice here the use of \verb+owl:equivalentClass+ since in CHEMINF molecules are defined as
+as classes rather than instances.
 
@@
 ChEMBL-RDF can be used to explore interrelated scientific datasets, and we here present a few applications. 
-The first application describes how it uses the SPARQL end point can be used to make the data available as linked data
+The first application describes how the SPARQL end point can be used to make the data available as linked data
 via the Bio2RDF platform. The second application takes advantage of the bibliographic information
@@
 
-This first applications shows how ChEMBL-RDF can be integrated into other linked data services.
+This first application shows how ChEMBL-RDF can be integrated into other linked data services.
 The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
-based services for standard URI resolution, along with text and link searches~\cite{Ansell2011}. 
+based services for standard URI resolution, along with textual and link-based searches~\cite{Ansell2011}. 
 A Java Web Archive (WAR) file along with the configuration files and build scripts for the ChEMBL-RDF Linked Data server are available on GitHub~\cite{WebAppGitHub}. 
 It proxies the standard ChEMBL-RDF URIs by translating URLs between those requested by users and the URIs that are available
-in SPARQL endpoints. For example, if the ChEMBL web application is running on the users
+in SPARQL endpoints. For example, if the ChEMBL web application is running on the user's
 local machine, e.g. \url{http://localhost:8080/chembl/}, then a request for the
@@
 however, if the user requested an HTML document, the results will contain both the
-original RDF triples, represented using RDFa, with links that resolve using the users local machine.
+original RDF triples, represented using RDFa, with links that resolve using the user's local machine.
 
-These links service enables the ChEMBL application to derive both forward links, originating in
+These link services enable the ChEMBL application to derive both forward links, originating in
 ChEMBL, e.g. \url{http://localhost:8080/chembl/linkstonamespace/-targetns/originalns:identifier},
@@
 
-A second resource in the linked open data network in which ChEMBL-RDF has been integrated is Chem2Bio2RDF.
+A second resource in the linked open data network into which ChEMBL-RDF has been integrated is Chem2Bio2RDF.
 Chem2Bio2RDF created a single RDF repository covering over 20 public data resources pertaining to drugs, chemical compounds, carcinogens,
 protein targets, genes, diseases, side effects, pathways and their relations~\cite{Chen2010}. The entities and their relations were
-further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative search~\cite{Chen2012a} and
-mining~\cite{Chen2012b}. The ChEMBL-RDF set has been uploaded into the Chem2Bio2RDF triple store, enabling
-queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChI keys to present chemical
-compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link each other. For instance,
-to investigate the relations between drug side effects and their targets (usually off-targets), a query was created to link the
+further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative searches~\cite{Chen2012a} and
+data mining~\cite{Chen2012b}. The ChEMBL-RDF set has been uploaded into the Chem2Bio2RDF triple store, enabling
+queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL-RDF and Chem2Bio2RDF use InChI keys to present chemical
+compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link the two. For instance,
+to investigate the relations between drug side effects and their targets (usually off-target effects), a query was created to link the
 targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
-In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets ($IC_50$ $<$ 10um). The top two most common targets
+In this case, 36 drugs causing heart disease were linked to 87 unique protein targets ($IC_50$ $<$ 10um). The top two most common targets
 of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
@@
 
-Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as in CitedIn, (http://www.citedin.org).
-The impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this narrowly defined calculation increasingly fails to capture the true reach of a paper.
+Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as CitedIn (http://www.citedin.org).
+While the impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this narrowly defined calculation increasingly fails to capture the true reach of a paper.
 CitedIn was developed to make scientific impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public resources.
@@
 Figure~\ref{fig:citedin} shows the result for the Nature paper by Gamo presenting the malaria structures, which is highly cited in ChEMBL~\cite{Gamo2010}.
-Counting citations in ChEMBL is done by the following SPARQL query.
+Counting citations in ChEMBL is done by the following SPARQL query:
 
@@
 This enables users browsing chemical structures in Bioclipse to look up near neighbors in ChemSpider 
-via the ChemSpider Web API (SOAP), and for the found compounds to query ChEMBL-RDF for 
+via the ChemSpider Web API (SOAP), and for the found compounds to be used to query ChEMBL-RDF for 
 available interaction data. The results are presented alongside predictive models in Bioclipse
 (see Fig~\ref{fig:bioclipse-ds}), and can be used for decision support when evaluating chemical
-structures and consider strategies for optimization.
+structures and considering strategies for optimization.
 
@@
 
-Designing a molecule, which is selective to one target over another will often be considered a 
-successful outcome in a drug design process. If the target in question is a protein, it is easy
-to understand that as sequence identity amongst residues, which contribute to potential small 
-molecule binding sites remains high, so does the issue of selectivity. That is, the molecule being 
+Designing a molecule that is selective in binding to one target rather than another is often considered a 
+successful outcome in a drug design process. If the target in question is a protein, high sequence similarity
+between the binding site and that of other proteins will make such selectivity more challenging. 
+That is, the molecule being 
 designed may also bind to the equivalent binding site in the closely related protein 
-target and may lead to undesirable consequences. The ChEMBL data model links molecules to 
-targets using different activity types recorded in the literature. Using activity types, 
-which act as a measure of binding affinity, such as $IC_{50}$ or $K_{i}$ and applying activity value cuts
-offs it is possible to identify molecules which have higher binding affinity to certain targets 
-compared to others. Taking this one step further it is possible to identify a set molecules
-which have a high affinity to protein A (e.g. $IC_{50}$ value $<$ 50~nM) and low affinity to protein 
+target, and this may lead to undesirable consequences. The ChEMBL data model links molecules to 
+targets using different activity types recorded in the literature. Using activity types 
+that act as a measure of binding affinity, such as $IC_{50}$ or $K_{i}$, and applying activity value 
+cut-offs, it is possible to identify molecules which have higher binding affinity to certain targets 
+compared to others. Taking this one step further, it is possible to identify a set of molecules
+that have a high affinity to protein A (e.g. $IC_{50}$ value $<$ 50~nM) and low affinity to protein 
 B (e.g. $IC_{50}$ value $>$ 200~nM).
 
-The following SPARQL query identifies a set of 65 molecules, which based on data curated from the 
-literature and stored in the ChEMBL-RDF data model, selectively bind Human Cyclin-Dependent Kinase 2 
-(UniProt: P24941) over Human Cyclin-Dependent Kinase 4 (UniProt: P11802).
+The following SPARQL query identifies a set of 65 molecules that selectively bind Human Cyclin-Dependent Kinase 2 
+(UniProt: P24941) over Human Cyclin-Dependent Kinase 4 (UniProt: P11802), based on data curated from the 
+literature and stored in the ChEMBL-RDF data model.
 
@@
 This paper shows the evolution of the ChEMBL-RDF resource in standardizing access to the
-ChEMBL data in a machine-readable manner, by using semantic web technologies. This standardization
+ChEMBL data in a machine-readable manner using semantic web technologies. This standardization
 applies to several aspects of data exchange, including the data serialization format and
-access options, the data structure, and the ontologies used to explain what the data is about.
+access options, the data structure, and the ontologies used to explain what the data are about.
 We touched upon choices made to implement these approaches, such as a selection of
@@
 For example, for linking to other databases we selected a few predicates to link to
-other databases: rdfs:seeAlso, skos:exactMatch, owl:equivalentClass, owl:sameAs. Each
-has an increasing level of implications, ranging from "you may also be interested in" to
-"both objects are identical and all properties apply to both objects". Additionally,
+other databases: \verb+rdfs:seeAlso+, \verb+skos:exactMatch+, \verb+owl:equivalentClass+, \verb+owl:sameAs+. 
+These are ordered according to their increasing level of implication, 
+ranging from ``you may also be interested in" to ``both objects are identical 
+and all properties apply to both objects". Additionally,
 the CiTO ontology has been used to link data to the literature it was extracted from,
-which provides links that lie in between rdfs:seeAlso and skos:exactMatch.
+which provides links that lie in between \verb+rdfs:seeAlso+ and \verb+skos:exactMatch+.
 
@@
 
-However, not all information is yet using standardized ontologies. For example, the
+However, not all information has thus far been annotated 
+using standardized ontologies. For example, the 
 activity types (IC50, EC50, Potency) are not exposed using an ontology, making it
-still impossible to reason or just group over similar activity types. Likewise,
+still impossible to reason or group over similar activity types. Likewise,
 measurement units are provided as literals, making it needlessly hard to compare
 concentrations in nM and mM. For this kind of data, two new ontologies were
-recently developed. For example, the BioAssay Ontology addresses many of our
-remaining concerns around making the experimental data machine readable~\cite{Visser2011},
-while the Quantity, Unit, Dimension and Type (QUDT) ontology makes it possible
+recently developed. Firstly, the BioAssay Ontology addresses many of our
+remaining concerns around making the experimental data machine readable~\cite{Visser2011}. Secondly,
+the Quantity, Unit, Dimension and Type (QUDT) ontology makes it possible
 to express measurements in a machine readable way~\cite{QUDT}. Both approaches are
@@
 
-Finally, we stress that nothing in this paper cannot be done with other approaches, such as
-more common relational databases. In fact, that is exactly what the ChEMBL website does.
-Similarly, there is nothing principle component analysis does that you cannot do with drawings
-on paper too; it is just inappropriate to do so when computer-based approaches are so much
+Finally, we stress that nothing reported on in this paper cannot be achieved within other approaches 
+or paradigms, such as the more common relational database paradigm that is in fact the backbone of the core ChEMBL offering. 
+Similarly, there is nothing principal component analysis does that one cannot do with drawings
+on paper; it is just inappropriate to use paper when computer-based approaches are so much
 more efficient. We have presented here a linked data approach to the ChEMBL data
-showing standardized and simplified integration to other resources, and that is what
-cheminformatics is about.
+showing standardized and simplified integration to other resources, and that is one of the requirements
+core to the efforts of cheminformatics.
 
@@
 support in hosting the SPARQL end point. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE. 
+JH acknowledges funding from the EU-OPENSCREEN project. 
 The data in ChEMBL is made available by funding from the Wellcome Trust [086151/Z/08/Z].

### 2012-08-09 article.tex
Added John's middle initial

@@
 Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk}\and
-John~Overington$^6$\email{John Overington - jpo@ebi.ac.uk}\and
+John~P.~Overington$^6$\email{John P. Overington - jpo@ebi.ac.uk}\and
 Anna~Gaulton$^6$\email{Anna Gaulton - agaulton@ebi.ac.uk}\and

### 2012-08-09 article.tex
Synched address fully now
Applied Anna and Mark's comments. The biggest change, which required source code changes too, is: all ChEMBL compounds are now snap:MaterialEntities to address the "a :Molecule" SPARQL problem.

@@
 \iid(4)Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A. \\
-\iid(5)Cheminformatics and Metabolism, European Bioinformatics Institute, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\
+\iid(5)Cheminformatics and Metabolism, European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\
 \iid(6)Chemogenomics, European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\ 
@@
 ChEMBL contains descriptions for biological activities involving over a million chemical 
-entities, extracted from the scientific literature, providing a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
+entities, extracted primarily from the scientific literature, providing a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
 It is updated on a fairly frequent basis as the existing data is further curated and new data is added. 
-The ChEMBL dataset is available for download and can be browsed using a web interface. 
+The ChEMBL dataset is available for download and can be browsed using a web interface and web services.
 The former requires scientists to import the data into a relational database, while the 
@@
 The new triple dataset uses standard ontologies
-including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology~\cite{Shotton2010} for literature
+including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
 references and citations, and domain ontologies such the Protein Ontology\cite{Sidhu2006} and the Chemical Information
@@
           & http://purl.obolibrary.org/obo/ \\
+bfo     & Basic Formal Ontology~\cite{Smith2004} \\
 
@@
 Each assay measures activities against one or more particular biological targets. To each
-target the ChEMBL database associates a confidence score. Because this score is specific
+target the ChEMBL database associates a confidence score (see the Compound Selectivity
+example). Because this score is specific
 for each target, the following construct is used:
@@
 
-For drug discovery, the drugs themselves are the main topic of study.
-ChEMBL contains many different drug types, mostly small molecules,
+For drug discovery, the drug-like compounds themselves are the main topic of study.
+ChEMBL contains many different compound types, mostly small molecules,
 but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
@@
 and encodes drugs as classes, rather than instances, and they subclass
-classes defined by several ontologies for these drug types. For example, proteins subclass the
+classes defined by several ontologies for these types. For example, proteins subclass the
 Protein class from the PRotein Ontology (PR\_000000001), small
@@
 subclass their respective matches in the CHEBI ontology (CHEBI\_50699
-and CHEBI\_7754).
+and CHEBI\_7754). To ensure we can look up all entities for which activities
+are reported, we subclass everything from the BFO 1.1 class MaterialEntity~\cite{Smith2004},
+which is a superclass to the CHEMINF Chemical Entity class.
 
-It should also be noted that the relationship of these structural classes of drug 
-to a general drug class is that of having a \textit{role} rather than a subclass relationship. 
+It should also be noted that the relationship of these structural classes 
+is that of having a \textit{role} of drug, rather than a subclass relationship.
+Moreover, we triples only give this drug role to approved drugs, as defined by the
+ChEMBL database. 
 This is triplified using the OBO and ChEBI ontologies, where CHEBI\_23888 is the drug role:
@@
    "Bis(3-[1 .... yl]-propionamide)" ;
+ rdfs:subClassOf bfo:MaterialEntity ;
  rdfs:subClassOf cheminf:CHEMINF_000000 ;
@@
 
-The following SPARQL query returns 1,143,682 ChEMBL molecules:
-
-%Mark Davies notes:
-%NOTE: Full set of ChEMBL molecules have been created by querying molecule\_dictionary table
-%NOTE: I have added extra triple for each molecule: <http://linkedchemistry.info/chembl/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
-%NOTE: Modified molecule compound structure can be queried here: https://wwwdev.ebi.ac.uk/chembl/sparql
+The following SPARQL query returns 1,143,682 entities for which activities are captured in ChEMBL:
 
@@
 select count(*) WHERE {
-  ?s ?p chembl:Molecule .
+  ?s ?p bfo:MaterialEntity .
 }
@@
 \end{small}
+
+It should be noted that this returns five entries less than the SQL database, because the
+triple generation requires the assay type to be defined, which is not the case for those 
+five assays.
 
@@
 The first application describes how the SPARQL end point can be used to make the data available as linked data
-via the Bio2RDF platform. The second application takes advantage of the bibliographic information
-exposed as machine readable data, in calculating citation statistics. The third application shows
-an integration of ChEMBL-RDF with ChemSpider to provide an extension for the decision support
-platform in Bioclipse. The last application shows how Chem2Bio2RDF can combine the ChEMBL data
+via the Bio2RDF platform. The second example shows how Chem2Bio2RDF can combine the ChEMBL data
 with other life sciences databases, showing the power of the Linked Data approach.
+The third application takes advantage of the bibliographic information exposed as machine readable
+data, in calculating citation statistics. The fourth use case shows an integration of ChEMBL-RDF with
+ChemSpider to provide an extension for the decision support platform in Bioclipse. The final application
+shows a more complex SPARQL query showing how information from various tables can be used to
+find compounds with a high selectivity.
 
@@
 Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as CitedIn (http://www.citedin.org).
-While the impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this narrowly defined calculation increasingly fails to capture the true reach of a paper.
-CitedIn was developed to make scientific impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public resources.
-ChEMBL contains many literature references.
-The SPARQL endpoint is queried by CitedIn and the results are presented to its users.
-Figure~\ref{fig:citedin} shows the result for the Nature paper by Gamo presenting the malaria structures, which is highly cited in ChEMBL~\cite{Gamo2010}.
-Counting citations in ChEMBL is done by the following SPARQL query:
+While the impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this
+narrowly defined calculation increasingly fails to capture the true reach of a paper. CitedIn was developed to make scientific
+impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public
+resources. ChEMBL contains many literature references. The SPARQL endpoint is queried by CitedIn and the results are presented to its users.
+Figure~\ref{fig:citedin} shows the results for the Nature paper by Gamo presenting the malaria structures~\cite{Gamo2010}
+which is cited as the reference for a large number of activity datapoints and compounds in ChEMBL.
+Counting all objects that reference this paper is done by the following SPARQL query:
 

### 2012-08-09 article.tex
More middle initials

@@
 Peter~Ansell$^3$\email{Peter Ansell - p\_ansell@yahoo.com}\and
-Antony~J.~Williams$^4$\email{Antony Williams - WilliamsA@rsc.org}\and
+Antony~J.~Williams$^4$\email{Antony J. Williams - WilliamsA@rsc.org}\and
 Valery~Tkachenko$^4$\email{Valery Tkachenko - TkachenkoV@rsc.org}\and
@@
 Bin~Chen$^7$\email{Bin Chen - binchen@indiana.edu}\and
-David~Wild$^7$\email{David Wild - djwild@indiana.edu}
+David~J.~Wild$^7$\email{David J. Wild - djwild@indiana.edu}
 }

### 2012-08-09 article.tex
Applied David's comments

@@
 Using the right to download, modify, and redistribute the ChEMBL data, two independent teams
-have previously mapped  the ChEMBL dataset to the Resource Description Framework (RDF),
-Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}. RDF is a framework where
+have previously mapped the ChEMBL dataset to the Resource Description Framework (RDF), resulting
+in Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}. RDF is a framework where
 knowledge is represented by small so-called triples, reflecting ``facts''. A fact can
@@
 A second resource in the linked open data network into which ChEMBL-RDF has been integrated is Chem2Bio2RDF.
-Chem2Bio2RDF created a single RDF repository covering over 20 public data resources pertaining to drugs, chemical compounds, carcinogens,
+Chem2Bio2RDF is a single RDF repository covering over twenty public data resources pertaining to drugs, chemical compounds, carcinogens,
 protein targets, genes, diseases, side effects, pathways and their relations~\cite{Chen2010}. The entities and their relations were

### 2012-08-10 article.tex
The authors themselves spell it all upper case: SIDER
quick check for grammar on article

@@
 This paper describes recent developments in an ongoing project converting data from the ChEMBL database into RDF triples.
-Relative to earlier versions, this updated version of ChEMBL-RDF uses recently introduced ontologies, including CHEMINF and CiTO,
-exposes more information from the database, and is now available as dereferencable, linked data.
+Relative to earlier versions, this updated version of ChEMBL-RDF uses recently introduced ontologies, including CHEMINF and CiTO;
+exposes more information from the database; and is now available as dereferencable, linked data.
 To demonstrate these new features, we present novel use cases showing further integration with
@@
 discover new, unique, and significant patterns in datasets that explain biological 
-phenomena not yet understood. This process requires integration of datasets that are both 
-sparse, and growing independently of one another. The discovery of new patterns showing the 
+phenomena not yet understood. The discovery of new patterns showing the 
 causes and effects of various biological phenomena is often beyond the scope of single datasets.
@@
 based on descriptions of its required biological properties. This process requires the effective 
-linkage of many scientific datasets~\cite{Samwald2011,Williams2012}.
+linkage of many scientific datasets that are both sparse, and growing independently of one another~\cite{Samwald2011,Williams2012}.
 
 ChEMBL contains descriptions for biological activities involving over a million chemical 
-entities, extracted primarily from the scientific literature, providing a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
+entities, extracted primarily from scientific literature. This provides a unique resource for drug researchers~\cite{Gaulton2012,Warr2009}.
 It is updated on a fairly frequent basis as the existing data is further curated and new data is added. 
@@
 latter limits the machine access to the data. ChEMBL also provides web services for programmatic
-access, but these are limited to specific types of query as encoded in the web service API.
+access, but these are limited to the specific types of queries which are encoded in the web service API.
 
@@
 be given on any topic, and RDF therefore talks about \textit{things} very generically~\cite{Miller:04:RP}. Things in RDF
-are represented by Universal Resource Identifiers (URI) and when the more specialized
-Universal Resource Locator (URL) is used, it creates \emph{linked data}~\cite{Samwald2011}. RDF is
+are represented by Universal Resource Identifiers (URI). When the URI can be resolved, 
+for example using the HTTP protocol, it creates \emph{linked data}~\cite{Samwald2011}. RDF is
 developed as an open standard by the World Wide Web Consortium, and provides a knowledge
@@
 is that of having a \textit{role} of drug, rather than a subclass relationship.
-Moreover, we triples only give this drug role to approved drugs, as defined by the
+Moreover, the triples only attach this drug role to approved drugs, as defined by the
 ChEMBL database. 
-This is triplified using the OBO and ChEBI ontologies, where CHEBI\_23888 is the drug role:
+This is triplified using the OBO and ChEBI ontologies in the following manner, where CHEBI\_23888 is the drug role:
 
@@
 The Bio2RDF project provides both resolvable Linked Data URIs using a generic Linked Data
-server~\cite{Ansell2011} to access SPARQL endpoints for a range of scientific databases~\cite{Belleau2008}.
+server~\cite{Ansell2011} and SPARQL endpoints containing RDF data from a range of scientific databases~\cite{Belleau2008}.
 A number of these databases are referenced in ChEMBL, including ChEBI, PubMed, and both the
@@
 utilize it as both a deposition platform for the community to contribute novel data, as
-well as a platform for annotation and curation for existing data. Studies have shown that there are data
-quality issues in many of the public compound databases~\cite{Williams2011}. ChemSpider has become a
-valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
+well as a platform for annotation and curation for existing data. ChemSpider has become a
+valuable resource for curated data, especially chemical-compound name mappings~\cite{Williams2011}. ChemSpider
 is presently providing the chemical structure, substructure and similarity searching services 
@@
 The data mappings between the CSIDs and ChEMBL IDs are released to the 
-community under the Creative Common Attribution-Share Alike license (CC-BY-SA 3.0). Attribution 
+community under the Creative Commons Attribution, Share Alike, license (CC-BY-SA 3.0). Attribution 
 should be made to the original ChEMBL database, Open PHACTS, and ChemSpider.
@@
 the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
-The targets can be also linked to side effects via protein-protein interactions by searching across ChEMBL, HPRD and Sider in
+The targets can be also linked to side effects via protein-protein interactions by searching across ChEMBL, HPRD and SIDER in
 Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the

### 2012-08-10 article.tex
Added twitter accounts

@@
 
-\author{Egon~L.~Willighagen$^{1,*}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl}\and
+\author{Egon~L.~Willighagen$^{1,*}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl, \url{http://twitter.com/egonwillighagen}}\and
 Andra~Waagmeester$^1$\email{Andra Waagmeester - andra.waagmeester@maastrichtuniversity.nl}\and
@@
 Peter~Ansell$^3$\email{Peter Ansell - p\_ansell@yahoo.com}\and
-Antony~J.~Williams$^4$\email{Antony J. Williams - WilliamsA@rsc.org}\and
+Antony~J.~Williams$^4$\email{Antony J. Williams - WilliamsA@rsc.org, \url{http://twitter.com/ChemConnector}}\and
 Valery~Tkachenko$^4$\email{Valery Tkachenko - TkachenkoV@rsc.org}\and
 Janna~Hastings$^5$\email{Janna Hastings - hastings@ebi.ac.uk}\and
-Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk}\and
+Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk\url{http://twitter.com/mdavies12}}\and
 John~P.~Overington$^6$\email{John P. Overington - jpo@ebi.ac.uk}\and
 Anna~Gaulton$^6$\email{Anna Gaulton - agaulton@ebi.ac.uk}\and
-Bin~Chen$^7$\email{Bin Chen - binchen@indiana.edu}\and
-David~J.~Wild$^7$\email{David J. Wild - djwild@indiana.edu}
+Bin~Chen$^7$\email{Bin Chen - binchen@indiana.edu, \url{http://twitter.com/binchenindiana}}\and
+David~J.~Wild$^7$\email{David J. Wild - djwild@indiana.edu, \url{http://twitter.com/davidjohnwild}}
 }

### 2012-08-11 article.tex
Added a missing separator

@@
 Janna~Hastings$^5$\email{Janna Hastings - hastings@ebi.ac.uk}\and
-Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk\url{http://twitter.com/mdavies12}}\and
+Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk, \url{http://twitter.com/mdavies12}}\and
 John~P.~Overington$^6$\email{John P. Overington - jpo@ebi.ac.uk}\and

### 2012-08-11 article.tex
Added Peter's twitter account

@@
 Ola Spjuth$^2$\email{Ola Spjuth - ola.spjuth@farmbio.uu.se}\and
-Peter~Ansell$^3$\email{Peter Ansell - p\_ansell@yahoo.com}\and
+Peter~Ansell$^3$\email{Peter Ansell - p\_ansell@yahoo.com, \url{http://twitter.com/p_ansell}}\and
 Antony~J.~Williams$^4$\email{Antony J. Williams - WilliamsA@rsc.org, \url{http://twitter.com/ChemConnector}}\and

### 2012-08-14 article.tex
Removed unused packages
Defaulted to URI

@@
 \usepackage[utf8]{inputenc} %unicode support
-%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
 \urlstyle{rm}
-%\usepackage[citestyle=numeric-comp,sorting=none,firstinits=true,doi=false,isbn=false,url=false]{biblatex}
 \usepackage{cite}
-%\setlength{\bibitemsep}{0cm}                                                                                                                
-%\renewbibmacro{in:}{%                                                                                                                       
-%  \ifentrytype{article}{}{%                                                                                                                 
-%  \printtext{\bibstring{in}\intitlepunct}}}                                                                                                 
-%\addbibresource{article.bib}
-
-%\usepackage{amsmath}
-%\usepackage{endnotes}
 \usepackage{graphicx}
-%\usepackage{tikz}
 
@@
 \newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}
-
 %Publication style settings
@@
 platform~\cite{kasabi} was used for a period of time, until that project was discontinued.
-The Uppsala SPARQL end point still uses resource IRIs based on the Kasabi approach, following a similar
-IRI pattern where linkedchemistry.info/chembl is replaced by data.kasabi.com/dataset/chembl-rdf.
+The Uppsala SPARQL end point still uses resource URIs based on the Kasabi approach, following a similar
+URI pattern where linkedchemistry.info/chembl is replaced by data.kasabi.com/dataset/chembl-rdf.
 
@@
 The InChI is a unique identifier for (small) organic molecules, and has been previously used
-to define unique IRIs for molecules~\cite{Bradley2009,Willighagen2011}. While IRIs are theoretically unlimited in length,
-in practice many web browsers and servers limit the length of IRIs. Virtuoso is, unfortunately,
-one such system that supports only IRIs of up to a certain length. Therefore, InChI-based links are only created for smaller molecules. 
+to define unique URIs for molecules~\cite{Bradley2009,Willighagen2011}. While URIs are theoretically unlimited in length,
+in practice many web browsers and servers limit the length of URIs. Virtuoso is, unfortunately,
+one such system that supports only URIs of up to a certain length. Therefore, InChI-based links are only created for smaller molecules. 
 Almost 1.3 million links to OpenMolecules RDF were created in a similar manner to:

### 2012-08-14 article.tex
Removed the bit about providing full detail, because that actually depends on how the URI is resolved. It may also redirect to a publisher's web page for that article.

@@
 the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) are used to link
-to authoritative databases that provide full detail about the references, as outlined later in this paper.
+to authoritative database, as outlined later in this paper.
 The properties are provided using the BIBO ontology:

### 2012-08-14 article.tex
Linking is primarily about linking to other databases, and not linking to other RDF versions of the same database :)
Added Andra's Twitter account

@@
 \author{Egon~L.~Willighagen$^{1,*}$\email{Egon~L.~Willighagen - egon.willighagen@maastrichtuniversity.nl, \url{http://twitter.com/egonwillighagen}}\and
-Andra~Waagmeester$^1$\email{Andra Waagmeester - andra.waagmeester@maastrichtuniversity.nl}\and
+Andra~Waagmeester$^1$\email{Andra Waagmeester - andra.waagmeester@maastrichtuniversity.nl, \url{http://twitter.com/andrawaag}}\and
 Ola Spjuth$^2$\email{Ola Spjuth - ola.spjuth@farmbio.uu.se}\and
@@
 
-To integrate ChEMBL-RDF with other RDF versions of scientific datasets, we link out to various resources.
+To integrate ChEMBL-RDF with RDF versions of other scientific datasets, we link out to various resources.
 These links are shown in Figure~\ref{2}. Triples for compounds link out to ChemSpider 

### 2012-08-14 article.tex
Added a citation for VOS
Added the Nature Precedings presentation for CitedIn

@@
 To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point is being
-hosted at Uppsala University, using the Open Source Virtuoso software. Usage is free, but the
+hosted at Uppsala University, using the Open-Source Edition of the Virtuoso software~\cite{Virtuoso}. Usage is free, but the
 allowed volume of querying is capped, based on the estimated computational effort. Second, resources have
@@
 
-Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as CitedIn (http://www.citedin.org).
+Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as CitedIn (http://www.citedin.org)~\cite{CitedIn2011}.
 While the impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this

### 2012-08-14 article.tex
Added some text on where the Details button points to

@@
 		\newline
-		\caption[wee]{Screenshot of the CitedIn web service showing a paper cited 94671 times in the ChEMBL database.}
+		\caption[wee]{Screenshot of the CitedIn web service showing a paper cited 94671 times in the ChEMBL database.
+		   The \textit{Details} button on the webpage links to a ChEMBL webpage with detail on what parts of the database are
+		   linked to that paper.}
 	\label{fig:citedin}
@@
 \end{small}
+
+The CitedIn summary for ChEMBL also provides a \textit{Details} button that links back to a ChEMBL webpage providing
+a \textit{Document Report Card} showing in more detail what targets, assays, compounds, and activities in the database
+are linked to that paper.
 

### 2012-08-15 article.tex
Fixed the link to Bio2RDF EC numbers: that is not an equivalence relationship, but a classification

@@
  chembl:hasTaxonomy <http://bio2rdf.org/taxonomy:9606> ;
- owl:sameAs <http://bio2rdf.org/ec:2.7.11.1> ;
+ rdfs:subClassOf <http://bio2rdf.org/ec:2.7.11.1> ;
  owl:sameAs <http://bio2rdf.org/uniprot:Q8IYT8> .

### 2012-08-17 article.tex
Restructured the discussion section, to actually make it read like a discussion

@@
 
+The predicates used to capture the links to these databases are
+\verb+rdfs:seeAlso+, \verb+skos:exactMatch+, \verb+owl:equivalentClass+, and \verb+owl:sameAs+. 
+These relations are ordered according to their increasing level of implication:
+ranging from ``you may also be interested in'' with \verb+rdfs:seeAlso+ to
+``both objects are identical  and all properties apply to both objects'' for
+\verb+owl:sameAs+.
+
 \begin{figure}[t]
@@
 
-For papers with PubMed identifiers we also link out to Bio2RDF:
+For resources from the biomedical with PubMed identifiers we also link out to Bio2RDF:
 
@@
 This paper shows the evolution of the ChEMBL-RDF resource in standardizing access to the
-ChEMBL data in a machine-readable manner using semantic web technologies. This standardization
-applies to several aspects of data exchange, including the data serialization format and
-access options, the data structure, and the ontologies used to explain what the data are about.
-We touched upon choices made to implement these approaches, such as a selection of
-RDF technologies, and chosen ontologies.
+ChEMBL data in a machine-readable manner using semantic web technologies. We showed here
+that the conversion into RDF allows a number of new integrations with other resources.
+It should be clear that these applications can be made using other technologies too, but
+using the RDF standards it is easier to implement them. More importantly, the solutions
+will be generally applicable to any system using these open standards. However, a number
+of choices need discussion.
 
-For example, for linking to other databases we selected a few predicates to link to
-other databases: \verb+rdfs:seeAlso+, \verb+skos:exactMatch+, \verb+owl:equivalentClass+, \verb+owl:sameAs+. 
-These are ordered according to their increasing level of implication, 
-ranging from ``you may also be interested in" to ``both objects are identical 
-and all properties apply to both objects". Additionally,
-the CiTO ontology has been used to link data to the literature it was extracted from,
-which provides links that lie in between \verb+rdfs:seeAlso+ and \verb+skos:exactMatch+.
+For example, much can be argued about the choice of ontologies, and the lack of an
+ontology at places where we decided to use custom classes and predicates. Indeed, the
+choices have only partly been decided on ontological arguments, but are mostly based
+on pragmatic approaches to fulfill our needs for the applications. It should be noted,
+however, that the ChEMBL data base captures a wide variety of experimental studies
+and that semantics are not fully specified in the source from which the data was
+extracted. Clearly, missing semantics must not be made up: it is better to leave
+some semantics undefined than making incorrect assumptions. Indeed, during the
+development of the applications some incorrect assumptions have been discovered and
+fixed. Therefore, the goal of this work is the conversion of the ChEMBL data into
+RDF, and not an ontological modeling of the domain.
 
-Another ontology that we adopted in this ChEMBL-RDF 13 is the cheminformatics ontology
-CHEMINF, allowing us to encode numeric and non-numeric descriptors of chemical entities
-in the same standardized way. Using this approach we expose identity descriptors such
+For example, this applies for the links to other databases: when linking between
+two databases based on a shared identifier, we make assumptions. Besides the mere
+aspect that either database may contain typing mistakes in databases, additional
+error may be introduced because at least the assumption is made in the ChEMBL database
+that entries in that database really matches the data linked to in the other database.
+That is the reason why we used different predicates for linking to other databases.
+This problem, however, is general, and applies to other methods to link databases with
+identifiers too~\cite{VanIersel2010}.
+
+Also debatable is the choice of the ontologies. For example, in this release of the
+ChEMBL-RDF data, we adopted the cheminformatics ontology, CHEMINF. Using this ontology
+we can encode numeric and non-numeric descriptors of chemical entities in the same
+standardized way. Using this approach we expose identity descriptors such
 as the SMILES string, InChI, and InChIKey, and molecular properties including the
 logP, rotatable bond count, and hydrogen bond donor and acceptor counts.
+The design of this annotation is more complex than strictly needed for a translation
+of the data in the ChEMBL relational database into RDF. Moreover, this also
+affects the querying of the data, where a more complex SPARQL query is needed.
+While we found that our selection of ontologies fit well with our applications,
+we anticipate that we will select more expressive ontologies in the future.
 
-However, not all information has thus far been annotated 
-using standardized ontologies. For example, the 
-activity types (IC50, EC50, Potency) are not exposed using an ontology, making it
-still impossible to reason or group over similar activity types. Likewise,
-measurement units are provided as literals, making it needlessly hard to compare
-concentrations in nM and mM. For this kind of data, two new ontologies were
-recently developed. Firstly, the BioAssay Ontology addresses many of our
-remaining concerns around making the experimental data machine readable~\cite{Visser2011}. Secondly,
+For example, not all information in the databases has so far been annotated using
+standard ontologies. That includes the activity types (``IC50'', ``EC50'', and ``Potency'' in ChEMBL), making it
+still impossible to reason or group over similar activity types. For example,
+the IC$_{50}$ and K$_i$ end point are closely related, but that is currently not
+ontologically expressed. Similarly, measurement units are currently provided as
+string literals, making it needlessly hard to compare concentrations in nM and mM.
+If we further allow to make some assumptions, we can even search for activities
+expressed in \mu g/ml with query concentration in nM.
+
+To address these issues, two ontologies were
+recently developed. First, the BioAssay Ontology addresses many of our
+remaining concerns around making the experimental data machine readable~\cite{Visser2011}. Second,
 the Quantity, Unit, Dimension and Type (QUDT) ontology makes it possible
@@
 
-Finally, we stress that nothing reported on in this paper cannot be achieved within other approaches 
-or paradigms, such as the more common relational database paradigm that is in fact the backbone of the core ChEMBL offering. 
-Similarly, there is nothing principal component analysis does that one cannot do with drawings
-on paper; it is just inappropriate to use paper when computer-based approaches are so much
-more efficient. We have presented here a linked data approach to the ChEMBL data
-showing standardized and simplified integration to other resources, and that is one of the requirements
+Taking these points into account, we believe that the adoption of the RDF standards provide
+novel use cases with the ChEMBL data. More importantly, the adoption of explicit
+semantics simplifies integration to other resources, and that is one of the requirements
 core to the efforts of cheminformatics.

### 2012-08-17 article.tex
Adding why RDF is important as a solution

@@
 If we further allow to make some assumptions, we can even search for activities
-expressed in \mu g/ml with query concentration in nM.
+expressed in $\mu$g/ml with query concentrations in nM. Without introducing machine-readable
+semantics, such conversions are less easily automated.
 

### 2012-08-18 article.tex
Small edits based on Antony's comments

@@
 and encodes drugs as classes, rather than instances, and they subclass
-classes defined by several ontologies for these types. For example, proteins subclass the
+other classes defined by several ontologies for these types. For example, proteins subclass the
 Protein class from the PRotein Ontology (PR\_000000001), small
@@
 these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
-CHEMINF ontology. Here is, for example, the ALogP value (as distributed with ChEMBL and
+CHEMINF ontology. Here is, for example, the logP value (as distributed with ChEMBL and
 calculated with ACD/Labs software):
@@
 
-The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 fewer than appears 
+The following SPARQL query returns the 44,681 ChEMBL documents. This is one fewer than appears 
 in the ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as the result does not include the doc\_id -1, as this refers to the 
@@
 Notice here the use of \verb+owl:equivalentClass+ since in CHEMINF molecules are defined as
-as classes rather than instances.
+classes rather than instances.
 
@@
 still impossible to reason or group over similar activity types. For example,
-the IC$_{50}$ and K$_i$ end point are closely related, but that is currently not
+the IC$_{50}$ and K$_i$ end points are closely related, but that is currently not
 ontologically expressed. Similarly, measurement units are currently provided as

### 2012-08-20 article.tex
edit discussion that was a little clumsy and possibly more negative than needed

@@
 
-For example, much can be argued about the choice of ontologies, and the lack of an
-ontology at places where we decided to use custom classes and predicates. Indeed, the
-choices have only partly been decided on ontological arguments, but are mostly based
-on pragmatic approaches to fulfill our needs for the applications. It should be noted,
-however, that the ChEMBL data base captures a wide variety of experimental studies
-and that semantics are not fully specified in the source from which the data was
-extracted. Clearly, missing semantics must not be made up: it is better to leave
-some semantics undefined than making incorrect assumptions. Indeed, during the
+Much could be argued about the choice of ontologies, and the lack of an
+ontology in places where we decided to use custom classes and predicates. Indeed, the
+choices have only partly been decided using ontological arguments. Most of the choices are based
+on pragmatic approaches to enable solutions to the range of use cases we describe here. It should be noted,
+however, that the ChEMBL database captures a wide variety of experimental studies
+and that semantics are not fully specified in the primary source from which the ChEMBL data was
+extracted. As it is not possible to infer the original intentions in these cases, 
+it is our opinion that it is better to leave some semantics undefined. During the 
 development of the applications some incorrect assumptions have been discovered and
-fixed. Therefore, the goal of this work is the conversion of the ChEMBL data into
+fixed, but in the main we have tried to avoid making assumptions if they are not generally known to be valid. 
+Therefore, the goal of this work is the conversion of the ChEMBL data into
 RDF, and not an ontological modeling of the domain.
 
-For example, this applies for the links to other databases: when linking between
-two databases based on a shared identifier, we make assumptions. Besides the mere
-aspect that either database may contain typing mistakes in databases, additional
-error may be introduced because at least the assumption is made in the ChEMBL database
-that entries in that database really matches the data linked to in the other database.
-That is the reason why we used different predicates for linking to other databases.
-This problem, however, is general, and applies to other methods to link databases with
-identifiers too~\cite{VanIersel2010}.
+Once case where Linked Data publishers must avoid implying strong semantics in many cases is
+in relation to links to other databases. When linking between two databases based on a shared identifier,
+Linked Data publishers make assumptions including syntax assumptions that the identifiers 
+are correct and present in the target database, along with semantic assumptions inferring that the
+that entries identified in the target database have a scientifically meaningful relationship with the current database.
+It is for this reason that the ChEMBL-RDF dataset uses different predicates for linking to various databases.
+This problem, however, is general, applying also to other cases where databases link out to third-party sources using identifiers~\cite{VanIersel2010}.
 
-Also debatable is the choice of the ontologies. For example, in this release of the
+Also debatable is the choice of ontologies used to describe the dataset in RDF. As an example, in this release of the
 ChEMBL-RDF data, we adopted the cheminformatics ontology, CHEMINF. Using this ontology
-we can encode numeric and non-numeric descriptors of chemical entities in the same
-standardized way. Using this approach we expose identity descriptors such
+we can encode numeric and non-numeric descriptors of chemical entities in a consistent
+manner. This method is used to consistently expose identity descriptors such
 as the SMILES string, InChI, and InChIKey, and molecular properties including the
 logP, rotatable bond count, and hydrogen bond donor and acceptor counts.
-The design of this annotation is more complex than strictly needed for a translation
-of the data in the ChEMBL relational database into RDF. Moreover, this also
-affects the querying of the data, where a more complex SPARQL query is needed.
-While we found that our selection of ontologies fit well with our applications,
-we anticipate that we will select more expressive ontologies in the future.
+However, the design of this annotation is more general than may be needed for a translation
+of the data in the ChEMBL relational database into RDF. Although this results in more 
+complex SPARQL queries, it allows generic queries that may not otherwise be easy to construct.
+While the selection of ontologies used was found to fit well with our applications,
+we anticipate that we will select more expressive ontologies in future iterations of the project.
 
-For example, not all information in the databases has so far been annotated using
-standard ontologies. That includes the activity types (``IC50'', ``EC50'', and ``Potency'' in ChEMBL), making it
-still impossible to reason or group over similar activity types. For example,
-the IC$_{50}$ and K$_i$ end points are closely related, but that is currently not
-ontologically expressed. Similarly, measurement units are currently provided as
-string literals, making it needlessly hard to compare concentrations in nM and mM.
-If we further allow to make some assumptions, we can even search for activities
-expressed in $\mu$g/ml with query concentrations in nM. Without introducing machine-readable
-semantics, such conversions are less easily automated.
+Some information in the databases has not yet been annotated using standard ontologies. 
+For example, the activity types in ChEMBL, ``IC50'', ``EC50'', and ``Potency'', are not exposed in the RDF dataset
+meaning that it is not yet possible to reason or group over similar activity types.
+When the activity types are exposed queries could be constructed based on the knowledge 
+that the IC$_{50}$ and K$_i$ end points are closely related. Similarly, measurement units are currently provided as
+string literals, making it difficult to compare concentrations when some are given in nM and others in mM.
+In the future it may also be possible, relying on some assumptions, to crossmatch activities
+expressed in $\mu$g/ml with query concentrations in nM. These additional semantics are necessary 
+to translate raw experimental data into semantically useful, machine-readable information.
 
-To address these issues, two ontologies were
-recently developed. First, the BioAssay Ontology addresses many of our
-remaining concerns around making the experimental data machine readable~\cite{Visser2011}. Second,
-the Quantity, Unit, Dimension and Type (QUDT) ontology makes it possible
-to express measurements in a machine readable way~\cite{QUDT}. Both approaches are
-currently explored to be applied to ChEMBL-RDF as part of work in the
-Open PHACTS project~\cite{Williams2012}.
+Two ontologies were recently developed that may address these issues for ChEMBL-RDF in future. Firstly, the BioAssay Ontology, which addresses many of our
+remaining concerns around making the experimental data machine readable~\cite{Visser2011}. Secondly,
+the Quantity, Unit, Dimension and Type (QUDT) ontology which makes it possible
+to express measurements in a standardised machine readable way~\cite{QUDT}. Both approaches are 
+being explored in the Open PHACTS project and will be applied to ChEMBL-RDF as the Open PHACTS project progresses~\cite{Williams2012}.
 
-Taking these points into account, we believe that the adoption of the RDF standards provide
-novel use cases with the ChEMBL data. More importantly, the adoption of explicit
-semantics simplifies integration to other resources, and that is one of the requirements
-core to the efforts of cheminformatics.
+Taking these points into account, we believe that the adoption of RDF and community based ontologies allows
+novel use cases to be solved using the ChEMBL data that would not otherwise be achievable. More importantly, the adoption of explicit
+semantics simplifies integration to other resources, as others can more easily understand the intent and purpose of each part of the ChEMBL dataset.
+This is a requirement which is essential to the efforts of cheminformatics researchers wishing to infer knowledge from both ChEMBL and other datasets.
 

### 2012-11-20 article.tex
Moved the figures and the table to the end
Fixed the syntax of the example

@@
 
-\begin{table*}
-\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
-\begin{center}
-\begin{tabular}{ll}
-\hline
-\multicolumn{2}{l}{\textbf{Common Vocabularies}} \\
-bibo    & Bibliography Ontology~\cite{Giasson2011} \\
-        & http://purl.org/ontology/bibo/ \\
-chebi   & Chemical Entities of Biological Interest~\cite{DeMatos2010} \\
-        & http://purl.org/obo/owl/CHEBI\# \\
-cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
-        & http://semanticscience.org/resource/ \\
-cito    & Citation Typing Ontology~\cite{Shotton2010} \\
-        & http://purl.org/spar/cito/ \\
-obo / pro & OBO \& PRotein Ontology~\cite{Sidhu2006} \\
-          & http://purl.obolibrary.org/obo/ \\
-bfo     & Basic Formal Ontology~\cite{Smith2004} \\
-
-\multicolumn{2}{l}{\textbf{ChEMBL-RDF Namespace}} \\
-chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
-
-\multicolumn{2}{l}{\textbf{ChEMBL-RDF Prefixes }}\\
-act    & http://linkedchemistry.info/chembl/activity/ \\
-assay  & http://linkedchemistry.info/chembl/assay/ \\
-mol    & http://linkedchemistry.info/chembl/molecule/ \\
-res    & http://linkedchemistry.info/chembl/resource/ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
 To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point is being
@@
 classes are linked together, detailed in this section at a triple level.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{figs/relations}
-\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
-of common classes, while others are instances.}\label{f1}
-\end{center}
-\end{figure}
 
@@
  cheminf:CHEMINF_000200
-  m41:inchikey , m41:smiles , m41:inchi ;
+  m41:inchikey , m41:smiles , m41:inchi .
 
@@
 
-\begin{figure}[t]
-{\center
-\includegraphics[width=0.45\textwidth]{figs/lodgraph}
-\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
-Edges are labeled by the predicates making the links.}\label{2}
-}
-\end{figure}
-
 \subsubsection*{Linking out to Bio2RDF}
@@
 
-\begin{figure*}[!ht]
-		\begin{center}
-		\includegraphics[width=15cm]{citedin.png}
-		\newline
-		\caption[wee]{Screenshot of the CitedIn web service showing a paper cited 94671 times in the ChEMBL database.
-		   The \textit{Details} button on the webpage links to a ChEMBL webpage with detail on what parts of the database are
-		   linked to that paper.}
-	\label{fig:citedin}
-		\end{center}
-\end{figure*}
-
 \subsection*{CitedIn}
@@
 structures and considering strategies for optimization.
-
-\begin{figure*}[!ht]
-		\begin{center}
-		\includegraphics[width=15cm]{bioclipse-ds.png}
-		\newline
-		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
-The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
-	\label{fig:bioclipse-ds}
-		\end{center}
-\end{figure*}
 
@@
 
+\newpage
+
+\begin{figure}[t]
+\begin{center}
+relations.pdf
+%\includegraphics[width=0.7\textwidth]{figs/relations}
+\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
+of common classes, while others are instances.}\label{f1}
+\end{center}
+\end{figure}
+
+\begin{figure}[t]
+{\center
+lodgraph.pdf
+%\includegraphics[width=0.45\textwidth]{figs/lodgraph}
+\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
+Edges are labeled by the predicates making the links.}\label{2}
+}
+\end{figure}
+
+\begin{figure*}[!ht]
+		\begin{center}
+		citedin.png
+		%\includegraphics[width=15cm]{citedin.png}
+		\caption[wee]{Screenshot of the CitedIn web service showing a paper cited 94671 times in the ChEMBL database.
+		   The \textit{Details} button on the webpage links to a ChEMBL webpage with detail on what parts of the database are
+		   linked to that paper.}
+	\label{fig:citedin}
+		\end{center}
+\end{figure*}
+
+\begin{figure*}[!ht]
+		\begin{center}
+		bioclipse-ds.png
+		%\includegraphics[width=15cm]{bioclipse-ds.png}
+		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
+The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
+	\label{fig:bioclipse-ds}
+		\end{center}
+\end{figure*}
+
+\newpage
+
+\begin{table*}
+\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
+\begin{center}
+\begin{tabular}{ll}
+\hline
+\multicolumn{2}{l}{\textbf{Common Vocabularies}} \\
+bibo    & Bibliography Ontology~\cite{Giasson2011} \\
+        & http://purl.org/ontology/bibo/ \\
+chebi   & Chemical Entities of Biological Interest~\cite{DeMatos2010} \\
+        & http://purl.org/obo/owl/CHEBI\# \\
+cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
+        & http://semanticscience.org/resource/ \\
+cito    & Citation Typing Ontology~\cite{Shotton2010} \\
+        & http://purl.org/spar/cito/ \\
+obo / pro & OBO \& PRotein Ontology~\cite{Sidhu2006} \\
+          & http://purl.obolibrary.org/obo/ \\
+bfo     & Basic Formal Ontology~\cite{Smith2004} \\
+
+\multicolumn{2}{l}{\textbf{ChEMBL-RDF Namespace}} \\
+chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
+
+\multicolumn{2}{l}{\textbf{ChEMBL-RDF Prefixes }}\\
+act    & http://linkedchemistry.info/chembl/activity/ \\
+assay  & http://linkedchemistry.info/chembl/assay/ \\
+mol    & http://linkedchemistry.info/chembl/molecule/ \\
+res    & http://linkedchemistry.info/chembl/resource/ \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
 \end{bmcformat}

### 2012-11-20 article.tex
Better English

@@
 based on descriptions of its required biological properties. This process requires the effective 
-linkage of many scientific datasets that are both sparse, and growing independently of one another~\cite{Samwald2011,Williams2012}.
+linkage of many scientific datasets that are both sparse and are growing independently from each
+other~\cite{Samwald2011,Williams2012}.
 

### 2012-11-20 article.tex
Better English

@@
 are reported, we subclass everything from the BFO 1.1 class MaterialEntity~\cite{Smith2004},
-which is a superclass to the CHEMINF Chemical Entity class.
+which is a superclass of the CHEMINF Chemical Entity class.
 

### 2012-11-20 article.tex
Better English

@@
 novel use cases to be solved using the ChEMBL data that would not otherwise be achievable. More importantly, the adoption of explicit
-semantics simplifies integration to other resources, as others can more easily understand the intent and purpose of each part of the ChEMBL dataset.
-This is a requirement which is essential to the efforts of cheminformatics researchers wishing to infer knowledge from both ChEMBL and other datasets.
+semantics simplifies integration with other resources, as others can more easily understand the intent and purpose of each part of the ChEMBL dataset.
+This is a requirement which is essential to the efforts of cheminformatics researchers wishing to infer knowledge
+from ChEMBL as well as other datasets.
 

### 2012-11-20 article.tex
Fixed the subscripting and the greek μ
Added the suggested 'such as', and added quotes the activities in the 'Data structure' section to be consistent with the 'such as' list

@@
 are captured by the ChEMBL database.
-The top five types are Potency (43\%), IC50 (13\%), MIC (4.6\%), Inhibition (3.7\%),
-and Ki (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
+The top five types are ``Potency'' (43\%), ``IC50'' (13\%), ``MIC'' (4.6\%), ``Inhibition'' (3.7\%),
+and ``Ki'' (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
 linked to, an ontology.
@@
 targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
-In this case, 36 drugs causing heart disease were linked to 87 unique protein targets ($IC_50$ $<$ 10um). The top two most common targets
+In this case, 36 drugs causing heart disease were linked to 87 unique protein targets ($IC_{50}$ $<$ 10 $\mu$m). The top two most common targets
 of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
@@
 Some information in the databases has not yet been annotated using standard ontologies. 
-For example, the activity types in ChEMBL, ``IC50'', ``EC50'', and ``Potency'', are not exposed in the RDF dataset
+For example, the activity types in ChEMBL, such as ``IC50'', ``EC50'', and ``Potency'', are not exposed in the RDF dataset
 meaning that it is not yet possible to reason or group over similar activity types.

### 2012-11-20 article.tex
Applied the linguistic suggestion

@@
 Taking these points into account, we believe that the adoption of RDF and community based ontologies allows
-novel use cases to be solved using the ChEMBL data that would not otherwise be achievable. More importantly, the adoption of explicit
+novel use cases to be addressed using the ChEMBL data that would not otherwise be achievable. More importantly, the adoption of explicit
 semantics simplifies integration with other resources, as others can more easily understand the intent and purpose of each part of the ChEMBL dataset.

### 2013-01-14 article.tex
Textual fix: s/once case/one case/
Added a pg on future versions and community support

@@
 
-Once case where Linked Data publishers must avoid implying strong semantics in many cases is
+One case where Linked Data publishers must avoid implying strong semantics in many cases is
 in relation to links to other databases. When linking between two databases based on a shared identifier,
@@
 from ChEMBL as well as other datasets.
+
+The ongoing ChEMBL-RDF project is Open Source making it possible to others to convert ChEMBL to RDF too.
+The ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
+exposing ChEMBL as RDF too (EW, PA). Various version may arise, exposing the data in different ontologies,
+supporting community needs, though convergence may arise after a period of experimenting with ontologies that
+will not survive the test of time. The availability of Open Source scripts ensure an easy switch from one
+ontology to another.
 

### 2013-01-14 article.tex
Extended the descriptions of types and the drug role more verbosely

@@
 The ChEMBL database
-recognizes various types: \verb+pro:PR_0000001+, \verb+chembl:ADMET+, \verb+chembl:CELL-LINE+,
+recognizes various target types: \verb+pro:PR_0000001+, \verb+chembl:ADMET+, \verb+chembl:CELL-LINE+,
 \verb+chembl:NUCLEIC-ACID+, \verb+chembl:ORGANISM+, \verb+chembl:SUBCELLULAR+, \verb+chembl:TISSUE+,
@@
 and encodes drugs as classes, rather than instances, and they subclass
-other classes defined by several ontologies for these types. For example, proteins subclass the
-Protein class from the PRotein Ontology (PR\_000000001), small
+other classes defined by several ontologies for these types. For example, protein drugs subclass the
+Protein class from the PRotein Ontology (PR\_000000001), instead that these drugs are
+of the Protein type. Each drug is a class itself. Likewise, small
 molecules subclass the Chemical Entity class from the CHEMINF
@@
 
-It should also be noted that the relationship of these structural classes 
-is that of having a \textit{role} of drug, rather than a subclass relationship.
+It should also be noted that all drug-like entities in ChEMBL are not ontologically
+subclassing a drug class. Instead, we give the entities the role of drug: only within a
+particular context, they would be drug. If not used as medication, they are not drug.
 Moreover, the triples only attach this drug role to approved drugs, as defined by the
-ChEMBL database. 
-This is triplified using the OBO and ChEBI ontologies in the following manner, where CHEBI\_23888 is the drug role:
+ChEMBL database. All drug-like entities in the database do not have the role of drug.
+The role is triplified using the OBO and ChEBI ontologies in the following manner,
+where CHEBI\_23888 is the ontological entry for of ``drug role'' in ChEBI:
 

### 2013-01-15 article.tex
fix minor grammar

@@
 be given on any topic, and RDF therefore talks about \textit{things} very generically~\cite{Miller:04:RP}. Things in RDF
-are represented by Universal Resource Identifiers (URI). When the URI can be resolved, 
+are represented by resources named with Universal Resource Identifiers (URI). When the URI can be resolved, 
 for example using the HTTP protocol, it creates \emph{linked data}~\cite{Samwald2011}. RDF is
@@
 
-This paper presents an update on the ChEMBL-RDF resource, including details of the latest structures
+This paper presents an update on the ChEMBL-RDF data set, including details of the latest structures
 and ontologies used to map ChEMBL to RDF, along with new use cases, showing 
@@
 The role is triplified using the OBO and ChEBI ontologies in the following manner,
-where CHEBI\_23888 is the ontological entry for of ``drug role'' in ChEBI:
+where CHEBI\_23888 is the ontological entry for ``drug role'' in ChEBI:
 
@@
 
-Each release of the ChEMBL database is accompanied by a set of release counts, which help 
+Each release of the ChEMBL database is accompanied by a set of release statistics, which help 
 provide a concise content overview of the release in question. The ChEMBL 13 release counts
@@
 
-The following SPARQL query returns 1,143,682 entities for which activities are captured in ChEMBL:
-
-\begin{small}
-\begin{verbatim}
-select count(*) WHERE {
-  ?s ?p bfo:MaterialEntity .
-}
-\end{verbatim}
-\end{small}
-
-The following SPARQL query returns 617,681 ChEMBL assays:
-
-\begin{small}
-\begin{verbatim}
-select count(*) WHERE {
-  ?s ?p chembl:Assay .
-}
-\end{verbatim}
-\end{small}
-
-It should be noted that this returns five entries less than the SQL database, because the
+A similar SPARQL query, replacing \verb+chembl:Target+ with \verb+bfo:MaterialEntity+, returns 1,143,682 entities for which activities are captured in ChEMBL, and by using \verb+chembl:Assay+, the query returns the total of 617,681 ChEMBL assays.
+It should be noted that last query returns five entries less than the SQL database, because the
 triple generation requires the assay type to be defined, which is not the case for those 
-five assays.
-
-The following SPARQL query returns the 6,933,068 ChEMBL activities:
-
-\begin{small}
-\begin{verbatim}
-select count(*) WHERE {
-  ?s ?p chembl:Activity .
-}
-\end{verbatim}
-\end{small}
-
-The following SPARQL query returns the 44,681 ChEMBL documents. This is one fewer than appears 
+five assays. With \verb+chembl:Activity+ and \verb+bibo:Article+, we count 6,933,068 ChEMBL activities and 44,681 documents,
+respectively. The document count is one fewer than appears 
 in the ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as the result does not include the doc\_id -1, as this refers to the 
 unpublished dataset, not itself a valid \verb+bibo:Article+.
-
-\begin{small}
-\begin{verbatim}
-select count(*) WHERE {
-  ?s ?p bibo:Article .
-}
-\end{verbatim}
-\end{small}
 
@@
 
-ChEMBL-RDF can be used to explore interrelated scientific datasets, and we here present a few applications. 
-The first application describes how the SPARQL end point can be used to make the data available as linked data
-via the Bio2RDF platform. The second example shows how Chem2Bio2RDF can combine the ChEMBL data
-with other life sciences databases, showing the power of the Linked Data approach.
-The third application takes advantage of the bibliographic information exposed as machine readable
-data, in calculating citation statistics. The fourth use case shows an integration of ChEMBL-RDF with
-ChemSpider to provide an extension for the decision support platform in Bioclipse. The final application
-shows a more complex SPARQL query showing how information from various tables can be used to
-find compounds with a high selectivity.
+ChEMBL-RDF can be used to explore interrelated scientific datasets, and we here present five applications,
+describing how the SPARQL end point can be used, the combination of the ChEMBL data
+with other life sciences databases, the use of the bibliographic information in calculating citation
+statistics, and how to integrate ChEMBL-RDF with ChemSpider to provide an extension for the decision
+support platform in Bioclipse.
 
@@
 
-A second resource in the linked open data network into which ChEMBL-RDF has been integrated is Chem2Bio2RDF.
+A second resource in the linked open data network into which ChEMBL-RDF was integrated is Chem2Bio2RDF.
 Chem2Bio2RDF is a single RDF repository covering over twenty public data resources pertaining to drugs, chemical compounds, carcinogens,
@@
 further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative searches~\cite{Chen2012a} and
-data mining~\cite{Chen2012b}. The ChEMBL-RDF set has been uploaded into the Chem2Bio2RDF triple store, enabling
+data mining~\cite{Chen2012b}. The ChEMBL-RDF set was uploaded into the Chem2Bio2RDF triple store, enabling
 queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL-RDF and Chem2Bio2RDF use InChI keys to present chemical
@@
 the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
-The targets can be also linked to side effects via protein-protein interactions by searching across ChEMBL, HPRD and SIDER in
+The targets can also be linked to side effects via protein-protein interactions by searching across ChEMBL, HPRD and SIDER in
 Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
@@
 impact more visible and more accurate by providing insight into where published work is being cited online in a variety of public
-resources. ChEMBL contains many literature references. The SPARQL endpoint is queried by CitedIn and the results are presented to its users.
+resources, as listed on the CitedIn website~\cite{CitedResources}.
+ChEMBL contains many literature references. The SPARQL endpoint is queried by CitedIn and the results are presented to its users.
 Figure~\ref{fig:citedin} shows the results for the Nature paper by Gamo presenting the malaria structures~\cite{Gamo2010}
@@
 
-Two ontologies were recently developed that may address these issues for ChEMBL-RDF in future. Firstly, the BioAssay Ontology, which addresses many of our
-remaining concerns around making the experimental data machine readable~\cite{Visser2011}. Secondly,
+Two ontologies were recently developed that may address these issues for ChEMBL-RDF in future.
+Firstly, the BioAssay Ontology, which addresses many of our
+remaining concerns around making the experimental data machine readable~\cite{Visser2011}. For example,
+it provides an ontology of activity types. Secondly,
 the Quantity, Unit, Dimension and Type (QUDT) ontology which makes it possible
@@
 exposing ChEMBL as RDF too (EW, PA). Various version may arise, exposing the data in different ontologies,
-supporting community needs, though convergence may arise after a period of experimenting with ontologies that
-will not survive the test of time. The availability of Open Source scripts ensure an easy switch from one
-ontology to another.
+supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that
+may not prove to be practically useful. The availability of Open Source scripts ensures that the community can contribute to the
+process of choosing a practically useful set of ontological terms.
 

### 2013-01-15 article.tex
reword the new discussion paragraph

@@
 The ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
-exposing ChEMBL as RDF too (EW, PA). Various version may arise, exposing the data in different ontologies,
-supporting community needs, though convergence may arise after a period of experimenting with ontologies that
-will not survive the test of time. The availability of Open Source scripts ensure an easy switch from one
-ontology to another.
+exposing ChEMBL as RDF (EW, PA). Different versions may arise which expose the data using different ontologies,
+supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that
+are not practically useful. The availability of Open Source scripts ensures that the community can contribute to the
+process of choosing a practically useful set of ontological terms.
 

### 2013-01-15 article.tex
minor change again to remove the implication that we have intentionally chosen bad ontologies

@@
 supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that
-are not practically useful. The availability of Open Source scripts ensures that the community can contribute to the
+may not prove to be practically useful. The availability of Open Source scripts ensures that the community can contribute to the
 process of choosing a practically useful set of ontological terms.

### 2013-01-22 article.tex
Directly linked BAO to the need of an 'activity type' ontology
Removed the repetitive, almost identical SPARQL queries

@@
 The role is triplified using the OBO and ChEBI ontologies in the following manner,
-where CHEBI\_23888 is the ontological entry for ``drug role'' in ChEBI:
+where CHEBI\_23888 is the ontological entry for of ``drug role'' in ChEBI:
 
@@
 The ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
-exposing ChEMBL as RDF (EW, PA). Different versions may arise which expose the data using different ontologies,
-supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that
-may not prove to be practically useful. The availability of Open Source scripts ensures that the community can contribute to the
-process of choosing a practically useful set of ontological terms.
+exposing ChEMBL as RDF too (EW, PA). Various version may arise, exposing the data in different ontologies,
+supporting community needs, though convergence may arise after a period of experimenting with ontologies that
+will not survive the test of time. The availability of Open Source scripts ensure an easy switch from one
+ontology to another.
 

### 2013-03-06 article.tex
Explained the design goals in more detail at the start of the Results section, and given more detail on the choice of some ontologies. But the fact is, there have not been that many design decisions otherwise.

@@
 We present here the updated ChEMBL-RDF, including a description of the mapping process 
-and the identification of links to other datasets.
+and the identification of links to other datasets. Because the primary purpose of of this
+paper is to expose the ChEMBL data as linked data, less focus has been placed on ontologically
+capturing the fine details of the pharmacological literature. This is the reason why
+many concepts are not mapped to existing ontologies, though also the lack of ontologies
+with a clear open license restrict the options.
 
@@
 For each of the common resource classes, a triple pattern was defined, following the
-data available in the relational database. Figure~\ref{f1} shows how the various resource
-classes are linked together, detailed in this section at a triple level.
+data available in the relational database. Unless the ontological meaning of the data
+in the database was very clear, the classes are annotated with concepts as implicitly
+defined by the relational database. For example, the activity concept in the triples
+merely follows the activity table in the database. As such, the concept of chembl:Activity
+is implied by the content of that table in the relational database.
 
+However, where the database content is well-defined, specific ontologies have been used.
+For example, where the database specifies literature from which data was extraced,
+the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
+compatibility with other Linked Data projects, rather than formal ontological arguments.
+We have not observed ontological contradictions based on these choices.
+
+Figure~\ref{f1} shows how the various resource classes are linked together, detailed in
+this section at a triple level.
 The core concept in the ChEMBL database is that of the biological activity.
@@
 and ``Ki'' (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
-linked to, an ontology.
+linked to, an ontology. The BioAssay Ontology could be a future option~\cite{Visser2011}.
 
@@
 The name and synonyms for each drug are
-provided as labels. When InChI and InChIKeys are available for a drug, then these are
-provided via the CHEMINF formalism (abbreviated here):
+provided as labels. If SMILES, InChI, or InChIKeys are given for a drug, then these are
+provided via the CHEMINF formalism.
+The CHEMINF ontology has been used here because of it's flexibility in defining
+descriptor types and their origin and meaning. By allowing an ontology of descriptors,
+we enable comparing values between databases semantically. Differences in, for example,
+the SMILES may then be explained by the software used for that descriptor.
+
+Example output for ChEMBL406142 looks like:
 
@@
 to authoritative database, as outlined later in this paper.
-The properties are provided using the BIBO ontology:
+The properties are provided using the BIBO ontology, which is an open sourced ontology
+for bibliographic information, compatible with the CiTO ontology:
 
@@
 Firstly, the BioAssay Ontology, which addresses many of our
-remaining concerns around making the experimental data machine readable~\cite{Visser2011}. For example,
+remaining concerns around making the experimental data machine readable. For example,
 it provides an ontology of activity types. Secondly,

### 2013-03-06 article.tex
Added a technical discussion section, answering the reviewer's question about RDF syntax errors and performance

@@
 to express measurements in a standardised machine readable way~\cite{QUDT}. Both approaches are 
-being explored in the Open PHACTS project and will be applied to ChEMBL-RDF as the Open PHACTS project progresses~\cite{Williams2012}.
+being explored in the Open PHACTS project and will be applied to ChEMBL-RDF as the Open PHACTS
+project progresses~\cite{Williams2012}.
 
@@
 The ongoing ChEMBL-RDF project is Open Source making it possible to others to convert ChEMBL to RDF too.
-The ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
+Technically, the choice of PHP makes it easy to share functionality between the
+here described conversion scripts, and PHP-enabled web services, used to provide the Linked
+Data interface to the ChEMBL-RDF data, as is fast enough for practical uses. These scripts do
+not use a library for handling RDF data. While using such libraries increases the dependencies
+for the scripts, they also ensure the generation of syntactically correct RDF triples. Continued
+development of the ChEMBL-RDF scripts, however, has started using RDF libraries for this purpose
+specificaly for triple generation of compound names, removing code in the scripts to escape
+special characters, like double quotes.
+
+Finally, the ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
 exposing ChEMBL as RDF too (EW, PA). Various version may arise, exposing the data in different ontologies,

### 2013-03-06 article.tex
Explained why custom scripts are used, and how the data is practically converted
Added a note about automated validation

@@
 The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
-were then executed to generate RDF triples using a set of PHP scripts. These scripts are Open Source 
+were then executed to generate RDF triples using a set of PHP scripts. The scripts use SQL to
+query the database and output triples using print statements. Custom PHP scripts have been used
+as a light-weight soluton to provide a Linked Data API via a Apache webserver, as well as to
+create data files with triples. These scripts are Open Source 
 and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
@@
 provides end users with a high degree of confidence that the ChEMBL-RDF contains the same 
-core content as the source relational database.
+core content as the source relational database. The development of a validation test suite
+is being discussed but not yet implemented; testing criteria other then the here presented
+summarizing counts would change with each release due to curation events. Interactive,
+manual curation of the conversion process as employed by the authors already is more effective
+at this moment.
 

### 2013-03-06 article.tex
Stronger word to link the causes to the implications: the challenges are the aforementioned data sizes and complexity

@@
 The current scientific data deluge, in which datasets grow faster in size and complexity than scientists can keep up with,
-provides several new challenges for information systems. Scientists wish to 
+defines several new challenges for information systems. Scientists wish to 
 discover new, unique, and significant patterns in datasets that explain biological 

### 2013-03-06 article.tex
Removed the last 'indeed'
Listed the references at the first occurance of the resource too.

@@
 These links are shown in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
-using the complementary index and OpenMolecules RDF using InChI values. Protein links are given to Bio2RDF~\cite{Belleau2008}
+using the complementary index and OpenMolecules RDF using InChI values~\cite{Pence2010,Bradley2009,Willighagen2011}.
+Protein links are given to Bio2RDF~\cite{Belleau2008}
 using the UniProt identifier~\cite{TheUniProtConsortium2010}. Literature references are directed to CrossRef and Bio2RDF 
-using the DOIs and PubMed identifiers, respectively.
+using the DOIs and PubMed identifiers, respectively~\cite{Bilder2011}.
 
@@
 Much could be argued about the choice of ontologies, and the lack of an
-ontology in places where we decided to use custom classes and predicates. Indeed, the
+ontology in places where we decided to use custom classes and predicates. These
 choices have only partly been decided using ontological arguments. Most of the choices are based

### 2013-03-06 article.tex
We now mention Bio2RDF as early adopter project semantically linking scientific data sets

@@
 linkage of many scientific datasets that are both sparse and are growing independently from each
-other~\cite{Samwald2011,Williams2012}.
+other~\cite{Belleau2008,Samwald2011,Williams2012}.
 

### 2013-03-06 article.tex
Better reference for 'linked data'
The source format is the relational database

@@
 Using the right to download, modify, and redistribute the ChEMBL data, two independent teams
-have previously mapped the ChEMBL dataset to the Resource Description Framework (RDF), resulting
+have previously mapped the ChEMBL relational database to the Resource Description Framework (RDF), resulting
 in Chem2Bio2RDF~\cite{Chen2010} and ChEMBL-RDF~\cite{Willighagen2011}. RDF is a framework where
@@
 are represented by resources named with Universal Resource Identifiers (URI). When the URI can be resolved, 
-for example using the HTTP protocol, it creates \emph{linked data}~\cite{Samwald2011}. RDF is
+for example using the HTTP protocol, it creates \emph{linked data}~\cite{Bizer2008}. RDF is
 developed as an open standard by the World Wide Web Consortium, and provides a knowledge

### 2013-03-06 article.tex
For PRO now refer to the 2011 NAR paper

@@
 including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
-references and citations, and domain ontologies such the Protein Ontology\cite{Sidhu2006} and the Chemical Information
+references and citations, and domain ontologies such the Protein Ontology\cite{Natale2011} and the Chemical Information
 Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 

### 2013-03-06 article.tex
Note about the nature of this 'ad-hoc' ontology
Fixed the number of assays in the RDF to match the UU end point

@@
 However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
-an ad-hoc ontology~\cite{Willighagen2011}. The current version, instead, uses community-proposed
+an ad-hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
+in the relational database, and are not formally defined. The current version, instead, uses community-proposed
 ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
@@
 
-A similar SPARQL query, replacing \verb+chembl:Target+ with \verb+bfo:MaterialEntity+, returns 1,143,682 entities for which activities are captured in ChEMBL, and by using \verb+chembl:Assay+, the query returns the total of 617,681 ChEMBL assays.
+A similar SPARQL query, replacing \verb+chembl:Target+ with \verb+bfo:MaterialEntity+, returns 1,143,682 entities for which activities are captured in ChEMBL, and by using \verb+chembl:Assay+, the query returns the total of 617,676 ChEMBL assays.
 It should be noted that last query returns five entries less than the SQL database, because the

### 2013-03-06 article.tex
Cleaned up the figure captions

@@
 %\includegraphics[width=0.7\textwidth]{figs/relations}
-\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
-of common classes, while others are instances.}\label{f1}
+\caption{The various resource types found in the ChEMBL-RDF triples. Some entities are
+subclasses of common classes, using a rdfs:subClassOf predicate, such as
+target:t100122, while others are instances using the rdf:type predicate, such as
+assay:a17. The predicates between classes are also provided, showing how the
+resources are semantically linked.}\label{f1}
 \end{center}
@@
 %\includegraphics[width=0.45\textwidth]{figs/lodgraph}
-\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
-Edges are labeled by the predicates making the links.}\label{2}
+\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud
+versions of various external databases. Edges are labeled by the predicates making
+the links.}\label{2}
 }
@@
 		%\includegraphics[width=15cm]{citedin.png}
-		\caption[wee]{Screenshot of the CitedIn web service showing a paper cited 94671 times in the ChEMBL database.
+		\caption[wee]{Screenshot of the CitedIn web service showing a 2010
+		Nature paper cited 94671 times in the ChEMBL database.
 		   The \textit{Details} button on the webpage links to a ChEMBL webpage with detail on what parts of the database are
@@
 		%\includegraphics[width=15cm]{bioclipse-ds.png}
-		\caption[wee]{Screenshot from Bioclipse Decision Support with results from a ChemSpider + ChEMBL-RDF search.
-The top left canvas contains the query structure, in this case the drug Carbamezapine, the top right canvas shows the near neighbors in ChemSpider (via a similarStructure search) that contain ChEMBL-RDF data, the lower right shows the chemical structure for the selected compound in the top right canvas, and the lower left canvas shows the found interactions for this compound using ChEMBL-RDF.}
+		\caption[wee]{Screenshot from the Bioclipse Decision Support with results from a combined ChemSpider and ChEMBL-RDF search.
+The top left canvas contains the query structure, carbamezapine, and the top right canvas shows its near neighbors found in ChemSpider for which ChEMBL-RDF data exists.
+The lower right canvas shows the chemical structure selected in the top right canvas, and the lower left canvas shows the available activities in ChEMBL-RDF for this compound.}
 	\label{fig:bioclipse-ds}

### 2013-03-06 article.tex
Spell checked

@@
 query the database and output triples using print statements. Custom PHP scripts have been used
-as a light-weight soluton to provide a Linked Data API via a Apache webserver, as well as to
+as a light-weight solution to provide a Linked Data API via a Apache webserver, as well as to
 create data files with triples. These scripts are Open Source 
@@
 However, where the database content is well-defined, specific ontologies have been used.
-For example, where the database specifies literature from which data was extraced,
+For example, where the database specifies literature from which data was extracted,
 the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
@@
 development of the ChEMBL-RDF scripts, however, has started using RDF libraries for this purpose
-specificaly for triple generation of compound names, removing code in the scripts to escape
+specifically for triple generation of compound names, removing code in the scripts to escape
 special characters, like double quotes.

### 2013-03-26 article.tex
convert a stray reference to the old PRO citation

@@
         & http://purl.org/spar/cito/ \\
-obo / pro & OBO \& PRotein Ontology~\cite{Sidhu2006} \\
+obo / pro & OBO \& PRotein Ontology~\cite{Natale2011} \\
           & http://purl.obolibrary.org/obo/ \\

### 2013-03-26 article.tex
fix reviewer 1 comment 5 to change from 1.3 million to 1.1 million
add school for my affiliation

@@
 \iid(2)Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden \\
-\iid(3)University of Queensland, St Lucia, Qld 4072, Australia \\
+\iid(3)School of Information Technology and Electronic Engineering, University of Queensland, St Lucia, Qld 4072, Australia \\
 \iid(4)Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A. \\
@@
 one such system that supports only URIs of up to a certain length. Therefore, InChI-based links are only created for smaller molecules. 
-Almost 1.3 million links to OpenMolecules RDF were created in a similar manner to:
+Almost 1.1 million links to OpenMolecules RDF were created in a similar manner to:
 

### 2013-03-26 article.tex
Fix Uniform Resource Identifiers (was Universal)

@@
 Making data available as Linked Data using Resource Description Framework (RDF) promotes integration with other web resources.
-RDF documents can natively link to related data, and others can link back using Universal Resource Identifiers (URIs).
+RDF documents can natively link to related data, and others can link back using Uniform Resource Identifiers (URIs).
 RDF makes the data machine-readable and uses extensible vocabularies for additional information, making it easier
@@
 be given on any topic, and RDF therefore talks about \textit{things} very generically~\cite{Miller:04:RP}. Things in RDF
-are represented by resources named with Universal Resource Identifiers (URI). When the URI can be resolved, 
+are represented by resources named with Uniform Resource Identifiers (URI). When the URI can be resolved, 
 for example using the HTTP protocol, it creates \emph{linked data}~\cite{Bizer2008}. RDF is

### 2013-03-26 article.tex
advantage to advantages
descriptions to a description

@@
 \paragraph*{Conclusions:}
-We have illustrated the advantage of using open standards and ontologies to link the ChEMBL database
+We have illustrated the advantages of using open standards and ontologies to link the ChEMBL database
 to other databases. Using those links and the knowledge encoded in standards and ontologies, the ChEMBL-RDF
@@
 Another prominent example is drug discovery, in which a new unique chemical entity is designed or discovered 
-based on descriptions of its required biological properties. This process requires the effective 
+based on a description of its required biological properties. This process requires the effective 
 linkage of many scientific datasets that are both sparse and are growing independently from each

### 2013-03-26 article.tex
independently from to independent of

@@
 based on a description of its required biological properties. This process requires the effective 
-linkage of many scientific datasets that are both sparse and are growing independently from each
+linkage of many scientific datasets that are both sparse and are growing independent of each
 other~\cite{Belleau2008,Samwald2011,Williams2012}.

### 2013-03-26 article.tex
remove extra "the"
reword the data integration in the life sciences statement

@@
 The former requires scientists to import the data into a relational database, while the 
-latter limits the machine access to the data. ChEMBL also provides web services for programmatic
+latter limits machine access to the data. ChEMBL also provides web services for programmatic
 access, but these are limited to the specific types of queries which are encoded in the web service API.
@@
 developed as an open standard by the World Wide Web Consortium, and provides a knowledge
-representation framework that applies to many, if not all, domains, making it a successful 
-approach to data integration in the life sciences.
+representation framework that applies to many, if not all, domains, and has been used successfully 
+to integrate data for the life sciences.
 

### 2013-03-26 article.tex
emphasise that the linked RDF URIs are all linked data compatible to our knowledge

@@
 and ontologies used to map ChEMBL to RDF, along with new use cases, showing 
-links to other datasets using their RDF URIs to further support research in the life sciences.
+links to other datasets using their RDF linked data URIs to further support research in the life sciences.
 

### 2013-03-26 article.tex
reword "not changed much" sentence
reword are to is, as the subject of the sentence seems to be the ontology and not the concepts

@@
 has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
-Group~\cite{Marshall2012}. This process has not changed much since its original description.
+Group~\cite{Marshall2012}. This process is similar to its original description.
 
@@
 an ad-hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
-in the relational database, and are not formally defined. The current version, instead, uses community-proposed
+in the relational database, and is not formally defined. The current version, instead, uses community-proposed
 ontologies, making the RDF more interoperable with the RDF offerings of different databases. 

### 2013-03-26 article.tex
rearrange sentence referring to interoperability with other databases

@@
 in the relational database, and is not formally defined. The current version, instead, uses community-proposed
-ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
+ontologies, making the resulting RDF triples more interoperable with RDF provided by other life sciences databases. 
 The new triple dataset uses standard ontologies

### 2013-03-26 article.tex
reword statement about community alternative ontologies not yet available

@@
 Although many parts of the ChEMBL-RDF 13 dataset now use community based ontologies, some 
-terms from the previous ad-hoc ontology are still used because no community alternative is yet available. 
+terms from the previous ad-hoc ontology are still used when a community alternative is not yet available. 
 These terms were created under the 

### 2013-03-26 article.tex
reference the Open Definition as the definition for open ontologies
the vast majority of so called "open ontologies", including OBO and
Gene Ontology, fail this definition by either restricting use to
academics, and/or restricting modification
of ontologies without onerous conditions like renaming everything

@@
 many concepts are not mapped to existing ontologies, though also the lack of ontologies
-with a clear open license restrict the options.
+with a clear open license, as defined by the Open Definition~\cite{OpenDefinition}, restrict the options.
 

### 2013-03-26 article.tex
grammar for sentence defining drugs to be only recognised as such within a specified context
add FIXME note for a possibly ambiguous sentence
an authoritative
in particular, there are multiple DOI authoritative sources, so an is more correct and easier to read

@@
 It should also be noted that all drug-like entities in ChEMBL are not ontologically
-subclassing a drug class. Instead, we give the entities the role of drug: only within a
-particular context, they would be drug. If not used as medication, they are not drug.
+subclassing a drug class. Instead, we give the entities the role of drug, and only within a
+particular context would they be a drug. If not used as medication, they are not a drug.
 Moreover, the triples only attach this drug role to approved drugs, as defined by the
-ChEMBL database. All drug-like entities in the database do not have the role of drug.
+ChEMBL database. 
+% FIXME: What does the following exactly mean? It isn't clear whether it actually means:
+% Not all drug-like entities in the database have the role of drug
+All drug-like entities in the database do not have the role of drug.
 The role is triplified using the OBO and ChEBI ontologies in the following manner,
@@
 the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) are used to link
-to authoritative database, as outlined later in this paper.
+to an authoritative database, as outlined later in this paper.
 The properties are provided using the BIBO ontology, which is an open sourced ontology

### 2013-03-26 article.tex
make URLs render using tt instead of rm

@@
 \usepackage[utf8]{inputenc} %unicode support
-\urlstyle{rm}
+\urlstyle{tt}
 \usepackage{cite}
@@
 local machine, e.g. \url{http://localhost:8080/chembl/}, then a request for the
-article with identifier ``a31863'', \url{http://localhost:8080/-chembl/article/a31863},
+article with identifier ``a31863'', \url{http://localhost:8080/chembl/article/a31863},
 will be resolved from the database using the full original URI,

### 2013-03-26 article.tex
reword sentence about resolution process

@@
 article with identifier ``a31863'', \url{http://localhost:8080/chembl/article/a31863},
-will be resolved from the database using the full original URI,
+will be resolved using the webapp from the database using the full original URI,
 \url{http://linkedchemistry.info/chembl/activity/a31863}. If the

### 2013-03-26 article.tex
SPARQLs isn't as easy to read as Example SPARQL queries

@@
 The targets can also be linked to side effects via protein-protein interactions by searching across ChEMBL, HPRD and SIDER in
-Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
+Chem2Bio2RDF. Example SPARQL queries are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
 SPARQL endpoint~\cite{Chen2010}.

### 2013-03-26 article.tex
emphasise that the sparql query is a template and cannot therefore be directly run
new to novel

@@
 which is cited as the reference for a large number of activity datapoints and compounds in ChEMBL.
-Counting all objects that reference this paper is done by the following SPARQL query:
+Counting all objects that reference this paper is done by the following SPARQL query template:
 
@@
 ChEMBL data in a machine-readable manner using semantic web technologies. We showed here
-that the conversion into RDF allows a number of new integrations with other resources.
+that the conversion into RDF allows a number of novel integrations with other resources.
 It should be clear that these applications can be made using other technologies too, but

### 2013-03-26 article.tex
add comma to split sentence up

@@
 Some information in the databases has not yet been annotated using standard ontologies. 
-For example, the activity types in ChEMBL, such as ``IC50'', ``EC50'', and ``Potency'', are not exposed in the RDF dataset
+For example, the activity types in ChEMBL, such as ``IC50'', ``EC50'', and ``Potency'', are not exposed in the RDF dataset,
 meaning that it is not yet possible to reason or group over similar activity types.

### 2013-03-26 article.tex
add comma to split sentence up
grammar fix

@@
 meaning that it is not yet possible to reason or group over similar activity types.
-When the activity types are exposed queries could be constructed based on the knowledge 
+When the activity types are exposed, queries could be constructed based on the knowledge 
 that the IC$_{50}$ and K$_i$ end points are closely related. Similarly, measurement units are currently provided as
@@
 Finally, the ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
-exposing ChEMBL as RDF too (EW, PA). Various version may arise, exposing the data in different ontologies,
+exposing ChEMBL as RDF too (EW, PA). In the future various versions may arise, exposing the data in different ontologies,
 supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that

### 2013-04-14 article.tex
Applied Antonty's linguistic fixes

@@
 query the database and output triples using print statements. Custom PHP scripts have been used
-as a light-weight solution to provide a Linked Data API via a Apache webserver, as well as to
+as a lightweight solution to provide a Linked Data API via a Apache webserver, as well as to
 create data files with triples. These scripts are Open Source 
@@
 However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
-an ad-hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
+an ad hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
 in the relational database, and is not formally defined. The current version, instead, uses community-proposed
@@
 including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
-references and citations, and domain ontologies such the Protein Ontology\cite{Natale2011} and the Chemical Information
+references and citations, and domain ontologies such as the Protein Ontology\cite{Natale2011} and the Chemical Information
 Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
@@
 
-To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point is being
+To expose the ChEMBL-RDF data two approaches have been adopted. First, a SPARQL endpoint is being
 hosted at Uppsala University, using the Open-Source Edition of the Virtuoso software~\cite{Virtuoso}. Usage is free, but the
@@
 platform~\cite{kasabi} was used for a period of time, until that project was discontinued.
-The Uppsala SPARQL end point still uses resource URIs based on the Kasabi approach, following a similar
+The Uppsala SPARQL endpoint still uses resource URIs based on the Kasabi approach, following a similar
 URI pattern where linkedchemistry.info/chembl is replaced by data.kasabi.com/dataset/chembl-rdf.
@@
 Although many parts of the ChEMBL-RDF 13 dataset now use community based ontologies, some 
-terms from the previous ad-hoc ontology are still used when a community alternative is not yet available. 
+terms from the previous ad hoc ontology are still used when a community alternative is not yet available. 
 These terms were created under the 
@@
 We present here the updated ChEMBL-RDF, including a description of the mapping process 
-and the identification of links to other datasets. Because the primary purpose of of this
+and the identification of links to other datasets. Because the primary purpose of this
 paper is to expose the ChEMBL data as linked data, less focus has been placed on ontologically
@@
 subclass their respective matches in the CHEBI ontology (CHEBI\_50699
-and CHEBI\_7754). To ensure we can look up all entities for which activities
+and CHEBI\_7754). To ensure that we can look up all entities for which activities
 are reported, we subclass everything from the BFO 1.1 class MaterialEntity~\cite{Smith2004},
@@
 subclassing a drug class. Instead, we give the entities the role of drug, and only within a
-particular context would they be a drug. If not used as medication, they are not a drug.
+particular context would they be a drug. If not used as a medication, they are not a drug.
 Moreover, the triples only attach this drug role to approved drugs, as defined by the
@@
 provided via the CHEMINF formalism.
-The CHEMINF ontology has been used here because of it's flexibility in defining
+The CHEMINF ontology has been used here because of its flexibility in defining
 descriptor types and their origin and meaning. By allowing an ontology of descriptors,
-we enable comparing values between databases semantically. Differences in, for example,
-the SMILES may then be explained by the software used for that descriptor.
+we enable the comparison of values between databases semantically. Differences in, for example,
+the SMILES string may then be explained by the software used for that descriptor.
 
@@
 
-Similarly, properties of small molecules are available from ChEMBL, and as of ChEMBL-RDF 13
+Similarly, the properties of small molecules are available from ChEMBL, and as of ChEMBL-RDF 13
 these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
@@
 to an authoritative database, as outlined later in this paper.
-The properties are provided using the BIBO ontology, which is an open sourced ontology
+The properties are provided using the BIBO ontology, which is an open source ontology
 for bibliographic information, compatible with the CiTO ontology:
@@
 
-Each release of the ChEMBL database is accompanied by a set of release statistics, which help 
+Each release of the ChEMBL database is accompanied by a set of release statistics which help 
 provide a concise content overview of the release in question. The ChEMBL 13 release counts
@@
 A similar SPARQL query, replacing \verb+chembl:Target+ with \verb+bfo:MaterialEntity+, returns 1,143,682 entities for which activities are captured in ChEMBL, and by using \verb+chembl:Assay+, the query returns the total of 617,676 ChEMBL assays.
-It should be noted that last query returns five entries less than the SQL database, because the
+It should be noted that this last query returns five entries less than the SQL database, because the
 triple generation requires the assay type to be defined, which is not the case for those 
@@
 core content as the source relational database. The development of a validation test suite
-is being discussed but not yet implemented; testing criteria other then the here presented
-summarizing counts would change with each release due to curation events. Interactive,
+is being discussed but not yet implemented; testing criteria other then the summarizing counts
+presented here, would change with each release due to curation events. Interactive,
 manual curation of the conversion process as employed by the authors already is more effective
@@
 
-ChEMBL-RDF can be used to explore interrelated scientific datasets, and we here present five applications,
-describing how the SPARQL end point can be used, the combination of the ChEMBL data
+ChEMBL-RDF can be used to explore interrelated scientific datasets, and we present here five applications,
+describing how the SPARQL endpoint can be used, the combination of the ChEMBL data
 with other life sciences databases, the use of the bibliographic information in calculating citation
@@
 
-Using the SPARQL end point, it is easy to integrate data from ChEMBL into other web resources, such as CitedIn (http://www.citedin.org)~\cite{CitedIn2011}.
+Using the SPARQL endpoint, it is easy to integrate data from ChEMBL into other web resources, such as CitedIn (http://www.citedin.org)~\cite{CitedIn2011}.
 While the impact of a scientific publication is traditionally measured by frequency of citation in subsequent publications, this
@@
 
-In this study we extended Bioclipse DS with remote access to ChEMBL-RDF and ChemSpider. 
-This enables users browsing chemical structures in Bioclipse to look up near neighbors in ChemSpider 
+In this study we extended Bioclipse DS with remote access to both ChEMBL-RDF and ChemSpider. 
+This enables users browsing chemical structures in Bioclipse to look up nearest neighbors in ChemSpider 
 via the ChemSpider Web API (SOAP), and for the found compounds to be used to query ChEMBL-RDF for 
@@
 When the activity types are exposed, queries could be constructed based on the knowledge 
-that the IC$_{50}$ and K$_i$ end points are closely related. Similarly, measurement units are currently provided as
+that the IC$_{50}$ and K$_i$ endpoint are closely related. Similarly, measurement units are currently provided as
 string literals, making it difficult to compare concentrations when some are given in nM and others in mM.
@@
 The authors acknowledge A. L\"ovgren at the BMC Computing Department at Uppsala University for his
-support in hosting the SPARQL end point. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE. 
+support in hosting the SPARQL endpoint. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE. 
 JH acknowledges funding from the EU-OPENSCREEN project. 

### 2013-04-14 article.tex
Longer version of Ola's acks
To reduce potential confusion, the ChEMBL team withdrew as authors, to make clear that this version is not to be confused with the RDF they will start distributing

@@
 Janna~Hastings$^5$\email{Janna Hastings - hastings@ebi.ac.uk}\and
-Mark~Davies$^6$\email{Mark Davies - mdavies@ebi.ac.uk, \url{http://twitter.com/mdavies12}}\and
-John~P.~Overington$^6$\email{John P. Overington - jpo@ebi.ac.uk}\and
-Anna~Gaulton$^6$\email{Anna Gaulton - agaulton@ebi.ac.uk}\and
-Bin~Chen$^7$\email{Bin Chen - binchen@indiana.edu, \url{http://twitter.com/binchenindiana}}\and
-David~J.~Wild$^7$\email{David J. Wild - djwild@indiana.edu, \url{http://twitter.com/davidjohnwild}}
+Bin~Chen$^6$\email{Bin Chen - binchen@indiana.edu, \url{http://twitter.com/binchenindiana}}\and
+David~J.~Wild$^6$\email{David J. Wild - djwild@indiana.edu, \url{http://twitter.com/davidjohnwild}}
 }
@@
 \iid(5)Cheminformatics and Metabolism, European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\
-\iid(6)Chemogenomics, European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\ 
-\iid(7)School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.
+\iid(6)School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.
 }
@@
 OS, AJW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
-JH supported the project with CHEMINF representations. PA integrated ChEMBL-RDF into Bio2RDF.
 JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
@@
 The authors acknowledge A. L\"ovgren at the BMC Computing Department at Uppsala University for his
-support in hosting the SPARQL endpoint. OS acknowledges funding from the Swedish VR (2011-6129) and eSSENCE. 
-JH acknowledges funding from the EU-OPENSCREEN project. 
-The data in ChEMBL is made available by funding from the Wellcome Trust [086151/Z/08/Z].
+support in hosting the SPARQL endpoint. OS acknowledges funding from the Swedish VR-2011-6129, and the Swedish strategic research program eSSENCE.
 

### 2013-04-20 article.tex
Removed a confusing line which was not adding new information
Applied the (minor) changes by Janna

@@
 However, where the database content is well-defined, specific ontologies have been used.
-For example, where the database specifies literature from which data was extracted,
+For example, where the database specifies literature from which data were extracted,
 the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
-compatibility with other Linked Data projects, rather than formal ontological arguments.
+compatibility with other Linked Data projects, rather than formal ontological considerations.
 We have not observed ontological contradictions based on these choices.
@@
 ChEMBL database.
-% FIXME: What does the following exactly mean? It isn't clear whether it actually means:
-% Not all drug-like entities in the database have the role of drug
-All drug-like entities in the database do not have the role of drug.
 The role is triplified using the OBO and ChEBI ontologies in the following manner,
@@
 Much could be argued about the choice of ontologies, and the lack of an
-ontology in places where we decided to use custom classes and predicates. These
-choices have only partly been decided using ontological arguments. Most of the choices are based
+ontology in places where we decided to use custom classes and predicates. Most of the choices are based
 on pragmatic approaches to enable solutions to the range of use cases we describe here. It should be noted,
@@
 
-Finally, the ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
+Finally, the ChEMBL project will privide RDF versions of ChEMBL, while others continue
 exposing ChEMBL as RDF too (EW, PA). In the future various versions may arise, exposing the data in different ontologies,
@@
 OS, AJW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
-JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
 BC, DW contributed the Chem2Bio2RDF use cases.
@@
         & http://purl.org/ontology/bibo/ \\
-chebi   & Chemical Entities of Biological Interest~\cite{DeMatos2010} \\
+chebi   & Chemical Entities of Biological Interest~\cite{Hastings2013} \\
         & http://purl.org/obo/owl/CHEBI\# \\
